---
title: "Exploring VTEs in ONCOTHROMB12-01 cohort"
author: "Bárbara Lobato Delgado"
output:
  html_document:
    df_print: paged
---


```{r, message=FALSE}
# Load packages
pacman::p_load(
  an9elproject,
  tidyverse,
  magrittr, 
  gridExtra, 
  ggpval, 
  ggstatsplot,
  ggsignif,
  RColorBrewer, 
  lubridate, 
  survival, 
  survminer, 
  cmprsk, 
  install = FALSE, update = FALSE
  )
```


# Load data set

```{r}
# Load ONCOTHROMB1 DB
# oncoth1 = get_project("oncothr1", version = "0.0.8003") 
load("/mnt/ir-bioinf02/home/blobato/oncothromb01/oncothr1_versions/oncothr1.RData_0.0.8003")
oncoth1 = obj
```


```{r}
# Get number of patients with VTE
n_patients_vte = unname(table(oncoth1$data$patient_group)[1])
print(n_patients_vte)
```


# Data exploration and visualisation

```{r}
# Get features of interest for the description of VTE episodes/events
VTE_data = oncoth1$data %>%
  select(id, 
         patient_group, 
         n_appointment_patient_became_case, 
         primary_tumor_simplified,
         date_cancer_dx,
         n_all_vte,
         n_vte_before_entering_study,
         n_total_vte_along_study,
         n_total_vte_recurrence,
         date_1st_vte_all,
         # Time until VTE (months) since cancer diagnosis
         tu_1st_vte_all_from_cancer_diagnosis_months,
         tu_1st_vte_all_from_cancer_diagnosis_days,
         tu_2nd_vte_all_from_cancer_diagnosis_months,
         tu_3rd_vte_all_from_cancer_diagnosis_months,
         tu_4th_vte_all_from_cancer_diagnosis_months,
         # Corrected times (to avoid having two different events at the same time)
         tu_1st_vte_all_from_cancer_diagnosis_months_corrected, 
         tu_2nd_vte_all_from_cancer_diagnosis_months_corrected,
         tu_3rd_vte_all_from_cancer_diagnosis_months_corrected,
         tu_4th_vte_all_from_cancer_diagnosis_months_corrected,
         tu_1st_rec_vte_from_cancer_diagnosis_months) %>%
  # Get patients with VTE
  filter(patient_group == "Case")
```


```{r}
# Check that the number of patients with VTE and df dimensions check out
stopifnot(dim(VTE_data)[1] == n_patients_vte)
```


```{r}
n_VTEs = sum(VTE_data$n_all_vte)
print(paste("There is a total of", n_VTEs, "recorded VTE events."))
```



## VTE events per month 

I am gonna ignore the distinction of incidental and symptomatic for now.  

```{r}
# Get time until VTE (months) for all VTEs
VTE_events_by_month = VTE_data %>%
  select(
    tu_1st_vte_all_from_cancer_diagnosis_months, 
    tu_2nd_vte_all_from_cancer_diagnosis_months, 
    tu_3rd_vte_all_from_cancer_diagnosis_months,
    tu_4th_vte_all_from_cancer_diagnosis_months
    ) %>%
  pivot_longer(everything(), values_to = "month") %>%
  filter(!is.na(month)) %>%
  group_by(month) %>%
  summarise(n_VTEs = n()) %>%
  mutate(percentage = round(n_VTEs/sum(n_VTEs), digits = 2))
```


```{r}
# Sanity check
stopifnot(sum(VTE_events_by_month$n_VTEs) == 120 - 1) # there is one VTE event with unknown date and symptomatology
```


```{r}
# Collapsing information of months -2, -1 and 0 into one
VTE_events_by_month_corrected <- VTE_events_by_month %>%
  mutate(across(2:3, ~ ifelse(row_number() == 3, sum(VTE_events_by_month[1:3, cur_column()]), .))) %>%
  slice(-c(1:2))
```


```{r}
# Sanity check
stopifnot(sum(VTE_events_by_month_corrected$n_VTEs) == 120 - 1) # there is one VTE event with unknown date and symptomatology
```


```{r}
# Barplot of by CAT over time
ggplot(
  data = VTE_events_by_month, 
  aes(x = month, y = percentage)) +
  geom_bar(stat = "identity", width = 0.8, color = "grey", fill = "lightgrey") +
  geom_line() +
  geom_point() +
  ggtitle("CAT occurrence by month with respect to cancer diagnosis date") + 
  xlab("Time (months)") + 
  ylab("Proportion of CAT events") + 
  scale_x_continuous(breaks = seq(from = -2, to = 18, by = 1)) + 
  scale_y_continuous(labels = scales::percent, breaks = scales::pretty_breaks(n = 8)) +
  theme_bw()+
  theme(plot.title = element_text(hjust = 0.5))

# Save plot
ggsave(
  filename = "../graphs/CAT_events_by_month_since_cancer_dx.jpeg", 
  width = 10,
  height = 6
)
```


```{r}
# Barplot of CAT over time (events prior to cancer diagnosis are collapsed into month zero)
ggplot(
  data = VTE_events_by_month_corrected, 
  aes(x = month, y = percentage)) +
  geom_bar(stat = "identity", width = 0.8, color = "grey", fill = "lightgrey") +
  geom_line() +
  geom_point() +
  ggtitle("CAT occurrence by month with respect to cancer diagnosis date") + 
  xlab("Time (months)") + 
  ylab("Proportion of CAT events") + 
  scale_x_continuous(breaks = seq(from = 0, to = 18, by = 1)) + 
  scale_y_continuous(labels = scales::percent, breaks = scales::pretty_breaks(n = 8)) +
  theme_bw()+
  theme(plot.title = element_text(hjust = 0.5))

# Save plot
ggsave(
  filename = "../graphs/CAT_events_by_month_since_cancer_dx_corrected.jpeg", 
  width = 10,
  height = 6
)
```


## Patients affected by CAT per month 

```{r}
# Transform data to count patients affected by CAT
patients_with_CAT_by_month = VTE_data %>% 
  select(
    id, 
    n_all_vte,
    # We only need the first VTE event
    tu_1st_vte_all_from_cancer_diagnosis_months
  ) %>%
  arrange(id) %>%
  # Patients with VTE prior to cancer diagnosis will be assigned time zero
  mutate(tu_1st_vte_all_from_cancer_diagnosis_months = case_when(
    tu_1st_vte_all_from_cancer_diagnosis_months <= 0 ~ 0, 
    TRUE ~ tu_1st_vte_all_from_cancer_diagnosis_months
  )) %>%
  rename(month_VTE_occurrence = tu_1st_vte_all_from_cancer_diagnosis_months) %>%
  # Count VTE ocurrences each month
  count(month_VTE_occurrence) %>%
  mutate(percentage = round(n/n_patients_vte, digits = 2))
```


```{r}
# Sanity check
stopifnot(sum(patients_with_CAT_by_month$n) == 88) 
```


```{r}
# Barplot with patients affected by CAT over time
ggplot(
  data = patients_with_CAT_by_month, 
  aes(x = month_VTE_occurrence, y = percentage)) +
  geom_bar(stat = "identity", width = 0.8, color = "grey", fill = "lightgrey") +
  geom_line() +
  geom_point() +
  ggtitle("Patients affected by CAT over time") + 
  xlab("Time (months)") + 
  ylab("% of patients") +
  scale_x_continuous(breaks = seq(from = 0, to = 18, by = 1)) + 
  scale_y_continuous(labels = scales::percent, breaks = scales::pretty_breaks(n = 8)) +
  theme_bw()+
  theme(plot.title = element_text(hjust = 0.5))

# Save plot
ggsave(
  filename = "../graphs/patients_with_CAT_by_month_since_cancer_dx_corrected.jpeg",
  width = 10,
  height = 6
)
```


## Frequency of VTE episode's type by location [corregir]


```{r}
# How many combinations of VTE events by location exist?
VTE_locs = c("CRT", "DVT", "SVT", "PTE", "VT")
combinations = combn(VTE_locs, 2)
print(combinations)
```


```{r}
# Get information on patients with VTE and the location
VTE_location = oncoth1$data %>%
  filter(patient_group == "Case") %>%
  select(
    id,
    primary_tumor_simplified,
    n_all_vte,
    n_total_crt_all,
    n_total_dvt_all, 
    n_total_svt_all,
    n_total_pte_all, 
    n_total_visceral_thromb_all
    ) 

# Check that total number of VTEs is the sum of the different types of VTE by location
stopifnot(VTE_location$n_all_vte == VTE_location$n_total_crt_all + VTE_location$n_total_dvt_all + VTE_location$n_total_svt_all + VTE_location$n_total_pte_all + VTE_location$n_total_visceral_thromb_all)

# Count number of events by location
VTE_location %<>%
  # Only CRTs
  mutate(n_crt = case_when(
    n_total_crt_all > 0 &
    n_total_dvt_all == 0 &
    n_total_svt_all == 0 &
    n_total_visceral_thromb_all == 0 &
    n_total_pte_all == 0 ~ 1, 
    TRUE ~ 0
  )) %>%
  # Only DVT
  mutate(n_dvt = case_when(
    n_total_crt_all == 0 &
    n_total_dvt_all > 0 &
    n_total_svt_all == 0 &
    n_total_visceral_thromb_all == 0 &
    n_total_pte_all == 0 ~ 1, 
    TRUE ~ 0
  )) %>%
  # Only VT
  mutate(n_vt = case_when(
    n_total_crt_all == 0 &
    n_total_dvt_all == 0 & 
    n_total_svt_all == 0 &
    n_total_visceral_thromb_all > 0 &
    n_total_pte_all == 0 ~ 1, 
    TRUE ~ 0
  )) %>%
  # Only SVT 
  mutate(n_svt = case_when(
    n_total_crt_all == 0 &
    n_total_dvt_all == 0 & 
    n_total_svt_all > 0 &
    n_total_visceral_thromb_all == 0 &
    n_total_pte_all == 0 ~ 1, 
    TRUE ~ 0
  )) %>%
  # Only PTE 
  mutate(n_pte = case_when(
    n_total_crt_all == 0 &
    n_total_dvt_all == 0 & 
    n_total_svt_all == 0 &
    n_total_visceral_thromb_all == 0 &
    n_total_pte_all > 0 ~ 1, 
    TRUE ~ 0
  )) %>%
  # DVT + PE
  mutate(n_dvt_pte = case_when(
    n_total_crt_all == 0 &
    n_total_dvt_all > 0 & 
    n_total_svt_all == 0 &
    n_total_visceral_thromb_all == 0 &
    n_total_pte_all > 0 ~ 1, 
    TRUE ~ 0
  )) %>%
  # DVT + VT
  mutate(n_dvt_vt = case_when(
    n_total_crt_all == 0 &
    n_total_dvt_all > 0 &
    n_total_svt_all == 0 &
    n_total_visceral_thromb_all > 0 &
    n_total_pte_all == 0 ~ 1, 
    TRUE ~ 0
  )) %>%
  # DVT + CRT
  mutate(n_dvt_crt = case_when(
    n_total_crt_all > 0 &
    n_total_dvt_all > 0 &
    n_total_svt_all == 0 &
    n_total_visceral_thromb_all == 0 &
    n_total_pte_all == 0 ~ 1, 
    TRUE ~ 0
  )) %>%
  # DVT + SVT
  mutate(n_dvt_svt = case_when(
    n_total_crt_all == 0 &
    n_total_dvt_all > 0 &
    n_total_svt_all > 0 &
    n_total_visceral_thromb_all == 0 &
    n_total_pte_all == 0 ~ 1, 
    TRUE ~ 0
  )) %>%
  # PE + VT
  mutate(n_pte_vt = case_when(
    n_total_crt_all == 0 &
    n_total_dvt_all == 0 &
    n_total_svt_all == 0 &
    n_total_visceral_thromb_all > 0 &
    n_total_pte_all > 0 ~ 1, 
    TRUE ~ 0
  )) %>%
  # PE + CRT
  mutate(n_pte_crt = case_when(
    n_total_crt_all > 0 &
    n_total_dvt_all == 0 &
    n_total_svt_all == 0 &
    n_total_visceral_thromb_all == 0 &
    n_total_pte_all > 0 ~ 1, 
    TRUE ~ 0
  )) %>%
  # PE + SVT
  mutate(n_pte_svt = case_when(
    n_total_crt_all == 0 &
    n_total_dvt_all == 0 &
    n_total_svt_all > 0 &
    n_total_visceral_thromb_all == 0 &
    n_total_pte_all > 0 ~ 1, 
    TRUE ~ 0
  )) %>%
  # CRT + SVT
  mutate(n_crt_svt = case_when(
    n_total_crt_all > 0 &
    n_total_dvt_all == 0 &
    n_total_svt_all > 0 &
    n_total_visceral_thromb_all == 0 &
    n_total_pte_all == 0 ~ 1, 
    TRUE ~ 0
  )) %>%
  # CRT + VT
  mutate(n_crt_vt = case_when(
    n_total_crt_all > 0 &
    n_total_dvt_all == 0 &
    n_total_svt_all == 0 &
    n_total_visceral_thromb_all > 0 &
    n_total_pte_all == 0 ~ 1, 
    TRUE ~ 0
  )) %>%
  # VT + SVT
  mutate(n_vt_svt = case_when(
    n_total_crt_all == 0 &
    n_total_dvt_all == 0 &
    n_total_svt_all > 0 &
    n_total_visceral_thromb_all > 0 &
    n_total_pte_all == 0 ~ 1, 
    TRUE ~ 0
  )) %>%
  # We don't do the combination for groups of three for all events, only for those
  # we observe in data
  # DVT + PE + SVT
  mutate(n_dvt_pte_svt = case_when(
    n_total_crt_all == 0 &
    n_total_dvt_all > 0 & 
    n_total_svt_all > 0 &
    n_total_visceral_thromb_all == 0 &
    n_total_pte_all > 0 ~ 1, 
    TRUE ~ 0
  )) %>%
  # DVT + PE + VT
  mutate(n_dvt_pte_vt = case_when(
    n_total_crt_all == 0 &
    n_total_dvt_all > 0 & 
    n_total_svt_all == 0 &
    n_total_visceral_thromb_all > 0 &
    n_total_pte_all > 0 ~ 1, 
    TRUE ~ 0
  ))
```


```{r}
# Total number of VTEs
stopifnot(sum(VTE_location$n_all_vte) == n_VTEs)
```


```{r}
print(paste("A total of", nrow(VTE_location), "patients had", n_VTEs, "VTEs"))
```


```{r}
# Count frequency of patients with one or more types of VTE by location
summary_n_patients_with_VTE_location = VTE_location %>%
  select(
    n_crt, 
    n_dvt, 
    n_vt, 
    n_svt, 
    n_pte, 
    n_dvt_pte, 
    n_dvt_vt, 
    n_dvt_crt,
    n_dvt_svt, 
    n_pte_vt, 
    n_pte_crt, 
    n_pte_svt,
    n_crt_svt,
    n_crt_vt, 
    n_vt_svt, 
    n_dvt_pte_svt, 
    n_dvt_pte_vt
    ) %>%
  summarise(across(everything(), sum, na.rm = TRUE)) %>%
  t() %>%
  as.data.frame() %>%
  rownames_to_column(var = "number_patients_with_vte_by_loc") %>%
  rename(freq_abs = V1) %>%
  arrange(desc(freq_abs)) %>%
  mutate(freq_rel = round(freq_abs/sum(freq_abs) * 100, digits = 1)) %>%
  # Filter out events with frequency zero
  filter(freq_abs != 0)

# Check that there is 88 patients accounted for
stopifnot(sum(summary_n_patients_with_VTE_location$freq_abs) == n_patients_vte)
```


```{r}
# jpeg("../graphs/paper/barplot_VTE_location_corrected.jpeg", width = 6.5, height = 4, units = 'in', res = 700)
# —
mycolors = colorRampPalette(brewer.pal(11, "Paired"))(11)

# display.brewer.all()
# my_palette <- colorRampPalette(c("cyan", "lightgreen"))
# colors <- my_palette(11)

ggplot(
  data = summary_n_patients_with_VTE_location, 
  aes(x = reorder(number_patients_with_vte_by_loc, -freq_abs), 
      y = freq_abs,
      fill = number_patients_with_vte_by_loc)) +  
  geom_col() + # color = "grey"
  geom_text(
    aes(label = paste0(freq_rel, "%")), 
    position = position_dodge(width=0.9), 
    vjust = -0.25, 
    color = "#36454F"
    ) + 
  theme_bw() +
  ggtitle("Patients with VTE types by location") +
  ylab("Number and % of patients") +
  theme(plot.title = element_text(hjust = 0.5),
        legend.position = "none", 
        #axis.text.x = element_text(angle = 30, vjust = 1, hjust = 1),
        axis.title.x = element_blank()) +
  scale_y_continuous(
    limits = c(0, 30),
    breaks = c(0, 5, 10, 15, 20, 25, 30),
    #labels = function(x) paste0(x,"%")
    ) + 
  scale_fill_manual(values = mycolors) + 
  scale_x_discrete(
    labels = c("VT",
               "PE", 
               "DVT", 
               "DVT\n&\nPE",
               "CRT",
               "DVT\n&\nVT", 
               "PE\n&\nVT",
               "DVT\n&\nCRT",
               "DVT\n&\nPE\n&\nVT", 
               "SVT", 
               "DVT\n&\nPE\n&\nSVT"))

# Save plot
ggsave(
  filename = "../graphs/oncoth1_n_patients_VTE_location.jpeg",
  width = 6, 
  height = 4,
  units = "in"
  )

# dev.off()
```


## VTE symptomatology by cancer type

```{r}
# Dataframe with incidental/symptomatic VTEs
VTE_symptomatology = oncoth1$data %>% 
  filter(patient_group == "Case") %>%
  select(
    id, 
    primary_tumor_simplified,
    n_all_vte,
    # cancer_concomitant_vte_types, 
    # cancer_concomitant_type_vte_event,          
    cancer_concomitant_crt_type_vte,
    cancer_concomitant_1st_ledvt_type_vte,
    cancer_concomitant_2nd_ledvt_type_vte,
    # dvt_1st_type_prior_study,
    # dvt_2nd_type_prior_study,
    cancer_concomitant_uedvt_type_vte,
    cancer_concomitant_pte_type_vte,
    cancer_concomitant_svt_type_vte,
    # cancer_concomitant_visceral_thromb_type_vte,
    # cancer_concomitant_other_thromb_type_vte,
    eval_crt_type_vte_along_study,
    crt_type_recurrence,
    eval_ledvt_type_vte_along_study,
    eval_uedvt_type_vte_along_study,
    # eval_pte_type_vte_along_study,              
    eval_svt_type_vte_along_study,
    # eval_visceral_thromb_type_vte_along_study,
    # eval_other_thromb_type_vte_along_study,      
    # dvt_type_along_study,
    dvt_type_recurrence,                         
    # pte_type_prior_study,                       
    pte_type_along_study,                        
    pte_type_recurrence,                        
    visceral_thromb_type_prior_study,            
    visceral_thromb_type_along_study,           
    visceral_thromb_type_recurrence
  ) %>%
  # Count the number of times string "Incidental" appears in multiple columns
  rowwise() %>%
  mutate(count_inc = sum(c_across(where(is.character) | where(is.factor)) == "Incidental", na.rm = TRUE), .after = n_all_vte) %>%
  # Count the number of times string "Symptomatic" appears in multiple columns
  mutate(count_sympt = sum(c_across(where(is.character) | where(is.factor)) == "Symptomatic", na.rm = TRUE), .after = count_inc) 
```


```{r}
# Patients where counts do not match
# Patient id = 258 will show here
# Patient id = 258 had 1 VT of unknown symptomatology in 2015-10-22
# xx = VTE_symptomatology_count[!(VTE_symptomatology_count$n_all_vte == VTE_symptomatology_count$count_inc + VTE_symptomatology_count$count_sympt), ]

# Patient (one by one)
# yy = oncoth1$data %>% filter(id == 397)


# Check that total number of VTEs is the same as the sum of incidental and symptomatic events
# Patient id = 258 had 1 VT of unknown symptomatology in 2015-10-22
# stopifnot(VTE_symptomatology_count$n_all_vte == VTE_symptomatology_count$count_inc + VTE_symptomatology_count$count_sympt)
```


```{r}
# Number of incidental and symptomatic events per cancer type
VTE_symptomatology_by_cancer_type = VTE_symptomatology %>%
  select(id, primary_tumor_simplified, count_inc, count_sympt) %>%
  # Count events by cancer type
  group_by(primary_tumor_simplified) %>%
  summarise(total_count_inc = sum(count_inc),
            total_count_sympt = sum(count_sympt)) 

# Count events for all cancer types
VTE_symptomatology_by_cancer_type %<>%
  bind_rows(
    data.frame(primary_tumor_simplified = "All",
               total_count_inc = sum(VTE_symptomatology_by_cancer_type$total_count_inc),
               total_count_sympt = sum(VTE_symptomatology_by_cancer_type$total_count_sympt))
  ) %>%
  # Calculate relative frequencies
  rename(abs_freq_inc = total_count_inc,
         abs_freq_sympt = total_count_sympt) %>%
  rowwise() %>%
  mutate(across(starts_with("abs_freq_"), list(rel_freq = ~./sum(c_across(starts_with("abs_freq_")))), .names = "rel_freq_{.col}")) %>%
  # Rename columns
  rename(rel_freq_inc = rel_freq_abs_freq_inc, 
         rel_freq_sympt = rel_freq_abs_freq_sympt) %>%
  # Calculate percentage
  mutate(rel_freq_inc = round(rel_freq_inc * 100, digits = 1)) %>%
  mutate(rel_freq_sympt = round(rel_freq_sympt * 100, digits = 1)) 

# Show table
VTE_symptomatology_by_cancer_type %<>%
  pivot_longer(cols = contains("freq"), 
               names_to = "vte_type") %>%
  mutate(VTE_symptomatology = ifelse(grepl("_inc", vte_type), "Incidental", "Symptomatic")) %>%
  mutate(abs_freq = ifelse(grepl("abs_freq", vte_type), as.numeric(gsub("[^0-9.]", "", value)), NA)) %>%
  mutate(rel_freq = ifelse(grepl("rel_freq", vte_type), as.numeric(gsub("[^0-9.]", "", value)), NA)) %>%
  select(-c(vte_type, value)) %>%
  group_by(primary_tumor_simplified, VTE_symptomatology) %>%
  mutate(abs_freq = coalesce(abs_freq, na.omit(abs_freq)[1]),
         rel_freq = coalesce(rel_freq, na.omit(rel_freq)[1])) %>%
  distinct()
```


```{r}
# Create dataframe with absolute frequencies for incidental and symptomatic events
# Tests will be applied
CAT_symp_cancer_type = VTE_symptomatology_by_cancer_type %>% 
  select(-rel_freq) %>%
  pivot_wider(names_from = VTE_symptomatology, values_from = abs_freq, values_fill = 0) %>%
  ungroup() %>%
  as.data.frame() 

rownames(CAT_symp_cancer_type) = CAT_symp_cancer_type$primary_tumor_simplified

CAT_symp_cancer_type %<>%
  select(-primary_tumor_simplified)
```


```{r}
# Fisher test to find differences in incidental/symptomatic VTEs across cancer types
fisher.test(CAT_symp_cancer_type)
```

There is an statistically significant difference in proportion of incidental and symptomatic VTEs across cancer types. 

```{r}
# Create dataframe from contingency table with as many observations as VTE types per cancer 
# This will be used to inspect differences in proportions with ggstatsplot library
x <- c()
for (row in rownames(CAT_symp_cancer_type)) {
  for (col in colnames(CAT_symp_cancer_type)) {
    x <- rbind(x, matrix(rep(c(row, col), CAT_symp_cancer_type[row, col]), ncol = 2, byrow = TRUE))
  }
}
VTE_sympt_data_longer <- as.data.frame(x)
colnames(VTE_sympt_data_longer) <- c("cancer_type", "vte_type")
```


```{r}
# Do Fisher's exact test to obtain p-value
fisher_test_VTE_sympt_by_cancer_type <- fisher.test(table(VTE_sympt_data_longer))

# Do barplot with several statistical test
ggbarstats(
  VTE_sympt_data_longer, vte_type, cancer_type,
  digits.perc = 1,
  label = "both",
  proportion.test = TRUE,
  results.subtitle = TRUE, 
  bf.message = FALSE
)
```


```{r}
# Calculate proportion test for each cancer type
proportion_test_VTE_sympt_cancer_type <- VTE_sympt_data_longer %>%
  group_by(cancer_type) %>%
  summarise(
    test_result = list(prop.test(table(vte_type))),
    .groups = 'drop'
  ) %>%
  mutate(
    p_value = map(test_result, ~ format.pval(.x$p.value, digits = 3))
  )
```


```{r}
# Barplot of incidental and symptomatic VTEs for each cancer type and all types
ggplot(
  data = VTE_symptomatology_by_cancer_type, 
  aes(x = primary_tumor_simplified, y = abs_freq, fill = VTE_symptomatology)) + 
  geom_col(position = "dodge") + 
  geom_text(
    aes(label = paste0(rel_freq, "%")), 
    position = position_dodge(width = 0.9), 
    vjust = -0.25, 
    color = "#36454F", 
    size = 3
    ) + 
  geom_vline(
    xintercept = 1.5,
    linetype = "dotted",
    color = "grey",
    size = 0.5) +
  geom_signif(
    comparisons = list(c("Incidental", "Symptomatic")),
    annotations = proportion_test_VTE_sympt_cancer_type$p_value,
    y_position = rep(75, length(proportion_test_VTE_sympt_cancer_type$p_value)),
    tip_length = 0.03,
    textsize = 3) +
  theme_bw() +
  ggtitle("VTE symptomatology by cancer type") +
  guides(fill = guide_legend(title = "")) + 
  ylab("Number of VTEs") +
  theme(plot.title = element_text(hjust = 0.5),
        axis.title.x = element_blank()) +
  scale_y_continuous(
    limits = c(0, 75),
    breaks = seq(from = 0, to = 75, by = 10)) +
  scale_fill_brewer(palette = "Set4") 

# # Save plot
# ggsave(
#   filename = "../graphs/oncoth1_n_VTEs_symptoms_by_cancer_type.jpeg",
#   width = 6, 
#   height = 4,
#   units = "in"
#   )
```





```{r}
aa
```


## Moment when CAT appears according to cancer type 

### VTE events

```{r}
# Count number of VTE episodes prior or concomitant to cancer diagnosis by cancer type
cat_prior_or_concomitant_by_cancer_type = VTE_data %>%
  group_by(primary_tumor_simplified) %>%
  summarise(n_vte_prior_or_concomitant = sum(n_vte_before_entering_study))
              
              
# Check all VTEs before study begins are accounted for
stopifnot(
  sum(cat_prior_or_concomitant_by_cancer_type$n_vte_prior_or_concomitant) ==
    # Sum of time until VTE that happened before month zero
    sum(VTE_data$tu_1st_vte_all_from_cancer_diagnosis_months <= 0, na.rm = TRUE) +
    sum(VTE_data$tu_2nd_vte_all_from_cancer_diagnosis_months <= 0, na.rm = TRUE) + 
    sum(VTE_data$tu_3rd_vte_all_from_cancer_diagnosis_months <= 0, na.rm = TRUE) + 
    sum(VTE_data$tu_4th_vte_all_from_cancer_diagnosis_months<= 0, na.rm = TRUE)
  )

cat_prior_or_concomitant_by_cancer_type
```


```{r}
# Count number of VTE episodes that are recurrences by cancer type
cat_recurrences_by_cancer_type = VTE_data %>%
  group_by(primary_tumor_simplified) %>%
  summarise(n_vte_recurrence = sum(n_total_vte_recurrence))

# sum(!is.na(tu_1st_rec_vte_from_cancer_diagnosis_months), na.rm = TRUE)

# Check all VTEs before study begins are accounted for
# stopifnot(
#   sum(cat_recurrences_by_cancer_type$n_vte_recurrence) ==
#     # Count non-NA 
#     sum(!is.na(VTE_data$tu_1st_rec_vte_from_cancer_diagnosis_months), na.rm = TRUE)
#   )

cat_recurrences_by_cancer_type
```


```{r}
# Count number of VTE episodes during study by cancer type
cat_during_study_by_cancer_type = VTE_data %>%
  group_by(primary_tumor_simplified) %>%
  summarise(n_vte_during_study = sum(n_total_vte_along_study))
  
  
# Check all VTEs during study begins are accounted for
# This not correct! Do it right!
# stopifnot(
#   sum(cat_during_study_by_cancer_type$n_vte_during_study) ==
#     # Sum variables of time until VTE that happened after 
#     sum(VTE_data$tu_1st_vte_all_from_cancer_diagnosis_months > 0, na.rm = TRUE) +
#     sum(VTE_data$tu_2nd_vte_all_from_cancer_diagnosis_months > 0, na.rm = TRUE) +
#     sum(VTE_data$tu_3rd_vte_all_from_cancer_diagnosis_months > 0, na.rm = TRUE) +
#     sum(VTE_data$tu_4th_vte_all_from_cancer_diagnosis_months > 0, na.rm = TRUE) +
#     sum(VTE_data$tu_5th_vte_all_from_cancer_diagnosis_months > 0, na.rm = TRUE) -
#     sum(VTE_data$n_total_vte_recurrence)
#   )

# Careful! Recurrences are included here!
# We have to substract the number of recurrences calculated prior
# cat_during_study_by_cancer_type$n_vte_during_study - cat_recurrences_by_cancer_type$n_vte_recurrence

cat_during_study_by_cancer_type
```


```{r}
# Put together number of episodes by moment of occurrence
cat_moment_cancer_type = rbind(
  t(cat_prior_or_concomitant_by_cancer_type$n_vte_prior_or_concomitant), 
  t(cat_during_study_by_cancer_type$n_vte_during_study), 
  t(cat_recurrences_by_cancer_type$n_vte_recurrence))

# Add column and row names
colnames(cat_moment_cancer_type) = t(cat_prior_or_concomitant_by_cancer_type)[1,]
rownames(cat_moment_cancer_type) = c("Prior or concomitant", "Throughout study", "Recurrences")

cat_moment_cancer_type
```

```{r}
# Make sure that number of VTE events and sum of all VTEs by moment and cancer type check out
stopifnot(n_vtes_oncoth1 == sum(cat_moment_cancer_type))
```


```{r}
print(paste0("The total number of episodes is: ", sum(cat_moment_cancer_type)))
```


```{r}
# Do Fisher's test to see if there is some category with significantly increased/decreased episodes
apply(
  cat_moment_cancer_type,
  1,
  function(x) fisher.test(matrix(x[-1], nrow = 2))$p.value)
```

There are significant differences in VTE episodes depending on cancer type when thrombosis occurs prior or concomitant to cancer diagnosis and in recurrences.

### Number of patients

```{r}
# Count number of VTE episodes prior or concomitant to cancer diagnosis by cancer type
cat_prior_or_concomitant_by_cancer_type_patients = VTE_data %>%
  group_by(primary_tumor_simplified) %>%
  filter(n_vte_before_entering_study > 0) %>%
  summarise(n_patients = n()) %>%
  arrange(primary_tumor_simplified)
  
              
# Check all patients with VTE before study begins are accounted for
# Incorrect code
# stopifnot(
#   sum(cat_prior_or_concomitant_by_cancer_type$n_vte_prior_or_concomitant) ==
#     # Sum of time until VTE that happened before month zero
#     sum(VTE_data$tu_1st_vte_all_from_cancer_diagnosis_months <= 0, na.rm = TRUE) +
#     sum(VTE_data$tu_2nd_vte_all_from_cancer_diagnosis_months <= 0, na.rm = TRUE) + 
#     sum(VTE_data$tu_3rd_vte_all_from_cancer_diagnosis_months <= 0, na.rm = TRUE) + 
#     sum(VTE_data$tu_4th_vte_all_from_cancer_diagnosis_months<= 0, na.rm = TRUE)
#   )

cat_prior_or_concomitant_by_cancer_type_patients
```


```{r}
# Count number of VTE episodes that are recurrences by cancer type
cat_recurrences_by_cancer_type_patients = VTE_data %>%
  group_by(primary_tumor_simplified) %>%
  filter(n_total_vte_recurrence > 0) %>%
  summarise(n_patients = n()) %>%
  arrange(primary_tumor_simplified)

# sum(!is.na(tu_1st_rec_vte_from_cancer_diagnosis_months), na.rm = TRUE)

# Check all VTEs before study begins are accounted for
# stopifnot(
#   sum(cat_recurrences_by_cancer_type$n_vte_recurrence) ==
#     # Count non-NA 
#     sum(!is.na(VTE_data$tu_1st_rec_vte_from_cancer_diagnosis_months), na.rm = TRUE)
#   )

# Add row with zero for NSCLC 
cat_recurrences_by_cancer_type_patients %<>%
  add_row(primary_tumor_simplified = "NSCLC", n_patients = 0) %>%
  arrange(primary_tumor_simplified)

cat_recurrences_by_cancer_type_patients
```


```{r}
# Count number of VTE episodes during study by cancer type
cat_during_study_by_cancer_type_patients = VTE_data %>%
  group_by(primary_tumor_simplified) %>%
  filter(n_total_vte_along_study > 0) %>% # & n_total_vte_recurrence == 0
  summarise(n_patients = n()) %>%
  arrange(primary_tumor_simplified)
  
  
# Check all VTEs during study begins are accounted for
# This not correct! Do it right!
# stopifnot(
#   sum(cat_during_study_by_cancer_type$n_vte_during_study) ==
#     # Sum variables of time until VTE that happened after 
#     sum(VTE_data$tu_1st_vte_all_from_cancer_diagnosis_months > 0, na.rm = TRUE) +
#     sum(VTE_data$tu_2nd_vte_all_from_cancer_diagnosis_months > 0, na.rm = TRUE) +
#     sum(VTE_data$tu_3rd_vte_all_from_cancer_diagnosis_months > 0, na.rm = TRUE) +
#     sum(VTE_data$tu_4th_vte_all_from_cancer_diagnosis_months > 0, na.rm = TRUE) +
#     sum(VTE_data$tu_5th_vte_all_from_cancer_diagnosis_months > 0, na.rm = TRUE) -
#     sum(VTE_data$n_total_vte_recurrence)
#   )

# Careful! Recurrences are included here!
# We have to substract the number of recurrences calculated prior
cat_during_study_by_cancer_type_patients$n_patients = cat_during_study_by_cancer_type_patients$n_patients - cat_recurrences_by_cancer_type_patients$n_patients

cat_during_study_by_cancer_type_patients
```


```{r}
# Put together number of episodes by moment of occurrence
cat_moment_cancer_type_patients = rbind(
  t(cat_prior_or_concomitant_by_cancer_type_patients$n_patients), 
  t(cat_during_study_by_cancer_type_patients$n_patients), 
  t(cat_recurrences_by_cancer_type_patients$n_patients))

# Add column and row names
colnames(cat_moment_cancer_type_patients) = t(cat_prior_or_concomitant_by_cancer_type)[1,]
rownames(cat_moment_cancer_type_patients) = c("Prior or concomitant", "Throughout study", "Recurrences")

cat_moment_cancer_type_patients
```


```{r}
# Make sure that number of VTE events and sum of all VTEs by moment and cancer type check out
stopifnot(n_patients_vte == sum(cat_moment_cancer_type_patients))
```

#### Symptomatology of VTE episodes according to cancer type




```{r}
# Create dataframe from contingency table with as many observations as VTE types per moment of detection 
# This will be used to inspect differences in proportions with ggstatsplot library
x <- c()
for (row in rownames(cat_moment_cancer_type)) {
  for (col in colnames(cat_moment_cancer_type)) {
    x <- rbind(x, matrix(rep(c(row, col), cat_moment_cancer_type[row, col]), ncol = 2, byrow = TRUE))
  }
}
df2 <- as.data.frame(x)
colnames(df2) <- c("vte_type", "Cancer_type")
df2
```


```{r}
# Plot proportion of cancer types according to moment of VTE detection
ggbarstats(
  df2, Cancer_type, vte_type,
  results.subtitle = TRUE,
  subtitle = paste0(
    "Fisher's exact test", ", p-value = ",
    ifelse(test$p.value < 0.001, "< 0.001", round(test$p.value, 3))
  )
)
```

Pancreatic cancer patients show by far the highest proportion of VTEs detected prior or concomitant to cancer diagnosis. This type is also responsible for the most part of VTE recurrences. However, it is colorectal cancer the one that showed the highest number of VTEs during the observation period of the study. 


```{r}
# Get information on VTE types by location and cancer type
# %<>%
# 
# yy = left_join(
#   VTE_location, 
#   VTE_symptomatology_count, 
#   by = c("id", "primary_tumor_simplified", "n_all_vte")
#   ) # %>%
  
cancer_vte_loc = VTE_location %>%
  select(c(primary_tumor_simplified, starts_with("n_total_")))

cancer_vte_loc %<>%
  group_by(primary_tumor_simplified) %>%
  summarise(across(starts_with("n_total_"), sum)) %>%
  rename(
    CRT = n_total_crt_all, 
    DVT = n_total_dvt_all, 
    SVT = n_total_svt_all, 
    PE = n_total_pte_all, 
    VT= n_total_visceral_thromb_all
    )


# Check that all VTE events have been accounted for
stopifnot(
  n_vtes_oncoth1 == cancer_vte_loc %>%
  select_if(is.numeric) %>%
  summarise(across(everything(), sum, na.rm = TRUE)) %>%
  rowSums()
  )
```


```{r}
# Reshape the data into a longer format
VTE_symptomatology_count_long <- VTE_symptomatology_count %>%
  group_by(primary_tumor_simplified) %>%
  pivot_longer(
    cols = contains("type"), 
    values_to = "Symptomatology", 
    names_to = "column_name") %>%
  filter(!is.na(Symptomatology))

# Count occurrences for each value
count_vte_loc_sympt <- VTE_symptomatology_count_long %>%
  group_by(primary_tumor_simplified, column_name, Symptomatology) %>%
  summarise(Count = n()) %>%
  mutate(VTE_loc = case_when(
    str_detect(column_name, "crt") ~ "CRT", 
    str_detect(column_name, "dvt") ~ "DVT", 
    str_detect(column_name, "pte") ~ "PE", 
    str_detect(column_name, "svt") ~ "SVT", 
    str_detect(column_name, "visceral_thromb") ~ "VT", 
    TRUE ~ column_name
  ), .after = column_name) %>%
  ungroup() %>%
  select(-c(column_name))

# Check all VTE events have been accounted for
# Note that it is unknown the symptomatology for 1 VTE
stopifnot(sum(count_vte_loc_sympt$Count) == n_vtes_oncoth1 - 1)
```


```{r}
# Simplified count of incidental/symptomatic VTEs and VTE location per cancer type
count_vte_loc_sympt_simplified <- count_vte_loc_sympt %>%
  group_by(primary_tumor_simplified, VTE_loc, Symptomatology) %>%
  summarise(number = sum(Count)) 

# Check all VTE events have been accounted for
# Note that it is unknown the symptomatology for 1 VTE
stopifnot(sum(count_vte_loc_sympt_simplified$number) == n_vtes_oncoth1 - 1)
```


```{r}
# Count number of patients with VTE, distinguishing by location, symptomatology and cancer type
patients_vte_loc_sympt_cancer_type = VTE_symptomatology_count_long %>%
  group_by(id, primary_tumor_simplified, column_name, Symptomatology) %>%
  mutate(VTE_loc = case_when(
    str_detect(column_name, "crt") ~ "CRT", 
    str_detect(column_name, "dvt") ~ "DVT", 
    str_detect(column_name, "pte") ~ "PE", 
    str_detect(column_name, "svt") ~ "SVT", 
    str_detect(column_name, "visceral_thromb") ~ "VT", 
    TRUE ~ column_name
  ), .after = column_name) %>%
  ungroup() %>%
  select(-c(starts_with("count_"), column_name))
```


```{r}
# Show counts of patients by VTE location and symptomatology across cancer types
# Careful! Patient id = 258 is does not appear here due to the symptomatology of VTE being
# unknown
# Patient id = 258 had pancreatic cancer and 
# Remember to manually account for them in paper tables
patients_vte_loc_combination_cancer = patients_vte_loc_sympt_cancer_type %>%
  group_by(id, primary_tumor_simplified) %>%
  summarise(concatenated_VTE_loc = paste(VTE_loc, collapse = "+")) %>%
  ungroup() %>%
  mutate(concatenated_VTE_loc = case_when(
    concatenated_VTE_loc == "CRT+CRT" ~ "CRT", 
    concatenated_VTE_loc == "DVT+DVT" ~ "DVT", 
    concatenated_VTE_loc == "DVT+DVT" ~ "DVT",
    concatenated_VTE_loc == "VT+VT" ~ "VT",
    concatenated_VTE_loc == "DVT+DVT+VT" ~ "DVT+VT",
    concatenated_VTE_loc == "PE+DVT+PE+VT" ~ "DVT+PE+VT",
    TRUE ~ concatenated_VTE_loc
  )) %>%
  count(primary_tumor_simplified, concatenated_VTE_loc)

# print(sum(patients_vte_loc_combination_cancer$n))

stopifnot(sum(patients_vte_loc_combination_cancer$n) + 1 == 88)
```




```{r}
aa
```


```{r}
# cancer_type_vte_type = data.frame(
#   "DVT" = c(8, 4, 4, 11),
#   "PE" = c(12, 7, 9, 10),
#   "CRT" = c(6, 0, 1, 1),
#   "SVT" = c(0, 1, 0, 0),
#   "VT" = c(9, 1, 3, 25),
#   "Other" = c(0, 3, 0, 7),
#   row.names = c("Colorectal",
#                 "NSCLC",
#                 "Oesophago-gastric",
#                 "Pancreatic")
# )
# 
# cancer_type_vte_type
```


```{r}
# create dataframe from contingency table
x <- c()
for (row in rownames(cancer_type_vte_type)) {
  for (col in colnames(cancer_type_vte_type)) {
    x <- rbind(x, matrix(rep(c(row, col), cancer_type_vte_type[row, col]), ncol = 2, byrow = TRUE))
  }
}
df_cancer_type_vte_type <- as.data.frame(x)
colnames(df_cancer_type_vte_type) <- c("cancer_type", "vte_type")
df_cancer_type_vte_type
```


```{r}
ggbarstats(
  df_cancer_type_vte_type, cancer_type, vte_type
)
```


```{r}
#test <- fisher.test(table(df_cancer_type_vte_type))

ggbarstats(
  df_cancer_type_vte_type, vte_type, cancer_type
)
```










