---
title: 'Survival analysis of ONCOTHROMB12-01 cohort'
author: "Bárbara Lobato Delgado"
output:
  html_document:
    df_print: paged
  pdf_document: default
---

```{r}
# Deactivate scientific notation
options(scipen = 999, max.print = 10000)
```


```{r}
# knitr::knit_hooks$set(output = function(x, options) {
#    max.lines <- options$max.lines
#    x <- truncate_to_lines(x, max.lines)
# 
#    hook_output_default(x, options)
# })
```


```{r}
# Load libraries
pacman::p_load(
  kableExtra,
  an9elproject, 
  tidyverse,
  lubridate, 
  magrittr,
  survival,
  survminer, 
  install = FALSE, update = FALSE
  )
```


```{r}
###########################################
## Customised ggplot for survfit objects ##
###########################################

survplot_custom = function(survfit_obj, title, n_no_vte, n_vte, logrank_pval, legend_labels, line_colours) {
  
  if (logrank_pval < 0.0001) {
    logrank_pval = as.character("< 0.001")
  }
  
  median_survival_days = surv_median(survfit_obj)[2, 2]
  
  p = autoplot(
  survfit_obj, 
  censor = FALSE, 
  conf.int = FALSE) +
  theme_bw() +
  labs(
    x = "Time (months)", 
    y = "OS probability", 
    title = title) +
  theme(
    plot.title = element_text(hjust = 0.5), 
    legend.title = element_blank()) +
  scale_x_continuous(
    limits = c(0, 605), 
    breaks = seq(from = 0, to = 600, by = 60), 
    labels = seq(from = 0, to = 20, by = 2)) +
  scale_color_manual(
    labels = legend_labels, 
    values = line_colours) + 
  # Log-rank p-value
  annotate(
    geom = "text", 
    x = 50,
    y = 0.24,
    label = "Log-rank",
    color = "black") + 
  annotate(
    geom = "text",
    x = 50, 
    y = 0.19, 
    label = logrank_pval) +
  # n stratified by CAT
  annotate(
    geom = "text", 
    x = 600,
    y = 1,
    label = paste0("n=", n_no_vte),
    color = "black") +
  annotate(
    geom = "text", 
    x = 600,
    y = 0.95,
    label = paste0("n=", n_vte),
    color = "red") + 
  # Median survival line
  geom_segment(
    aes(x = 0, y = 0.5, xend = median_survival_days, yend = 0.5), 
    linetype = "dashed", 
    size = 0.4) +
  geom_segment(
    aes(x = median_survival_days, y = 0, xend = median_survival_days, yend = 0.5), 
    linetype = "dashed", 
    size = 0.5)  
  
  return(p)
}
```


# Load data

```{r}
# Load DB
oncoth1 = get_project("oncothr1", version = "0.0.8003") # 0.0.8004
```


```{r}
# Get data slot
oncoth1_data = oncoth1$data
```


```{r}
table(oncoth1_data$patient_group)
```


```{r}
table(oncoth1_data$patient_group, oncoth1_data$vte_recurrence)
```


```{r}
# UPDATE an9leproject!!!
# Correct information for patient id = 118
oncoth1_data %<>%
  mutate(vte_recurrence = case_when(
    id == 118 ~ "No VTE", 
    TRUE ~ vte_recurrence
    ))
```


```{r}
# Group tobacco use into 'Never smoked' vs 'Former/current'
oncoth1_data %<>%
  mutate(tobacco_use_imp_grouped = case_when(
    tobacco_use_imp == "Never has smoked" ~ "Never", 
    tobacco_use_imp == "Former smoker" ~ "Former/current",
    tobacco_use_imp == "Active smoker" ~ "Former/current"), 
    .after = tobacco_use_imp)

# Create quartiles for age 
oncoth1_data %<>%
  mutate(age_quartiles = case_when(
    age_when_cancer_dx <= 60 ~ "33-60", 
    age_when_cancer_dx > 60 & age_when_cancer_dx <= 69 ~ "61-69",
    age_when_cancer_dx >= 70 ~ "70-93"
  ), .after = age_when_cancer_dx) %>%
  mutate(age_quartiles = as.factor(age_quartiles))
```


```{r}
# Correct ties between death date and DVT date
# THIS IS NOT NECESSARY IF TIES ARE BETWEEN EVENT AND TDC
#oncoth1_data %<>%
#  mutate(follow_up_length_from_cancer_diagnosis_days = case_when(
#    id == 60 | 
#      id == 246 | 
#      id == 248 | 
#      id == 286 |
#      id == 296 |
#      id == 363 |
#      id == 441 ~ follow_up_length_from_cancer_diagnosis_days + 0.5, 
#    TRUE ~ follow_up_length_from_cancer_diagnosis_days))
```


# Create dataset for survival analysis

This dataset must be in the proper format to deal with time-dependent covariates and left- and right-censoring. The function `survival::tmerge()` helps us with it. 

```{r}
# Get subset of covariates for survival analysis
covariates_df = oncoth1_data %>%
  select(id, 
         patient_group,
         age_when_cancer_dx,
         age_quartiles,
         age_when_cancer_dx_50,
         age_when_cancer_dx_65,
         gender, 
         ethnicity,
         ethnicity_grouped,
         menopausal_status_imp,
         weight, 
         height, 
         body_surface_area,
         bmi_value,
         bmi_category, 
         performance_status_category_corrected_imp,
         hormone_therapy,
         # albumin_imp, 
         # albumin_imp_category,
         # aptt_imp, 
         # aptt_imp_category,
         # aptt_ratio_imp,
         # aptt_ratio_imp_category,
         # bilirubin_imp, 
         # bilirubin_imp_category,
         # creatinine_imp,
         # creatinine_imp_category,
         # alkaline_phosphatase_imp, 
         # alkaline_phosphatase_imp_category,
         # hemoglobin, 
         # hemoglobin_category,
         # inr_imp, 
         # inr_imp_category,
         # leukocytes_imp, 
         # leukocytes_imp_category,
         # platelets, 
         # platelets_category, 
         # leukocytosis_imp,
         # thrombocytosis, 
         # low_hemoglobin, 
         # personal_background, 
         major_surgery_in_medical_history, 
         pregnancy_imp, # constant
         oral_contraceptive_tx, # constant
         diabetes_mellitus, 
         dyslipidemia, 
         tobacco_use_imp, 
         tobacco_use_imp_grouped,
         copd_imp, 
         arterial_hypertension, 
         chronic_cardiac_insufficiency, 
         renal_insufficiency, 
         khorana_risk_score_imp,
         krs_category, 
         two_groups_krs, 
         tic_onco_imp,
         previous_onco_surgery, 
         primary_tumor_simplified, 
         tumor_surgically_removed, 
         progression_according_to_clinical_stage_imp, 
         tnm_stage, 
         tnm_stage_detailed, 
         tnm_stage_grouped, 
         tnm_stage_two_groups,
         t_stage_imp, 
         t_stage_grouped,
         n_stage_imp, 
         n_stage_grouped,
         m_stage, 
         m_stage_grouped,
         histology_type_imp, 
         mucinous_histology_imp, 
         grade_histological_differentiation_imp, 
         n_metastases,
         n_metastases_grouped,
         metastasis_adrenal_glands, 
         metastasis_bones,                             
         metastasis_brain,                               
         metastasis_liver,                               
         metastasis_lung,                                
         metastasis_lymph_nodes,                         
         metastasis_peritoneum,                          
         metastasis_pleura,                              
         metastasis_skin,                                
         metastasis_soft_tissues,                        
         other_metastases,
         catheter_device_imp,
         venous_insufficiency_imp,
         n_comorbidities,
         family_background_vte,
         family_background_vte_n,
         family_background_vte_all_relatives,
         family_background_VTE_father,    
         family_background_VTE_mother,  
         family_background_VTE_siblings, 
         family_background_VTE_offspring, 
         family_background_VTE_other_relatives, 
         previous_vte, 
         previous_ate,
         n_previous_ate,
         previous_ate_type,
         cva_ate,
         myocardial_infarction_ate, 
         pad_ate, 
         tia_ate, 
         other_thromb_ate, 
         anticoag_tx_vte,                                
         anticoag_tx_vte_drugs,                          
         anticoag_tx_cardiopathy,
         anticoag_tx_other_causes, 
         anticoag_tx_reason, 
         anticoag_tx_apixaban,                           
         anticoag_tx_dabigatran,                         
         anticoag_tx_lmwh,                               
         anticoag_tx_lmwh_dosage,                        
         anticoag_tx_vit_k_antag,                        
         anticoag_tx_other_drugs,                        
         anticoag_tx_currently,
         n_antiaggreg_tx, 
         antiaggreg_tx_reason,
         antiaggreg_tx_cardiopathy,                      
         antiaggreg_tx_cva,                              
         antiaggreg_tx_other_causes,
         antiaggreg_tx_currently, 
         tu_1st_blood_transfusion_from_cancer_diagnosis_days, 
         tu_2nd_blood_transfusion_from_cancer_diagnosis_days, 
         tu_3rd_blood_transfusion_from_cancer_diagnosis_days, 
         tu_1st_major_trauma_from_cancer_diagnosis_days, 
         tu_1st_immobilisation_from_cancer_diagnosis_days, 
         tu_2nd_immobilisation_from_cancer_diagnosis_days, 
         tu_3rd_immobilisation_from_cancer_diagnosis_days, 
         tu_1st_recent_major_surgery_from_cancer_diagnosis_days, 
         tu_2nd_recent_major_surgery_from_cancer_diagnosis_days, 
         tu_3rd_recent_major_surgery_from_cancer_diagnosis_days, 
         tu_1st_thromboprophylaxis_hospital_from_cancer_diagnosis_days, 
         tu_1st_erythropoietin_tx_from_cancer_diagnosis_days, 
         tu_2nd_erythropoietin_tx_from_cancer_diagnosis_days, 
         tu_3rd_erythropoietin_tx_from_cancer_diagnosis_days, 
         vte_recurrence)
```


```{r}
# Get subset of variables containing information about patient status and time until events
survival_data = oncoth1_data %>%
  select(id, 
         patient_group, 
         date_cancer_dx, 
         date_study_begins, 
         date_study_ends,
         date_death,
         patient_status_at_end_study, 
         censored_patient, 
         follow_up_length_from_cancer_diagnosis_days, 
         follow_up_length_from_cancer_diagnosis_months,
         tu_1st_vte_all_from_cancer_diagnosis_days_corrected,
         tu_1st_vte_all_from_cancer_diagnosis_months_corrected, 
         tu_2nd_vte_all_from_cancer_diagnosis_days_corrected, 
         tu_2nd_vte_all_from_cancer_diagnosis_months_corrected, 
         tu_3rd_vte_all_from_cancer_diagnosis_days_corrected, 
         tu_3rd_vte_all_from_cancer_diagnosis_months_corrected, 
         tu_4th_vte_all_from_cancer_diagnosis_days_corrected, 
         tu_4th_vte_all_from_cancer_diagnosis_months_corrected
         )
```


```{r}
# Create object tmerge for managing time-dependant covariates 
# Add time until death/censoring and time until VTE
# For interval censored data, the status indicator is 
# 0=right censored
# 1=event at time 
# 2=left censored
# 3=interval censored
survival_vte_tmerge = tmerge(
  data1 = covariates_df, 
  data2 = survival_data, 
  id = id, 
  tstop = follow_up_length_from_cancer_diagnosis_days, 
  # event() is the survival analysis event of interest (in our case, death)
  death = event(follow_up_length_from_cancer_diagnosis_days, censored_patient), 
  # tdc() are for time-dependent covariates
  vte_event = tdc(tu_1st_vte_all_from_cancer_diagnosis_days_corrected),
  vte_event = tdc(tu_2nd_vte_all_from_cancer_diagnosis_days_corrected),
  vte_event = tdc(tu_3rd_vte_all_from_cancer_diagnosis_days_corrected),
  vte_event = tdc(tu_4th_vte_all_from_cancer_diagnosis_days_corrected)
  )
```


```{r}
# Add other time-dependent covariates (i.e., blood transfusion, immobilisation, etc.)
survival_vte_tmerge = tmerge(
  survival_vte_tmerge, 
  survival_vte_tmerge, 
  id = id, 
  blood_transfusion = tdc(tu_1st_blood_transfusion_from_cancer_diagnosis_days),
  blood_transfusion = tdc(tu_2nd_blood_transfusion_from_cancer_diagnosis_days),
  blood_transfusion = tdc(tu_3rd_blood_transfusion_from_cancer_diagnosis_days),
  immobilisation = tdc(tu_1st_immobilisation_from_cancer_diagnosis_days), 
  immobilisation = tdc(tu_2nd_immobilisation_from_cancer_diagnosis_days), 
  immobilisation = tdc(tu_3rd_immobilisation_from_cancer_diagnosis_days), 
  recent_major_surgery = tdc(tu_1st_recent_major_surgery_from_cancer_diagnosis_days),
  recent_major_surgery = tdc(tu_2nd_recent_major_surgery_from_cancer_diagnosis_days),
  recent_major_surgery = tdc(tu_3rd_recent_major_surgery_from_cancer_diagnosis_days),
  thromboprophylaxis_hospital = tdc(tu_1st_thromboprophylaxis_hospital_from_cancer_diagnosis_days), 
  erythropoietin_tx = tdc(tu_1st_erythropoietin_tx_from_cancer_diagnosis_days),
  erythropoietin_tx = tdc(tu_2nd_erythropoietin_tx_from_cancer_diagnosis_days),
  erythropoietin_tx = tdc(tu_3rd_erythropoietin_tx_from_cancer_diagnosis_days)
  )
```


# Survival curves 

We will fit several survival curves using the Kaplan-Meier estimator. 

```{r}
# Fit Kaplan-Meier estimator 
km_estimator_fit = survfit(
  Surv(time = tstart, time2 = tstop, event = death) ~ 
    vte_event + 
    cluster(id), # ensures that the survival estimates account for the correlation within individuals, providing more robust results in the presence of repeated measurements
  data = survival_vte_tmerge)

km_estimator_fit
```

This output contains several items:

- records: The total number of records in each group (vte_event=0 and vte_event=1).
- n.max: The maximum number of subjects at risk at any time during the observation period.
- n.start: The number of subjects at the start of the observation period.
- events: The number of events (deaths) observed.
- median: The median survival time (days).
- 0.95LCL and 0.95UCL: The lower and upper 95% confidence limits for the median survival time (days).


```{r}
# Get information for all time points from Kaplan-Meier estimation
summary_fit = summary(km_estimator_fit)
formatted_summary = data.frame(
  time = summary_fit$time,
  n_risk = summary_fit$n.risk,
  n_event = summary_fit$n.event,
  survival = summary_fit$surv,
  std_error = summary_fit$std.err,
  lower_95CI = summary_fit$lower,
  upper_95CI = summary_fit$upper
)

# Show table with summary
formatted_summary %>%
  kbl(caption = "Survival Analysis Summary") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"),
                full_width = FALSE) %>%
  column_spec(ncol(formatted_summary)) 
```


```{r}
# Fit the Cox Proportional-Hazards Model
cox_model <- coxph(Surv(time = tstart, 
  time2 = tstop, 
  event = death) ~ vte_event + cluster(id), data = survival_vte_tmerge)

# Generate Survival Curves
surv_fit <- survfit(cox_model)

# Create a summary dataframe for the survival curves
surv_summary <- summary(surv_fit)

# Plot Survival Curves
plot(surv_fit, xlab = "Time", ylab = "Survival Probability", main = "Survival Curves")

```




```{r}
# We perform the log-rank test separately, because we need to manually insert the
# p-value into the KM plot
cph_fit = coxph(Surv(
  time = tstart, 
  time2 = tstop, 
  event = death) ~ 
    vte_event + 
    #strata(tnm_stage_grouped) + #test this primary_tumor_simplified
    cluster(id),
  data = survival_vte_tmerge)

cph_tests_summary = summary(cph_fit)

cph_tests_summary$robscore[["pvalue"]]

```


```{r}
# jpeg("../graphs/time_dependent_covariate_analysis/vte_td_surv_analysis_KM_curve.jpeg", width = 6.5, height = 4, units = 'in', res = 700)
td_survival_km_plot = 
  ggsurvplot(fit = km_estimator_fit,
           data = survival_vte_tmerge,
           xscale = 30,
           break.time.by = 60,
           censor = FALSE,
           #pval = TRUE, 
           #pval.method = TRUE,
           #pval.size = 4, 
           conf.int = TRUE,
           conf.int.alpha = 0.15,
           xlim = c(0, 540),
           surv.plot.height = 1.25, 
           tables.height = 0.25,
           legend.title = " ",
           legend.labs = c("No VTE", "VTE"),
           risk.table = "abs_pct", 
           risk.table.height = 0.3, 
           fontsize = 3,
           risk.table.col = "strata", 
           risk.table.y.text.col = TRUE,
           #linetype = "strata", 
           surv.median.line = "hv",
           xlab = "Time to death (months)",
           ylab = "Overall Survival Probability",
           ggtheme = theme_bw(),
           palette = c("black", "red")
           )

# Add p-value
td_survival_km_plot$plot = td_survival_km_plot$plot + 
  annotate(
    geom = "text", 
    x = 26,
    y = 0.3,
    label = "Log-rank",
    color = "black") + 
  annotate(
    geom = "text", 
    x = 28,
    y = 0.2,
    label = as.character(format(
      cph_tests_summary$robscore[["pvalue"]], 
      scientific = TRUE, 
      digits = 1)),
    color = "black")

# Customize risk table title 
td_survival_km_plot$table = 
  td_survival_km_plot$table +
  theme(plot.title = element_text(size = 10))

td_survival_km_plot

# dev.off()
```
```{r}
aa
```


### Stratified by TNM staging

```{r}
# Fit survival analysis model
stratified_stage_km_estimator_fit = survfit(
  Surv(
    time = tstart, 
    time2 = tstop, 
    event = death) ~ 
    vte_event + 
    tnm_stage_two_groups + #test this primary_tumor_simplified
    cluster(id), 
  data = survival_vte_tmerge)

stratified_stage_km_estimator_fit
```


```{r}
# We perform the log-rank test separately, because we need to manually insert the
# p-value into the KM plot
stratified_stage_cph_fit = coxph(Surv(
  time = tstart, 
  time2 = tstop, 
  event = death) ~ 
    vte_event + 
    tnm_stage_two_groups + #test this primary_tumor_simplified
    cluster(id),
  data = survival_vte_tmerge)

stratified_stage_cph_tests_summary = summary(stratified_stage_cph_fit)

stratified_stage_cph_tests_summary$robscore[["pvalue"]]

```


```{r}
#jpeg("./graphs/paper/KM_cat_and_stage.jpeg", width = 8, height = 4, units = 'in', res = 700)

# Custom KM 

# Count n patients in each subgroup
n_stages_i_ii_iii_no_vte = sum(
  oncoth1_data$tnm_stage_two_groups == "I-II-III" & 
    oncoth1_data$patient_group == "Control")

n_stages_i_ii_iii_vte = sum(
  oncoth1_data$tnm_stage_two_groups == "I-II-III" & 
    oncoth1_data$patient_group == "Case")

n_stages_iv_no_vte = sum(
  oncoth1_data$tnm_stage_two_groups == "IV" & 
    oncoth1_data$patient_group == "Control")

n_stages_iv_vte = sum(
  oncoth1_data$tnm_stage_two_groups == "IV" & 
    oncoth1_data$patient_group == "Case")

legend_labels_and_n = c(
  paste0("Stages I-II-III w/o CAT (n=", n_stages_i_ii_iii_no_vte, ")"), 
  paste0("Stage IV w/o CAT (n=", n_stages_iv_no_vte, ")"),
  paste0("Stages I-II-III with CAT (n=", n_stages_i_ii_iii_vte, ")"), 
  paste0("Stage IV with CAT (n=", n_stages_iv_vte, ")"))

#median_survival_days = surv_median(stratified_stage_km_estimator_fit)
  
# survplot_cat_stage = ggfortify::autoplot(
#   stratified_stage_km_estimator_fit, 
#   censor = FALSE, 
#   conf.int = FALSE, 
#   surv.size = 0.5, 
#   surv.linetype = "strata") +
#   theme_bw() +
#   labs(
#     x = "Time (months)", 
#     y = "OS probability", 
#     title = "") +
#   theme(
#     plot.title = element_text(hjust = 0.5), 
#     legend.title = element_text(hjust = 0.5, face = "bold"),
#     legend.text = element_text(size = 11),
#     legend.position = "right") +
#   scale_x_continuous(
#     limits = c(0, 605), 
#     breaks = seq(from = 0, to = 600, by = 60), 
#     labels = seq(from = 0, to = 20, by = 2)) +
#   scale_colour_manual(
#     name = "CAT and Stage",
#     labels = legend_labels_and_n,
#     values = c("black", "black", "red", "red")) +   
#   scale_linetype_manual(
#     name = "CAT and Stage",
#     labels = legend_labels_and_n, 
#     values = c("longdash", "dotted", "longdash", "dotted"))

# Add log-rank p-value
# survplot_cat_stage = survplot_cat_stage + 
#   # Log-rank p-value
#   annotate(
#     geom = "text", 
#     size = 4,
#     x = 50,
#     y = 0.2,
#     label = "Log-rank",
#     color = "black") + 
#   annotate(
#     geom = "text",
#     size = 4,
#     x = 50, 
#     y = 0.15, 
#     label = "<0.0001") 

# Add median survival line
# survplot_cat_stage = survplot_cat_stage +
#   # Horizontal line
#   geom_segment(
#     aes(x = 0, y = 0.5, xend = 478, yend = 0.5), 
#     linetype = "dashed", 
#     size = 0.4) +
#   # Vertical lines
#   geom_segment(
#     aes(x = 478, y = 0, xend = 478, yend = 0.5), 
#     linetype = "dashed", 
#     size = 0.4) +
#   geom_segment(
#     aes(x = 231, y = 0, xend = 231, yend = 0.5), 
#     linetype = "dashed", 
#     size = 0.4)

# survplot_cat_stage

#dev.off()
```


```{r}
summary(stratified_stage_km_estimator_fit, times = 18*30.25)
```


Stage IV heavily impacts the decrease in survival, more than having a VTE. Nevertheless, within the subgroups defined by these stages, having a VTE also significantly decreases survival.  


### Stratified by primary tumor

```{r}
# Fit survival analysis model
stratified_primary_tumor_km_estimator_fit = survfit(
  Surv(
    time = tstart, 
    time2 = tstop, 
    event = death) ~ 
    vte_event + 
    primary_tumor_simplified +
    cluster(id), 
  data = survival_vte_tmerge)

stratified_primary_tumor_km_estimator_fit
```


```{r}
# We perform the log-rank test separately, because we need to manually insert the
# p-value into the KM plot
stratified_primary_tumor_cph_fit = coxph(Surv(
  time = tstart, 
  time2 = tstop, 
  event = death) ~ 
    vte_event + 
    age_when_cancer_dx +
    gender + 
    tnm_stage_grouped +
    primary_tumor_simplified + #test this primary_tumor_simplified
    cluster(id),
  id = id, 
  data = survival_vte_tmerge)

stratified_primary_tumor_cph_tests_summary = summary(stratified_primary_tumor_cph_fit)

stratified_primary_tumor_cph_tests_summary$robscore[["pvalue"]]
```


```{r}
#jpeg("./graphs/time_dependent_covariate_analysis/stratified_primary_tumor_vte_td_surv_analysis_KM_curve.jpeg", width = 6.5, height = 4, units = 'in', res = 700)
stratified_primary_tumor_td_survival_km_plot = 
  ggsurvplot(
    fit = stratified_primary_tumor_km_estimator_fit,
    data = survival_vte_tmerge,
    xscale = 30,
    break.time.by = 60,
    censor = FALSE,
    conf.int = FALSE,
    #conf.int.alpha = 0.15,
    xlim = c(0, 540),
    #surv.plot.height = 1.25, 
    #tables.height = 0.25,
    legend.title = " ",
    legend.labs = c("No VTE - Colorectal",
                    "No VTE - NSCLC",
                    "No VTE - Oesophago-gastric",
                    "No VTE - Pancreatic",
                    "VTE - Colorectal", 
                    "VTE - NSCLC", 
                    "VTE - Oesophago-gastric",
                    "VTE - Pancreatic"),
    linetype = "vte_event", 
    surv.median.line = "hv",
    xlab = "Time to death (months)",
    ylab = "Overall Survival Probability",
    ggtheme = theme_bw(), 
    palette = c("blue", "magenta", "orange", "steelblue", "blue", "magenta", "orange", "steelblue")
    )

# Add p-value
stratified_primary_tumor_td_survival_km_plot$plot = 
  stratified_primary_tumor_td_survival_km_plot$plot + 
  annotate(
    geom = "text", 
    x = 26,
    y = 0.3,
    label = "Log-rank",
    color = "black") + 
  annotate(
    geom = "text", 
    x = 28,
    y = 0.2,
    label = as.character(format(
      stratified_primary_tumor_cph_tests_summary$robscore[["pvalue"]], 
      scientific = TRUE, 
      digits = 1)),
    color = "black")

# Customize risk table title 
#stratified_stage_td_survival_km_plot$table = 
#  stratified_stage_td_survival_km_plot$table +
#  theme(plot.title = element_text(size = 10))

stratified_primary_tumor_td_survival_km_plot

#dev.off()
```

```{r}
#jpeg("./graphs/paper/KM_cat_and_primary_tumor.jpeg", width = 8, height = 4, units = 'in', res = 700)

# Custom KM 

# Count n patients in each subgroup
n_colorectal_no_vte = sum(
  oncoth1_data$primary_tumor_simplified == "Colorectal" & 
    oncoth1_data$patient_group == "Control")

n_colorectal_vte = sum(
  oncoth1_data$primary_tumor_simplified == "Colorectal" & 
    oncoth1_data$patient_group == "Case")

n_nsclc_no_vte = sum(
  oncoth1_data$primary_tumor_simplified == "NSCLC" & 
    oncoth1_data$patient_group == "Control")

n_nsclc_vte = sum(
  oncoth1_data$primary_tumor_simplified == "NSCLC" & 
    oncoth1_data$patient_group == "Case")

n_oesophagogastric_no_vte = sum(
  oncoth1_data$primary_tumor_simplified == "Oesophago-gastric" & 
    oncoth1_data$patient_group == "Control")

n_oesophagogastric_vte = sum(
  oncoth1_data$primary_tumor_simplified == "Oesophago-gastric" & 
    oncoth1_data$patient_group == "Case")

n_pancreatic_no_vte = sum(
  oncoth1_data$primary_tumor_simplified == "Pancreatic" & 
    oncoth1_data$patient_group == "Control")

n_pancreatic_vte = sum(
  oncoth1_data$primary_tumor_simplified == "Pancreatic" & 
    oncoth1_data$patient_group == "Case")


legend_labels_and_n = c(
  paste0("Colorectal w/o CAT (n=", n_colorectal_no_vte, ")"), 
  paste0("Colorectal with CAT (n=", n_colorectal_vte, ")"),
  paste0("NSCLC w/o CAT (n=", n_nsclc_no_vte, ")"), 
  paste0("NSCLC with CAT (n=", n_nsclc_vte, ")"),
  paste0("Oesophago-gastric w/o CAT (n=", n_oesophagogastric_no_vte, ")"), 
  paste0("Oesophago-gastric with CAT (n=", n_oesophagogastric_vte, ")"),
  paste0("Pancreatic w/o CAT (n=", n_pancreatic_no_vte, ")"), 
  paste0("Pancreatic with CAT (n=", n_pancreatic_vte, ")")
  )

#median_survival_days = surv_median(stratified_stage_km_estimator_fit)
  
# survplot_cat_cancer_type = autoplot(
#   stratified_primary_tumor_km_estimator_fit, 
#   censor = FALSE, 
#   conf.int = FALSE, 
#   surv.size = 0.5, 
#   surv.linetype = "strata") +
#   theme_bw() +
#   labs(
#     x = "Time (months)", 
#     y = "OS probability", 
#     title = "") +
#   theme(
#     plot.title = element_text(hjust = 0.5), 
#     legend.title = element_text(hjust = 0.5, face = "bold"),
#     legend.text = element_text(size = 11),
#     legend.position = "right") +
#   scale_x_continuous(
#     limits = c(0, 605), 
#     breaks = seq(from = 0, to = 600, by = 60), 
#     labels = seq(from = 0, to = 20, by = 2)) +
#   scale_colour_manual(
#     name = "CAT and Cancer type",
  #   labels = legend_labels_and_n,
  #   values = c("#F8766D", "#F8766D", 
  #              "#00BA42", "#00BA42",
  #              "#00BBDD", "#00BBDD", 
  #              "#FF62BC", "#FF62BC")) +   
  # scale_linetype_manual(
  #   name = "CAT and Cancer type",
  #   labels = legend_labels_and_n, 
  #   values = c("solid", "longdash", 
  #              "solid", "longdash",
  #              "solid", "longdash", 
  #              "solid", "longdash"))

# Add log-rank p-value
# survplot_cat_cancer_type = survplot_cat_cancer_type + 
#   # Log-rank p-value
#   annotate(
#     geom = "text", 
#     size = 4,
#     x = 50,
#     y = 0.2,
#     label = "Log-rank",
#     color = "black") + 
#   annotate(
#     geom = "text",
#     size = 4,
#     x = 50, 
#     y = 0.15, 
#     label = "<0.0001") 

# Add median survival line
# survplot_cat_cancer_type = survplot_cat_cancer_type +
#   # Horizontal line
#   geom_segment(
#     aes(x = 0, y = 0.5, xend = 478, yend = 0.5), 
#     linetype = "dashed", 
#     size = 0.4) +
#   # Vertical lines
#   geom_segment(
#     aes(x = 478, y = 0, xend = 478, yend = 0.5), 
#     linetype = "dashed", 
#     size = 0.4) +
#   geom_segment(
#     aes(x = 231, y = 0, xend = 231, yend = 0.5), 
#     linetype = "dashed", 
#     size = 0.4)

# survplot_cat_cancer_type

#dev.off()
```


This plot is messy and I intend to transform it into something easier to understand. But we can see that for all 4 types of tumor, median survival is lower when patients have one or more VTEs than when they have none.  

Among the subgroup of patients that suffer VTE, pancreatic patients are the ones that die sooner (median survival time of almost 6 months). It is followed by NSCLC patients (median survival of 10.5 months, approximately); colorectal (median survival of 15.5 months approximately) and oesophago-gastric patients (median survival time of almost 19 months).  

However, in the non-VTE patient's subgroup, pancreatic, NSCLC and oesophago-gastric have similar median survival times; whilst colorectal cancer patients have the best prognostic among all of them. This could also be related to the fact that the majority of colorectal cancer patients are in stage III (45.7%) and a lesser percentage (36.4%) are in stage IV. In the previous section, we have seen that stage IV cancer patients, regardless of the type of tumor, are the ones with shorter survival times.    

The p-value shows that having a VTE is significantly associated with a difference in survival between patients subgroups, regardless of the type of cancer.  


## KM curves of other time-dependent covariates

### Blood transfusions

```{r}
# Fit survival analysis model
blood_transf_km_estimator_fit = survfit(
  Surv(
    time = tstart, 
    time2 = tstop, 
    event = death) ~ 
    blood_transfusion + 
    #strata(primary_tumor_simplified, tnm_stage_grouped) +
    cluster(id), 
  id = id,
  data = survival_vte_tmerge)

blood_transf_km_estimator_fit
```


```{r}
# We perform the log-rank test separately, because we need to manually insert the
# p-value into the KM plot
blood_transf_cph_fit = coxph(Surv(
  time = tstart, 
  time2 = tstop, 
  event = death) ~ 
    blood_transfusion + 
    #strata(primary_tumor_simplified) + #test this primary_tumor_simplified
    cluster(id),
  id = id,
  data = survival_vte_tmerge)

blood_transf_cph_tests_summary = summary(blood_transf_cph_fit)

blood_transf_cph_tests_summary$robscore[["pvalue"]]

```


```{r}
blood_transf_td_survival_km_plot = ggsurvplot(
    fit = blood_transf_km_estimator_fit,
    data = survival_vte_tmerge,
    xscale = 30,
    break.time.by = 60,
    censor = FALSE,
    conf.int = FALSE,
    #conf.int.alpha = 0.15,
    xlim = c(0, 540),
    surv.plot.height = 1.25, 
    tables.height = 0.25,
    risk.table = "abs_pct", 
    risk.table.height = 0.3, 
    fontsize = 3,
    risk.table.col = "strata", 
    risk.table.y.text.col = TRUE,
    legend.title = " ",
    legend.labs = c("No blood transfusions", 
                    "One or more blood transfusions"),
    #linetype = "Stage IV", 
    surv.median.line = "hv",
    xlab = "Time to death (months)",
    ylab = "Overall Survival Probability",
    ggtheme = theme_bw(), 
    #palette = c("black", "black", "red", "red")
    )

# Add p-value
blood_transf_td_survival_km_plot$plot = 
  blood_transf_td_survival_km_plot$plot + 
  annotate(
    geom = "text", 
    x = 26,
    y = 0.3,
    label = "Log-rank",
    color = "black") + 
  annotate(
    geom = "text", 
    x = 28,
    y = 0.2,
    label = as.character(format(
      blood_transf_cph_tests_summary$robscore[["pvalue"]], 
      scientific = TRUE, 
      digits = 1)),
    color = "black")

blood_transf_td_survival_km_plot

```


```{r}
coxph(Surv(
  time = tstart, 
  time2 = tstop, 
  event = death) ~ 
    age_when_cancer_dx + 
    gender +
    primary_tumor_simplified + 
    tnm_stage_grouped +
    blood_transfusion + 
    #strata(primary_tumor_simplified, tnm_stage_two_groups) + 
    cluster(id),
  id = id,
  data = survival_vte_tmerge) %>% 
  gtsummary::tbl_regression(exp = TRUE)
```

```{r}
cox.zph(coxph(Surv(
  time = tstart, 
  time2 = tstop, 
  event = death) ~ 
    age_when_cancer_dx + 
    gender +
    primary_tumor_simplified + 
    tnm_stage_grouped +
    blood_transfusion + 
    #strata(primary_tumor_simplified, tnm_stage_two_groups) + 
    cluster(id),
  id = id,
  data = survival_vte_tmerge))
```



### Immobilisation

```{r}
# Fit survival analysis model
immobilisation_km_estimator_fit = survfit(
  Surv(
    time = tstart, 
    time2 = tstop, 
    event = death) ~ 
    immobilisation + 
    #strata(primary_tumor_simplified, tnm_stage_grouped) +
    cluster(id), 
  id = id,
  data = survival_vte_tmerge)

immobilisation_km_estimator_fit
```


```{r}
# We perform the log-rank test separately, because we need to manually insert the
# p-value into the KM plot
immobilisation_cph_fit = coxph(Surv(
  time = tstart, 
  time2 = tstop, 
  event = death) ~ 
    immobilisation + 
    #strata(primary_tumor_simplified) + #test this primary_tumor_simplified
    cluster(id),
  id = id,
  data = survival_vte_tmerge)

immobilisation_cph_tests_summary = summary(immobilisation_cph_fit)

immobilisation_cph_tests_summary$robscore[["pvalue"]]

```


```{r}
immobilisation_td_survival_km_plot = ggsurvplot(
    fit = immobilisation_km_estimator_fit,
    data = survival_vte_tmerge,
    xscale = 30,
    break.time.by = 60,
    censor = FALSE,
    conf.int = FALSE,
    #conf.int.alpha = 0.15,
    xlim = c(0, 540),
    surv.plot.height = 1.25, 
    tables.height = 0.25,
    risk.table = "abs_pct", 
    risk.table.height = 0.3, 
    fontsize = 3,
    risk.table.col = "strata", 
    risk.table.y.text.col = TRUE,
    legend.title = " ",
    legend.labs = c("No immobilisation", 
                    "Immobilised one or more times"),
    #linetype = "Stage IV", 
    surv.median.line = "hv",
    xlab = "Time to death (months)",
    ylab = "Overall Survival Probability",
    ggtheme = theme_bw(), 
    #palette = c("black", "black", "red", "red")
    )

# Add p-value
immobilisation_td_survival_km_plot$plot = 
  immobilisation_td_survival_km_plot$plot + 
  annotate(
    geom = "text", 
    x = 26,
    y = 0.3,
    label = "Log-rank",
    color = "black") + 
  annotate(
    geom = "text", 
    x = 28,
    y = 0.2,
    label = as.character(format(
      immobilisation_cph_tests_summary$robscore[["pvalue"]], 
      scientific = TRUE, 
      digits = 1)),
    color = "black")

immobilisation_td_survival_km_plot

```


```{r}
coxph(Surv(
  time = tstart, 
  time2 = tstop, 
  event = death) ~ 
    age_when_cancer_dx + 
    gender +
    primary_tumor_simplified +
    tnm_stage_grouped + 
    immobilisation + 
    cluster(id),
  id = id,
  data = survival_vte_tmerge) %>% 
  gtsummary::tbl_regression(exp = TRUE)
```


### Recent major surgeries

```{r}
# Fit survival analysis model
recent_major_surgery_km_estimator_fit = survfit(
  Surv(
    time = tstart, 
    time2 = tstop, 
    event = death) ~ 
    recent_major_surgery + 
    #strata(primary_tumor_simplified, tnm_stage_grouped) +
    cluster(id), 
  id = id,
  data = survival_vte_tmerge)

recent_major_surgery_km_estimator_fit
```


```{r}
# We perform the log-rank test separately, because we need to manually insert the
# p-value into the KM plot
recent_major_surgery_cph_fit = coxph(Surv(
  time = tstart, 
  time2 = tstop, 
  event = death) ~ 
    recent_major_surgery + 
    #strata(primary_tumor_simplified) + #test this primary_tumor_simplified
    cluster(id),
  id = id,
  data = survival_vte_tmerge)

recent_major_surgery_cph_tests_summary = summary(recent_major_surgery_cph_fit)

recent_major_surgery_cph_tests_summary$robscore[["pvalue"]]

recent_major_surgery_cph_fit 
```


```{r}
recent_major_surgery_td_survival_km_plot = ggsurvplot(
    fit = recent_major_surgery_km_estimator_fit,
    data = survival_vte_tmerge,
    xscale = 30,
    break.time.by = 60,
    censor = FALSE,
    conf.int = FALSE,
    #conf.int.alpha = 0.15,
    xlim = c(0, 540),
    surv.plot.height = 1.25, 
    tables.height = 0.25,
    risk.table = "abs_pct", 
    risk.table.height = 0.3, 
    fontsize = 3,
    risk.table.col = "strata", 
    risk.table.y.text.col = TRUE,
    legend.title = " ",
    legend.labs = c("No major surgery", 
                    "Major surgery"),
    #linetype = "Stage IV", 
    surv.median.line = "hv",
    xlab = "Time to death (months)",
    ylab = "Overall Survival Probability",
    ggtheme = theme_bw(), 
    #palette = c("black", "black", "red", "red")
    )

# Add p-value
recent_major_surgery_td_survival_km_plot$plot = 
  recent_major_surgery_td_survival_km_plot$plot + 
  annotate(
    geom = "text", 
    x = 26,
    y = 0.3,
    label = "Log-rank",
    color = "black") + 
  annotate(
    geom = "text", 
    x = 28,
    y = 0.2,
    label = as.character(format(
      recent_major_surgery_cph_tests_summary$robscore[["pvalue"]], 
      scientific = TRUE, 
      digits = 1)),
    color = "black")

recent_major_surgery_td_survival_km_plot

```


```{r}
coxph(Surv(
  time = tstart, 
  time2 = tstop, 
  event = death) ~ 
    age_when_cancer_dx + 
    gender +
    primary_tumor_simplified +
    tnm_stage_grouped +
    recent_major_surgery +
    cluster(id),
  id = id,
  data = survival_vte_tmerge) %>% 
  gtsummary::tbl_regression(exp = TRUE)
```


### Thromboprophylaxis received during hospitalization

```{r}
# Fit survival analysis model
thromboprophylaxis_hospital_km_estimator_fit = survfit(
  Surv(
    time = tstart, 
    time2 = tstop, 
    event = death) ~ 
    thromboprophylaxis_hospital + 
    #strata(primary_tumor_simplified, tnm_stage_grouped) +
    cluster(id), 
  id = id,
  data = survival_vte_tmerge)

thromboprophylaxis_hospital_km_estimator_fit
```


```{r}
# We perform the log-rank test separately, because we need to manually insert the
# p-value into the KM plot
thromboprophylaxis_hospital_cph_fit = coxph(Surv(
  time = tstart, 
  time2 = tstop, 
  event = death) ~ 
    thromboprophylaxis_hospital + 
    #strata(primary_tumor_simplified) + #test this primary_tumor_simplified
    cluster(id),
  id = id,
  data = survival_vte_tmerge)

thromboprophylaxis_hospital_cph_tests_summary = summary(thromboprophylaxis_hospital_cph_fit)

thromboprophylaxis_hospital_cph_tests_summary$robscore[["pvalue"]]

```


```{r}
thromboprophylaxis_hospital_td_survival_km_plot = 
  ggsurvplot(
    fit = thromboprophylaxis_hospital_km_estimator_fit,
    data = survival_vte_tmerge,
    xscale = 30,
    break.time.by = 60,
    censor = FALSE,
    conf.int = FALSE,
    #conf.int.alpha = 0.15,
    xlim = c(0, 540),
    surv.plot.height = 1.25, 
    tables.height = 0.25,
    risk.table = "abs_pct", 
    risk.table.height = 0.3, 
    fontsize = 3,
    risk.table.col = "strata", 
    risk.table.y.text.col = TRUE,
    legend.title = " ",
    legend.labs = c("No thromboprophylaxis", 
                    "Thromboprophylaxis"),
    #linetype = "Stage IV", 
    surv.median.line = "hv",
    xlab = "Time to death (months)",
    ylab = "Overall Survival Probability",
    ggtheme = theme_bw(), 
    #palette = c("black", "black", "red", "red")
    )

# Add p-value
thromboprophylaxis_hospital_td_survival_km_plot$plot = 
  thromboprophylaxis_hospital_td_survival_km_plot$plot + 
  annotate(
    geom = "text", 
    x = 26,
    y = 0.3,
    label = "Log-rank",
    color = "black") + 
  annotate(
    geom = "text", 
    x = 28,
    y = 0.2,
    label = as.character(format(
      thromboprophylaxis_hospital_cph_tests_summary$robscore[["pvalue"]], 
      scientific = TRUE, 
      digits = 1)),
    color = "black")

thromboprophylaxis_hospital_td_survival_km_plot

```


```{r}
coxph(Surv(
  time = tstart, 
  time2 = tstop, 
  event = death) ~ 
    age_when_cancer_dx + 
    gender +
    primary_tumor_simplified +
    tnm_stage_grouped +
    thromboprophylaxis_hospital + 
    #strata(primary_tumor_simplified, tnm_stage_grouped) + 
    cluster(id),
  id = id,
  data = survival_vte_tmerge)  %>% 
  gtsummary::tbl_regression(exp = TRUE)
```


### Erythropoietin treatment

```{r}
# Fit survival analysis model
erythropoietin_tx_km_estimator_fit = survfit(
  Surv(
    time = tstart, 
    time2 = tstop, 
    event = death) ~ 
    erythropoietin_tx + 
    #strata(tnm_stage_two_groups) +
    cluster(id), 
  id = id,
  data = survival_vte_tmerge)

erythropoietin_tx_km_estimator_fit
```


```{r}
# We perform the log-rank test separately, because we need to manually insert the
# p-value into the KM plot
erythropoietin_tx_cph_fit = coxph(Surv(
  time = tstart, 
  time2 = tstop, 
  event = death) ~ 
    erythropoietin_tx + 
    #strata(tnm_stage_two_groups) + 
    cluster(id),
  id = id,
  data = survival_vte_tmerge)

erythropoietin_tx_cph_tests_summary = summary(erythropoietin_tx_cph_fit)

erythropoietin_tx_cph_tests_summary$robscore[["pvalue"]]

```


```{r}
erythropoietin_tx_td_survival_km_plot = 
  ggsurvplot(
    fit = erythropoietin_tx_km_estimator_fit,
    data = survival_vte_tmerge,
    xscale = 30,
    break.time.by = 60,
    censor = FALSE,
    conf.int = FALSE,
    #conf.int.alpha = 0.15,
    xlim = c(0, 540),
    surv.plot.height = 1.25, 
    tables.height = 0.25,
    risk.table = "abs_pct", 
    risk.table.height = 0.3, 
    fontsize = 3,
    risk.table.col = "strata", 
    risk.table.y.text.col = TRUE,
    legend.title = " ",
    legend.labs = c("No erythropoietin", 
                    "Erythropoietin"),
    #linetype = "Stage IV", 
    surv.median.line = "hv",
    xlab = "Time to death (months)",
    ylab = "Overall Survival Probability",
    ggtheme = theme_bw(), 
    #palette = c("black", "black", "red", "red")
    )

# Add p-value
erythropoietin_tx_td_survival_km_plot$plot = 
  erythropoietin_tx_td_survival_km_plot$plot + 
  annotate(
    geom = "text", 
    x = 26,
    y = 0.3,
    label = "Log-rank",
    color = "black") + 
  annotate(
    geom = "text", 
    x = 28,
    y = 0.2,
    label = as.character(format(
      erythropoietin_tx_cph_tests_summary$robscore[["pvalue"]], 
      scientific = TRUE, 
      digits = 1)),
    color = "black")

erythropoietin_tx_td_survival_km_plot

```

```{r}
coxph(Surv(
  time = tstart, 
  time2 = tstop, 
  event = death) ~ 
    age_when_cancer_dx + 
    gender +
    primary_tumor_simplified +
    tnm_stage_grouped +
    erythropoietin_tx + 
    #strata(primary_tumor_simplified, tnm_stage_grouped) + 
    cluster(id),
  id = id,
  data = survival_vte_tmerge)  %>% 
  gtsummary::tbl_regression(exp = TRUE)
```


```{r}
coxph(Surv(
  time = tstart, 
  time2 = tstop, 
  event = death) ~ 
    age_when_cancer_dx + 
    gender +
    vte_event:erythropoietin_tx + 
    #strata(primary_tumor_simplified, tnm_stage_grouped) +
    cluster(id),
  id = id,
  data = survival_vte_tmerge)  %>% 
  gtsummary::tbl_regression(exp = TRUE)
```

#### Erythropoietin treatment stratified by stage


```{r}
# Fit survival analysis model
stratified_stage_erythropoietin_tx_km_estimator_fit = survfit(
  Surv(
    time = tstart, 
    time2 = tstop, 
    event = death) ~ 
    erythropoietin_tx + 
    strata(tnm_stage_two_groups) + 
    cluster(id), 
  id = id,
  data = survival_vte_tmerge)

stratified_stage_erythropoietin_tx_km_estimator_fit
```


```{r}
# We perform the log-rank test separately, because we need to manually insert the
# p-value into the KM plot
stratified_stage_erythropoietin_tx_cph_fit = coxph(Surv(
  time = tstart, 
  time2 = tstop, 
  event = death) ~ 
    erythropoietin_tx + 
    strata(tnm_stage_two_groups) + 
    cluster(id),
  id = id,
  data = survival_vte_tmerge)

stratified_stage_erythropoietin_tx_cph_tests_summary = summary(stratified_stage_erythropoietin_tx_cph_fit)

stratified_stage_erythropoietin_tx_cph_tests_summary$robscore[["pvalue"]]

```


```{r}
stratified_stage_erythropoietin_tx_td_survival_km_plot = 
  ggsurvplot(
    fit = stratified_stage_erythropoietin_tx_km_estimator_fit,
    data = survival_vte_tmerge,
    xscale = 30,
    break.time.by = 60,
    censor = FALSE,
    conf.int = FALSE,
    #conf.int.alpha = 0.15,
    xlim = c(0, 540),
    #surv.plot.height = 1.25, 
    #tables.height = 0.25,
    #risk.table = "abs_pct", 
    #risk.table.height = 0.3, 
    #fontsize = 3,
    #risk.table.col = "strata", 
    #risk.table.y.text.col = TRUE,
    legend.title = " ",
    legend.labs = c("No erythropoietin - Stages I-III",
                    "No erythropoietin - Stage IV",
                    "Erythropoietin - Stages I-III", 
                    "Erythropoietin - Stage IV"),
    #linetype = "vte_event", 
    surv.median.line = "hv",
    xlab = "Time to death (months)",
    ylab = "Overall Survival Probability",
    ggtheme = theme_bw(), 
    #palette = c("black", "black", "red", "red")
    )

# Add p-value
stratified_stage_erythropoietin_tx_td_survival_km_plot$plot = 
  stratified_stage_erythropoietin_tx_td_survival_km_plot$plot + 
  annotate(
    geom = "text", 
    x = 26,
    y = 0.3,
    label = "Log-rank",
    color = "black") + 
  annotate(
    geom = "text", 
    x = 28,
    y = 0.2,
    label = as.character(format(
      erythropoietin_tx_cph_tests_summary$robscore[["pvalue"]], 
      scientific = TRUE, 
      digits = 1)),
    color = "black")

stratified_stage_erythropoietin_tx_td_survival_km_plot

```

Stage IV acts as a confounder variable, since it is the one responsible for the false statistical difference in survival between patients that had erythropoietin and those who did not. Actually, the treatment with erythropoietin does not affect survival. 


```{r}
coxph(Surv(
  time = tstart, 
  time2 = tstop, 
  event = death) ~ 
    age_when_cancer_dx + 
    gender +
    erythropoietin_tx + 
    strata(tnm_stage_grouped) + 
    cluster(id),
  id = id,
  data = survival_vte_tmerge)  %>% 
  gtsummary::tbl_regression(exp = TRUE)
```


```{r}
coxph(Surv(
  time = tstart, 
  time2 = tstop, 
  event = death) ~ 
    age_when_cancer_dx + 
    gender +
    vte_event:erythropoietin_tx + 
    strata(tnm_stage_grouped) +
    cluster(id),
  id = id,
  data = survival_vte_tmerge)  %>% 
  gtsummary::tbl_regression(exp = TRUE)
```



# Multivariate CPH regression


## Exploratory CPH regression

For example, in a Cox model, you generally should limit yourself to about 1 candidate predictor per 15 events in your data set.

https://stats.stackexchange.com/questions/178944/precisely-how-does-rs-coxph-handle-repeated-measures  

One approach to dealing with a violation of the proportional hazards assumption is to stratify by that variable. Including a strata() term will result in a separate baseline hazard function being fit for each level in the stratification variable. It will be no longer possible to make direct inference on the effect associated with that variable.  


```{r}

all_covariates_td_cph_reg = coxph(
  Surv(time = tstart, time2 = tstop, event = death) ~ 
    cluster(id) + 
    vte_event + 
    primary_tumor_simplified + # No significance along with       vte_event:tnm_stage_grouped
    #vte_event:primary_tumor_simplified:tnm_stage_two_groups + # No significance
    #vte_event:age_when_cancer_dx:menopausal_status_imp + 
    age_when_cancer_dx + 
    menopausal_status_imp +
    #age_when_cancer_dx:menopausal_status_imp + # No significance
    #age_when_cancer_dx:gender + # No significance
    #gender + 
    #gender:menopausal_status_imp +
    #ethnicity_grouped + 
    #weight + 
    #height + 
    #body_surface_area + 
    #bmi_value +
    #bmi_category + 
    performance_status_category_corrected_imp + 
    #strata(albumin_imp) + 
    #aptt_imp + 
    #aptt_ratio_imp + 
    #bilirubin_imp + 
    #alkaline_phosphatase_imp + 
    #hemoglobin + 
    #inr + # causes problems
    #leukocytes_imp + 
    #platelets + 
    #leukocytosis_imp + 
    #thrombocytosis + 
    #low_hemoglobin + 
    #personal_background + 
    #major_surgery_in_medical_history + 
    #hormone_therapy + 
    #pregnancy + 
    #oral_contraceptive_tx + 
    #diabetes_mellitus + 
    #dyslipidemia + 
    #tobacco_use + 
    #copd + 
    #arterial_hypertension + 
    #chronic_cardiac_insufficiency + 
    #renal_insufficiency + 
    #khorana_risk_score_imp + 
    #two_groups_krs + 
    #tic_onco_imp + 
    #previous_onco_surgery + 
    tumor_surgically_removed + 
    #primary_tumor_simplified +
    #progression_according_to_clinical_stage + 
    #tnm_stage + 
    #tnm_stage_detailed + 
    tnm_stage_grouped + 
    #strata(tnm_stage_grouped, vte_event) + 
    #t_stage_grouped + 
    #n_stage_grouped + 
    #m_stage_grouped + 
    #histology_type_imp + 
    #mucinous_histology_imp + 
    #grade_histological_differentiation_imp + 
    #metastasis_dx + 
    #n_metastases + 
    #metastasis_adrenal_glands + 
    #metastasis_bones +
    #metastasis_brain + 
    #metastasis_liver + 
    #metastasis_lung + 
    #metastasis_lymph_nodes + 
    #metastasis_peritoneum + 
    #metastasis_pleura + 
    #metastasis_skin + 
    #metastasis_soft_tissues + 
    #other_metastases + 
    #catheter_device_imp + 
    #venous_insufficiency_imp + 
    family_background_vte + 
    #family_background_vte_n + 
    #family_background_vte_all_relatives +
    #family_background_VTE_father + 
    #family_background_VTE_mother + 
    #family_background_VTE_siblings + 
    #family_background_VTE_offspring + 
    #family_background_VTE_other_relatives +
    previous_vte, # + 
    ##n_previous_ate +
    #previous_ate_type +
    #previous_ate + 
    #cva_ate + 
    #myocardial_infarction_ate + 
    #pad_ate + 
    #tia_ate + 
    #other_thromb_ate + 
    #anticoag_tx_vte + 
    #anticoag_tx_vte_drugs + 
    #anticoag_tx_cardiopathy + 
    #anticoag_tx_other_causes + 
    #anticoag_tx_reason + 
    #anticoag_tx_apixaban + 
    #anticoag_tx_dabigatran + 
    #anticoag_tx_lmwh + 
    #anticoag_tx_lmwh_dosage + 
    #anticoag_tx_vit_k_antag + 
    #anticoag_tx_other_drugs + 
    #anticoag_tx_currently + 
    #n_antiaggreg_tx + 
    #n_antiaggreg_tx,
    #antiaggreg_tx_reason + 
    #antiaggreg_tx_cardiopathy + 
    #antiaggreg_tx_cva + 
    #antiaggreg_tx_other_causes + 
    #antiaggreg_tx_currently
    #blood_transfusion + # Correlated with leukocytosis, low red cells, low hemoglobin, etc.
    #immobilisation + # Correlated with PS
    #recent_major_surgery + 
    #thromboprophylaxis_hospital, # + 
    #erythropoietin_tx, # Correlated with low red blood cells
  #singular.ok = TRUE,
  id = id,
  data = survival_vte_tmerge)


all_covariates_td_cph_reg %>% 
  gtsummary::tbl_regression(exp = TRUE)

```


```{r}
cox.zph(all_covariates_td_cph_reg)
```


[Investigate primary_tumor:previous_onco_surgery y tnm_stage:previous_onco_surgery.]

```{r}
concordance(all_covariates_td_cph_reg)
```



```{r}
#glm(death ~ ., family=binomial(link=logit), data=survival_vte_tmerge)
```


```{r}
# Multicollinearity
rms::vif(all_covariates_td_cph_reg)
```


```{r}
cox.zph(all_covariates_td_cph_reg)
```

## Stepwise feature selection CPH

```{r}
# Create dataset
survival_vte_tmerge_stepwise = survival_vte_tmerge %>%
  filter(!is.na(n_comorbidities))

# Do one-hot encoding with primary tumour
survival_vte_tmerge_stepwise = fastDummies::dummy_cols(survival_vte_tmerge_stepwise, select_columns = "primary_tumor_simplified")

# Rename with proper format
survival_vte_tmerge_stepwise %<>%
  rename(primary_tumor_simplified_Oesophagogastric = `primary_tumor_simplified_Oesophago-gastric`)

# Delete variables with time until event 
# They contain NAs that throw an error when training model
# This did not happen before
# survival_vte_tmerge_stepwise %<>%
#   select(-starts_with("tu_"))

```

Using `stepReg` package.  

```{r}
StepReg::stepwiseCox(
  formula = Surv(time = tstart, time2 = tstop, event = death) ~ 
    vte_event + 
    age_when_cancer_dx + 
    gender + # Multicollinear with menopausal status
    #menopausal_status_imp + 
    #weight + 
    #height + 
    #body_surface_area + 
    # bmi_value +
    bmi_category + 
    # performance_status_category_corrected_imp + 
    # albumin_imp + # Correlated with hemoglobin
    #aptt_imp_category + # correlated with aptt_ratio
    # aptt_ratio_imp + # it's best to use aptt ratio instead of aptt 
    # bilirubin_imp + # correlated with INR and alkaline phosphatase
    # creatinine_imp + 
    # alkaline_phosphatase_imp + # correlated with bilirubin and INR
    # low_hemoglobin + #_category + 
    # inr_imp + #_category +
    # leukocytes_imp + # correlated with platelets
    # platelets + # correlated with leukocytes
    # leukocytosis_imp + # correlated with leukocytes
    # thrombocytosis + # correlated with platelet levels
    # low_hemoglobin + # correlated with hemoglobin
    #major_surgery_in_medical_history + 
    #hormone_therapy + 
    #pregnancy + 
    #oral_contraceptive_tx + 
    diabetes_mellitus +
    dyslipidemia +
    tobacco_use_imp +
    # tobacco_use_imp_grouped +
    copd_imp +
    arterial_hypertension +
    chronic_cardiac_insufficiency + # Very little variance
    renal_insufficiency + # Very little variance
    venous_insufficiency_imp +
    n_comorbidities +
    #khorana_risk_score_imp + 
    #two_groups_krs + 
    #tic_onco_imp + 
    previous_onco_surgery + 
    tumor_surgically_removed + 
    primary_tumor_simplified +
    #progression_according_to_clinical_stage_imp + 
    #tnm_stage + 
    #tnm_stage_detailed + 
    tnm_stage_grouped + 
    #tnm_stage_two_groups +
    #t_stage_grouped + 
    #n_stage_grouped + 
    #m_stage_grouped + 
    histology_type_imp +
    mucinous_histology_imp +
    # grade_histological_differentiation_imp + 
    # n_metastases_grouped +
    #metastasis_adrenal_glands + 
    #metastasis_bones +
    #metastasis_brain + 
    #metastasis_liver + 
    #metastasis_lung + 
    #metastasis_lymph_nodes + 
    #metastasis_peritoneum + 
    #metastasis_pleura + 
    #metastasis_skin + 
    #metastasis_soft_tissues + 
    #other_metastases + 
    catheter_device_imp + 
    family_background_vte +
    #family_background_vte_n + 
    #family_background_vte_all_relatives +
    #family_background_VTE_father + 
    #family_background_VTE_mother + 
    #family_background_VTE_siblings + 
    #family_background_VTE_offspring + 
    #family_background_VTE_other_relatives +
    previous_vte +
    # previous_ate +
    #n_previous_ate +
    #previous_ate_type +
    anticoag_tx_currently +  
    antiaggreg_tx_currently +
    # blood_transfusion + # Correlated with low blood cells, low hemoglobin, etc. CONFOUNDER
    # immobilisation + # Correlated with PS. Also, CONFOUNDER
    # recent_major_surgery + # few observations
    # thromboprophylaxis_hospital + # CONFOUNDER
    erythropoietin_tx,
  data = survival_vte_tmerge_stepwise,
  selection = "bidirection", 
  select = "AIC", # either AIC, AICc or SBC
  method = "efron"
)
```


```{r}
final_cph_model = coxph(
  Surv(time = tstart, 
       time2 = tstop, 
       event = death) ~ 
    cluster(id) + 
    tnm_stage_grouped + 
    primary_tumor_simplified +
    vte_event +
    tumor_surgically_removed +
    catheter_device_imp +
    family_background_vte +
    previous_onco_surgery +    
    n_comorbidities +
    mucinous_histology_imp +
    diabetes_mellitus +
    erythropoietin_tx,
  id = id,
  data = survival_vte_tmerge_stepwise)

cox.zph(final_cph_model)

final_cph_model %>% 
  gtsummary::tbl_regression(exp = TRUE)
```


```{r}
summary(final_cph_model)
```


```{r}
AIC(final_cph_model)
```


```{r}
#AICcmodavg::aictab(final_cph_model)
```


### Training/test set partition and CV

See if there is consistency in variable selection and predictive capability of CV partitions, as well as with test set.

```{r}
# Set seed for reproducibility and consistency in results
set.seed(2828)
```


```{r}
# Target distribution when event of interest is death
table(oncoth1_data$patient_status_at_end_study_simplified, useNA = "always")
```


```{r}
# Partition data with repeated rows per patient into training and test set

# Get unique IDs
unique_ids <- unique(survival_vte_tmerge_stepwise$id)

# Randomly assign each unique ID to either train or test
train_ids <- sample(unique_ids, size = floor(0.8 * length(unique_ids)))  # 80% for training
test_ids <- setdiff(unique_ids, train_ids)  # Remaining IDs for testing

# Partition the data into train and test sets based on the assigned IDs
stepwise_training_set <- survival_vte_tmerge_stepwise %>% filter(id %in% train_ids)
stepwise_testing_set <- survival_vte_tmerge_stepwise %>% filter(id %in% test_ids)

# Check that IDs are not in training and test sets
stopifnot(
  identical(
    sort(unique(stepwise_training_set$id)), sort(unique(stepwise_testing_set$id))) == FALSE)
```


```{r}
# Create folds for 5-fold CV

# Create a grouping variable for each ID
group_var <- as.numeric(factor(stepwise_training_set$id))

# Determine the unique IDs and shuffle them randomly to ensure a balanced distribution across the folds
# Get unique IDs
folds_unique_ids <- unique(group_var)

# Shuffle the unique IDs randomly
folds_shuffled_ids <- sample(folds_unique_ids)

# Divide the shuffled IDs into five bins
stepwise_folds <- cut(folds_shuffled_ids, breaks = 10, labels = FALSE)
```


```{r}
# Initialize a vector to store the evaluation results
stepwise_cv_vars_results <- vector("list", length = 10)
stepwise_cv_cindex_results <- vector("list", length = 10)
# stepwise_cv_vif_results <- vector("list", length = 10)

for (i in 1:10) {
  # Get the validation IDs
  validation_ids <- unique_ids[stepwise_folds == i]
  
  # Get the validation fold
  validation_fold <- stepwise_training_set[stepwise_training_set$id %in% validation_ids, ]
  
  # Get the training fold
  training_fold <- stepwise_training_set[!(stepwise_training_set$id %in% validation_ids), ]
  
  # print("Training")
  # print(head(training_fold, n = 10))
  # 
  # print("Validation")
  # print(head(validation_fold, n = 10))
  
  # Check that IDs are not in training and test sets
  stopifnot(
    identical(
      sort(unique(training_fold$id)), sort(unique(validation_fold$id))) == FALSE)
  
  # Train your model on the training fold
  stepwise_model = StepReg::stepwiseCox(
    formula = Surv(time = tstart, time2 = tstop, event = death) ~
    vte_event +
    # vte_recurrence +
    # vte_event:vte_recurrence +
    age_when_cancer_dx +
    # age_quartiles + 
    # age_when_cancer_dx_50 +
    # age_when_cancer_dx_65 +
    gender + # Collinear with menopausal status
    # menopausal_status_imp +
    # weight +
    # height +
    # body_surface_area +
    bmi_value +
    # bmi_category +
    performance_status_category_corrected_imp +
    # major_surgery_in_medical_history +
    # hormone_therapy +
    # pregnancy +
    # oral_contraceptive_tx +
    # diabetes_mellitus + # confounder
    dyslipidemia +
    tobacco_use_imp +
    # tobacco_use_imp_grouped +
    copd_imp +
    arterial_hypertension +
    # chronic_cardiac_insufficiency + # Very little variance
    # renal_insufficiency + # Very little variance
    venous_insufficiency_imp +
    # n_comorbidities +
    # khorana_risk_score_imp +
    # two_groups_krs +
    # tic_onco_imp +
    # previous_onco_surgery +
    tumor_surgically_removed + # uncomment
    primary_tumor_simplified +
    # primary_tumor_simplified_Colorectal +
    # primary_tumor_simplified_NSCLC +
    # primary_tumor_simplified_Oesophagogastric +
    # primary_tumor_simplified_Pancreatic +
    # progression_according_to_clinical_stage_imp + # too wide CI
    #tnm_stage +
    #tnm_stage_detailed +
    # tnm_stage_grouped +
    tnm_stage_two_groups +
    # t_stage_grouped +
    # n_stage_grouped +
    # m_stage_grouped +
    # histology_type_imp +
    mucinous_histology_imp +
    # grade_histological_differentiation_imp + # violates PH assumption
    # n_metastases_grouped + # confounder, possibly correlated to pTNM stage
    #metastasis_adrenal_glands +
    #metastasis_bones +
    #metastasis_brain +
    #metastasis_liver +
    #metastasis_lung +
    #metastasis_lymph_nodes +
    #metastasis_peritoneum +
    #metastasis_pleura +
    #metastasis_skin +
    #metastasis_soft_tissues +
    #other_metastases +
    # catheter_device_imp + # quite correlated with tumor type, as Cramer's V indicates
    family_background_vte +
    #family_background_vte_n +
    #family_background_vte_all_relatives +
    #family_background_VTE_father +
    #family_background_VTE_mother +
    #family_background_VTE_siblings +
    #family_background_VTE_offspring +
    #family_background_VTE_other_relatives +
    previous_vte +
    previous_ate +
    # n_previous_ate +
    # previous_ate_type +
    anticoag_tx_currently +
    antiaggreg_tx_currently, # +
    # blood_transfusion + # Correlated with low blood cells, low hemoglobin, etc. CONFOUNDER
    # immobilisation + # Correlated with PS. Also, CONFOUNDER
    # recent_major_surgery + # few observations
    # thromboprophylaxis_hospital + # CONFOUNDER
    # erythropoietin_tx,
    data = training_fold,
    include = "primary_tumor_simplified",
    selection = "bidirection",
    select = "AIC", # either AIC, AICc or SBC
    method = "efron"
    )
  
  # Save selected variables
  selected_variables = as.vector(unlist(stepwise_model$`Selected Varaibles`))
  # print(selected_variables)
  
  # Fit model with selected variables in this fold
  # Create formula
  fold_pcox_formula = as.formula(
    sprintf('%s~%s', 'Surv(time = tstart, time2 = tstop, event = death)',
            paste(selected_variables, collapse = "+ ")))
  # Fit model
  fold_cph_model = coxph(
    formula = fold_pcox_formula, 
    id = id,
    data = validation_fold
  )
  
  # VIF
  # vif_results = car::vif(fold_cph_model)
  # print(vif_results)
  
  # Make predictions on the validation fold
  # The result is the hazard of suffering the event. It's difficult to interpret
  predictions <- predict(fold_cph_model, newdata = validation_fold)
  # print(predictions)
  
  # Evaluate the predictions
  # C-index
  fold_cindex <- concordance(fold_cph_model, newdata = validation_fold)
  # print(fold_cindex$concordance)
  
  # Likelihood ratio test (LRT)
  # fold_lrt <- anova(fold_cph_model)
  # fold_lrt_p_value <- fold_lrt$`Pr(>Chisq)`[2]  # P-value for the model
  # print(fold_lrt_p_value)
  
  # Store the evaluation result
  stepwise_cv_vars_results[[i]] = selected_variables
  stepwise_cv_cindex_results[[i]] = fold_cindex$concordance
  # stepwise_cv_vif_results[[i]] = as.data.frame(vif_results)

}
```

Age by itself is not considered relevant.

Patients grouped by aged with 65 years old as threshold is not a very important variable. However, when the threshold is 50 years old, it is selected as important, with patients equal or older than 50 y.o. being protected against death. 

This strange result may be due to the tumor type:

                              Colorectal NSCLC Oesophago-gastric Pancreatic
  Younger than 50 years old         11     5                 9          8
  Older or 50 years old            151    82                60         64


```{r}
# Variable selection frequency across 10-fold CV
stepwise_var_selection_freq = as.data.frame(table(unlist(stepwise_cv_vars_results)))
stepwise_var_selection_freq %<>% arrange(desc(Freq))
knitr::kable(stepwise_var_selection_freq)
```


```{r}
# Mean C-index across 10-fold CV
mean(unlist(stepwise_cv_cindex_results))
```


```{r}
# Do correlation between CVC and tumor surgically removed with primary tumor
oncoth1_corr_vars = oncoth1_data %>% 
  select(id, primary_tumor_simplified, catheter_device, tumor_surgically_removed)

# Calculate Cramér's V
# Close to 0: Weak or no association.
# Around 0.1: Small association.
# Around 0.3: Moderate association.
# Around 0.5: Strong association.
# Close to 1: Very strong or perfect association.

# Primary tumor and CVC
contingency_table1 <- table(
  oncoth1_corr_vars$primary_tumor_simplified,
  oncoth1_corr_vars$catheter_device
  )

cramers_v1 <- round(sqrt(chisq.test(contingency_table1)$statistic / sum(contingency_table1)), digits = 4)
print(paste("Cramer's V for cancer type and CVC =", cramers_v1))


# Primary tumor and tumor surgically removed
contingency_table2 <- table(
  oncoth1_corr_vars$primary_tumor_simplified,
  oncoth1_corr_vars$tumor_surgically_removed
  )

cramers_v2 <- round(sqrt(chisq.test(contingency_table2)$statistic / sum(contingency_table2)), digits = 4)
print(paste("Cramer's V for cancer type and surgical resection of tumor =", cramers_v2))
```



```{r}
# Convert variable from factor to binary
# stepwise_training_set$mucinous_histology_imp = ifelse(stepwise_training_set$mucinous_histology_imp == "No", 0, 1)
```


```{r}
# Relevel factor
# stepwise_training_set$mucinous_histology_imp = relevel(stepwise_training_set$mucinous_histology_imp, ref = "Yes")
```


```{r}
# Final model with most frequently selected variables
stepwise_cv_cph_final_model = coxph(
  Surv(time = tstart, 
       time2 = tstop, 
       event = death) ~ 
    cluster(id) + 
    vte_event + # :vte_recurrence + # 
    performance_status_category_corrected_imp +
    primary_tumor_simplified +
    tnm_stage_two_groups +
    tumor_surgically_removed +
    venous_insufficiency_imp +
    mucinous_histology_imp,
  id = id,
  data = stepwise_training_set
  )

# Check proportional hazards assumption for Cox model
cox.zph(stepwise_cv_cph_final_model)

# Show HR and 95% CI
stepwise_cv_cph_final_model %>% 
  gtsummary::tbl_regression(exp = TRUE)
```


```{r}
# Does not work with models where there is interaction between variables
vif_results = car::vif(stepwise_cv_cph_final_model)
vif_results
```


```{r}
# C-index in testing set
concordance(stepwise_cv_cph_final_model, newdata = stepwise_testing_set)
```


```{r}
# # Predict survival probabilities for different groups
# # Does not work with models where there is interaction between variables
# stepwise_model_predicted_survival = survfit(stepwise_cv_cph_final_model)
# 
# # Plot survival curves
# # It cannot be done because I cannot apply survfit() to model with variable's interacions
# ggsurvplot(
#   stepwise_model_predicted_survival,
#   data = stepwise_training_set,
#   title = "Survival Curves with Interaction",
#   xlab = "Time",
#   ylab = "Survival Probability",
#   risk.table = TRUE,
#   risk.table.title = "Group Comparison"
#   )
```


```{r}
# https://stat.ethz.ch/R-manual/R-devel/library/survival/html/predict.coxph.html
# Choices are the linear predictor ("lp"), the risk score exp(lp) ("risk"), the expected number of events given the covariates and follow-up time ("expected"), and the terms of the linear predictor ("terms"). The survival probability for a subject is equal to exp(-expected)
# stepwise_pred = predict(
#   object = stepwise_cv_cph_final_model, 
#   newdata = stepwise_testing_set, 
#   se.fit = TRUE,
#   type = "survival"
#   )
# 
# # Get probabilities and SE
# stepwise_testing_set$prob = stepwise_pred$fit
# stepwise_testing_set$lcl = stepwise_pred$fit - 1.96 * stepwise_pred$se.fit
# stepwise_testing_set$ucl = stepwise_pred$fit + 1.96 * stepwise_pred$se.fit
```


```{r}
yhat <- predict(
  stepwise_cv_cph_final_model,
  newdata = stepwise_testing_set,
  se.fit = TRUE,
  reference = "zero"
  )

yy <- yhat$fit + outer(yhat$se.fit, c(0, -1.96, 1.96), '*')
yy
```



## Adjusted survival curves to account for confounder covariates

### VTE event

```{r}
ggadjustedcurves(
  fit = all_covariates_td_cph_reg, 
  variable = "vte_event",
  data = survival_vte_tmerge, 
  method = "conditional", 
  #xscale = 30,
  #break.time.by = 60,
  xlim = c(0, 540), 
  surv.median.line = "hv",
  ggtheme = theme_bw()
  )
```


### Primary tumor

```{r}
# Primary tumor
ggadjustedcurves(
  fit = all_covariates_td_cph_reg, 
  data = survival_vte_tmerge, 
  variable = "primary_tumor_simplified",
  #individual.curves = TRUE,
  method = "conditional",
  #xscale = 30,
  #break.time.by = 60,
  #censor = FALSE,
  #conf.int = FALSE,
  #conf.int.alpha = 0.15,
  xlim = c(0, 540),
  legend.title = "Cancer type",
  #legend.labs = c("No VTE", "VTE"),
  #linetype = "strata", 
  #surv.median.line = "hv",
  xlab = "Time (days)",
  ylab = "Overall Survival Probability",
  ggtheme = theme_bw(),
  #palette = c("black", "red")
  )

```


### TNM stage

```{r}
# Stage
ggadjustedcurves(
  fit = all_covariates_td_cph_reg, 
  data = survival_vte_tmerge, 
  variable = "tnm_stage_grouped",
  #individual.curves = TRUE,
  method = "conditional",
  #xscale = 30,
  #break.time.by = 60,
  #censor = FALSE,
  #conf.int = FALSE,
  #conf.int.alpha = 0.15,
  xlim = c(0, 540),
  legend.title = "Stage",
  #legend.labs = c("No VTE", "VTE"),
  #linetype = "strata", 
  #surv.median.line = "hv",
  xlab = "Time (days)",
  ylab = "Overall Survival Probability",
  ggtheme = theme_bw(),
  #palette = c("black", "red")
  )

```


```{r}
# Table of times from ggadjustedcurves()
# I can extract the median survival time
surv_adjustedcurves(
  fit = all_covariates_td_cph_reg, 
  data = survival_vte_tmerge, 
  variable = "tnm_stage_grouped",
  method = "conditional", 
  ggtheme = theme_bw()
)
```


# Univariate CPH regression

First, we perform univariate CPH regression on all the covariates of the data set in order to see which ones are correlated with survival.  

```{r}
# I need to define manually the covariates to do univariate CPH
covariates_names = c(
  "vte_event",
  "age_when_cancer_dx",
  "age_when_cancer_dx_50",
  "age_when_cancer_dx_65",
  "gender", 
  "ethnicity",
  "ethnicity_grouped",
  "menopausal_status_imp",
  "weight", 
  "height", 
  "body_surface_area",
  "bmi_value",
  "bmi_category", 
  "performance_status_category_corrected_imp",
  # "albumin_imp_category", 
  # "aptt_imp_category", 
  # "aptt_ratio", 
  # "bilirubin_imp",
  # "creatinine_imp",
  # "alkaline_phosphatase_imp", 
  # "hemoglobin", 
  # "inr_imp", 
  # "leukocytes_imp", 
  # "platelets", 
  # "leukocytosis_imp", 
  # "thrombocytosis", 
  # "low_hemoglobin", 
  "major_surgery_in_medical_history", 
  "hormone_therapy", 
  "pregnancy_imp", 
  #"oral_contraceptive_tx", 
  "diabetes_mellitus", 
  "dyslipidemia", 
  "tobacco_use_imp", 
  "copd", 
  "arterial_hypertension", 
  "chronic_cardiac_insufficiency", 
  "renal_insufficiency", 
  "n_comorbidities",
  "khorana_risk_score_imp",
  "krs_category", 
  "two_groups_krs", 
  "tic_onco_imp",
  "previous_onco_surgery", 
  "primary_tumor_simplified", 
  "tumor_surgically_removed", 
  "progression_according_to_clinical_stage_imp", 
  "tnm_stage", 
  "tnm_stage_detailed", 
  "tnm_stage_grouped", 
  "t_stage_imp", 
  "t_stage_grouped",
  "n_stage_imp", 
  "n_stage_grouped",
  "m_stage", 
  "m_stage_grouped",
  "histology_type_imp", 
  "mucinous_histology_imp", 
  "grade_histological_differentiation_imp", 
  "n_metastases",
  "metastasis_adrenal_glands", 
  "metastasis_bones", 
  "metastasis_brain", 
  "metastasis_liver",  
  "metastasis_lung",   
  "metastasis_lymph_nodes", 
  "metastasis_peritoneum",  
  "metastasis_pleura", 
  "metastasis_skin", 
  "metastasis_soft_tissues",  
  "other_metastases",
  "catheter_device_imp",
  "venous_insufficiency_imp",
  "family_background_vte",
  "family_background_vte_n",
  "family_background_vte_all_relatives",
  "family_background_VTE_father",
  "family_background_VTE_mother", 
  "family_background_VTE_siblings", 
  "family_background_VTE_offspring", 
  "family_background_VTE_other_relatives", 
  "previous_vte", 
  "previous_ate",
  "n_previous_ate",
  "previous_ate_type",
  "cva_ate",
  "myocardial_infarction_ate", 
  "pad_ate", 
  "tia_ate", 
  "other_thromb_ate", 
  "anticoag_tx_vte", 
  "anticoag_tx_vte_drugs", 
  "anticoag_tx_cardiopathy",
  "anticoag_tx_other_causes", 
  "anticoag_tx_reason", 
  #"anticoag_tx_apixaban",  
  #"anticoag_tx_dabigatran",     
  #"anticoag_tx_lmwh",     
  #"anticoag_tx_lmwh_dosage",  
  #"anticoag_tx_vit_k_antag",  
  #"anticoag_tx_other_drugs", 
  #"anticoag_tx_currently",
  "n_antiaggreg_tx", 
  "antiaggreg_tx_reason",
  "antiaggreg_tx_cardiopathy", 
  "antiaggreg_tx_cva", 
  "antiaggreg_tx_other_causes",
  "antiaggreg_tx_currently", 
  "blood_transfusion", 
  #"major_trauma", 
  "immobilisation",
  "recent_major_surgery",
  "thromboprophylaxis_hospital",
  "erythropoietin_tx"
  )

```


```{r}
univariate_cph_reg_results = list()

for (covariate in covariates_names) {
  i_formula <- as.formula(
        # Returns a character vector containing a formatted combination of text and
        # variable values
        sprintf("Surv(time = tstart, time2 = tstop, event = death) ~ %s ", covariate))
  
  i_cph <- coxph(i_formula, data = survival_vte_tmerge, na.action = na.omit)
  
  print(covariate)
  
  univariate_cph_reg_results[[covariate]] <- summary(i_cph)
  
  i_coxzph = cox.zph(i_cph)
  
  pval_coxzph = i_coxzph$table[1, 3]
  
  univariate_cph_reg_results[[covariate]]["pval_ph_check"] <- pval_coxzph

  print(summary(i_cph))
  cat("\n")
  
  print("P-value of proportional hazards testing for each covariate: ")
  print(covariate)
  print(pval_coxzph)
  cat("\n")
  
  if (pval_coxzph < 0.05) {
    print(ggcoxzph(cox.zph(i_cph)))
    cat("\n")
    cat("\n")
  }
}

```



# Impact of VTE by location on survival

```{r}
# Load latest version of oncoth1
oncoth1 = get_project("oncothr1", version = "0.0.8003")
```


```{r}
covariates_df = oncoth1_data %>%
  select(id, 
         patient_group,
         age_when_cancer_dx,
         age_when_cancer_dx_50, 
         age_when_cancer_dx_65,
         gender, 
         ethnicity_grouped,
         menopausal_status_imp,
         bmi_category, 
         performance_status_category_corrected_imp,
         khorana_risk_score_imp,
         krs_category, 
         two_groups_krs, 
         tic_onco_imp,
         previous_onco_surgery, 
         primary_tumor_simplified, 
         tumor_surgically_removed, 
         tnm_stage_grouped, 
         tnm_stage_two_groups,
         n_metastases,
         catheter_device_imp,
         family_background_vte,
         previous_vte, 
         previous_ate, 
         vte_recurrence
         )
```


```{r}
# Get events and time until events for VTE according to locations
all_vte_types_survival_data = oncoth1_data %>%
  select(id, 
         patient_group, 
         date_cancer_dx, 
         date_study_begins, 
         date_study_ends,
         date_death,
         patient_status_at_end_study, 
         censored_patient, 
         follow_up_length_from_cancer_diagnosis_days, 
         follow_up_length_from_cancer_diagnosis_months,
         # CRT
         starts_with("tu_1st_crt_"),
         # DVT
         starts_with("tu_1st_dvt_"),
         starts_with("tu_2nd_dvt_"),
         # PTE
         starts_with("tu_1st_pte_"),
         starts_with("tu_2nd_pte_"),
         # Visceral thrombosis
         starts_with("tu_1st_visceral_thromb_"),
         starts_with("tu_2nd_visceral_thromb_")
         # SVT
         # tu_1st_svt...
         # faltan variables
         # Others
         # These features should be erased
         # tu_1st_other_thromb_grouped_days_corrected,
         # tu_2nd_other_thromb_grouped_days, 
         # tu_1st_other_thromb_grouped_months_corrected,
         # tu_2nd_other_thromb_grouped_months, 
         # tu_1st_other_thromb_grouped_from_cancer_diagnosis_days_corrected,
         # tu_2nd_other_thromb_grouped_from_cancer_diagnosis_days,  
         # tu_1st_other_thromb_grouped_from_cancer_diagnosis_months_corrected,
         # tu_2nd_other_thromb_grouped_from_cancer_diagnosis_months
         )

```


```{r}
# Create object tmerge for time-dependant covariate survival analysis
# For interval censored data, the status indicator is 0=right censored, 1=event at time, 
# 2=left censored, 3=interval censored.
all_types_vte_tmerge = tmerge(
  data1 = covariates_df, 
  data2 = all_vte_types_survival_data, 
  id = id, 
  tstop = follow_up_length_from_cancer_diagnosis_days, 
  # event() is the survival analysis event of interest (in our case, death)
  death = event(follow_up_length_from_cancer_diagnosis_days, censored_patient), 
  # CRTs
  crt_event = tdc(tu_1st_crt_from_cancer_diagnosis_days_corrected), 
  # DVTs
  dvt_event = tdc(tu_1st_dvt_from_cancer_diagnosis_days_corrected),
  dvt_event = tdc(tu_2nd_dvt_from_cancer_diagnosis_days), 
  # PTEs
  pte_event = tdc(tu_1st_pte_from_cancer_diagnosis_days_corrected), 
  pte_event = tdc(tu_2nd_pte_from_cancer_diagnosis_days), 
  # Visceral thrombosis
  visceral_thromb_event = 
    tdc(tu_1st_visceral_thromb_from_cancer_diagnosis_days_corrected), 
  visceral_thromb_event = 
    tdc(tu_2nd_visceral_thromb_from_cancer_diagnosis_days) # ,
  # Other thromboses grouped
  # There is still information about other events
  # other_thromb_grouped_event = 
  #   tdc(tu_1st_other_thromb_grouped_from_cancer_diagnosis_days_corrected),
  # other_thromb_grouped_event =
  #   tdc(tu_2nd_other_thromb_grouped_from_cancer_diagnosis_days)
  )
```


```{r}
# Get patient's ids for each type of VTE types and combinations
# Also patients with no VTE 

# Temporal df
tmp_id_types_vte = oncoth1_data %>%
  select(id, 
         any_crt,
         any_dvt, 
         any_pte, 
         any_svt,
         any_visceral_thromb)

# Only CRT
id_only_crt = tmp_id_types_vte %>%
  filter(!any_pte == "Yes" & 
           any_crt == "Yes" &
           !any_svt == "Yes" &
           !any_visceral_thromb == "Yes" &
           !any_dvt == "Yes") %>%
  distinct(id) %>%
  unlist()

# Only DVT
id_only_dvt = tmp_id_types_vte %>%
  filter(!any_pte == "Yes" & 
           !any_crt == "Yes" &
           !any_svt == "Yes" &
           !any_visceral_thromb == "Yes" &
           any_dvt == "Yes") %>%
  distinct(id) %>%
  unlist()

# Only PTE
id_only_pte = tmp_id_types_vte %>%
  filter(!any_dvt == "Yes" &
           !any_crt == "Yes" &
           !any_svt == "Yes" &
           !any_visceral_thromb == "Yes" &
           any_pte == "Yes") %>%
  distinct(id) %>%
  unlist()

# Only VT
id_only_visceral_thromb = tmp_id_types_vte %>%
  filter(!any_pte == "Yes" & 
           !any_crt == "Yes" &
           !any_svt == "Yes" &
           !any_dvt == "Yes" & 
           any_visceral_thromb == "Yes") %>%
  distinct(id) %>%
  unlist()

# Only SVT
id_only_svt = tmp_id_types_vte %>%
  filter(!any_pte == "Yes" & 
           !any_crt == "Yes" &
           any_svt == "Yes" &
           !any_dvt == "Yes" & 
           !any_visceral_thromb == "Yes") %>%
  distinct(id) %>%
  unlist()


# DVT + PTE
id_dvt_and_pte = tmp_id_types_vte %>%
  filter(!any_visceral_thromb == "Yes" & 
           !any_crt == "Yes" &
           !any_svt == "Yes" &
           any_dvt == "Yes" & 
           any_pte == "Yes") %>%
  distinct(id) %>%
  unlist()

# DVT + visceral thrombosis
id_dvt_and_visceral_thromb = tmp_id_types_vte %>%
  filter(!any_pte == "Yes" & 
           !any_crt == "Yes" &
           !any_svt == "Yes" &
           any_dvt == "Yes" & 
           any_visceral_thromb == "Yes") %>%
  distinct(id) %>%
  unlist()


# PTE + VT
id_pte_and_visceral_thromb = tmp_id_types_vte %>%
  filter(
    !any_dvt == "Yes" & 
      !any_crt == "Yes" &
           !any_svt == "Yes" &
      any_pte == "Yes" & 
      any_visceral_thromb == "Yes") %>%
  distinct(id) %>%
  unlist()


# DVT + PTE + VT (there no patients)
id_dvt_and_pte_and_vt = tmp_id_types_vte %>%
  filter(
      any_dvt == "Yes" & 
      any_pte == "Yes" & 
      any_visceral_thromb == "Yes") %>%
  distinct(id) %>%
  unlist()


# All types
id_all_types_vte = tmp_id_types_vte %>%
  filter(
    any_dvt == "Yes" & 
      any_visceral_thromb == "Yes" & 
      any_pte == "Yes") %>%
  distinct(id) %>%
  unlist()


# No VTEs
id_no_vte = tmp_id_types_vte %>%
  filter(
    any_dvt == "No" & 
      any_visceral_thromb == "No" & 
      any_pte == "No") %>%
  distinct(id) %>%
  unlist()

```


# DVT vs no VTE

```{r}
multivariate_km_log_rank_test = function(tu1_var, tu2_var, event_var, covariates, dataset, plot_curve = TRUE) {
  
  # Define survival object
  surv_object = Surv(
    time = dataset[[tu1_var]], 
    time2 = dataset[[tu2_var]], 
    event = as.numeric(dataset[[event_var]])
    )
  
  # Build formula
  # formula = reformulate(covariate, response = "Surv(tu1_var, event_var)")
  
  # Returns a character vector containing a formatted combination of text and variable values
  formula = as.formula(sprintf("surv_object ~ %s ", paste(covariates, collapse = "+")))
  print(formula)
  
  # Create survfit object
  # It estimates the survival function (probability of survival) over time for censored data
  survfit_object = survfit(formula = formula, data = dataset)
  
  # Show summary of survival function
  cat("Median OS from KM fitted survival curve:")
  print(summary(survfit_object)$table/30.25)
  cat("\n")
  
  # Overall survival at 18 months
  cat("OS rate at 18 months:")
  cat("\n")
  print(summary(survfit_object, times = 18*30.25))
  cat("\n")
  
  # In the case of interval censored data, log-rank test cannot be computed
  # Instead, we fit a Cox model and extract the score test, which indicates 
  # whether there are significant differences in survival between the groups
  # Cox model 
  cph_model = coxph(
    formula = formula,
    # id = id,
    data = dataset
    )
  
  # Get summary of results
  cph_model_summary = summary(cph_model)
  print(cph_model_summary)
  
  # Get score test p-value
  score_test_pval = format(
    unname(cph_model_summary$sctest["pvalue"])
  )
  
  print(score_test_pval)
  
  # Do log-rank test to compare two or more survival curves
  # logrank_test = survdiff(formula = formula, data = dataset) # survdiff
  # Get log-rank test p-value
  # lrtest_pval = logrank_test$pvalue
  # pvalue = surv_pvalue(fit = survfit_object, data = dataset, method = "1")
  
  
  # Print p-value
  if (score_test_pval > 0.05) {
    pval = "p-value is NS"
    print(paste("Cox score test (approximation of log-rank test) p-value is NS"))
  } else if (score_test_pval < 0.05 & score_test_pval > 0.01) {
    pval = "p-value < 0.05"
    print(paste("Cox score test (approximation of log-rank test) p-value < 0.05"))
  } else if (score_test_pval < 0.01 & score_test_pval > 0.001) {
    pval = "p-value < 0.01"
    print(paste("Cox score test (approximation of log-rank test) p-value < 0.01"))
  } else if (score_test_pval < 0.001 & score_test_pval > 0.0001) {
    pval = "p-value < 0.001"
    print(paste("Cox score test (approximation of log-rank test) p-value < 0.001"))
  } else if (score_test_pval < 0.0001) {
    pval = "p-value < 0.0001"
    print(paste("Cox score test (approximation of log-rank test) p-value < 0.0001"))
  }
  
  # Return survfit object, needed to plot the survival curve
  if (plot_curve) {

    km_plot = ggsurvplot(
      fit = surv_fit(formula = formula, data = dataset),
      data = dataset,
      break.time.by = 60,
      censor = FALSE,
      # color = "black",
      conf.int = TRUE,
      conf.int.alpha = 0.5,
      xscale = 30,
      xlim = c(0, 540),
      surv.plot.height = 1.25,
      legend.title = " ",
      legend.labs = c("No CAT", "CAT"),
      risk.table = "abs_pct",
      risk.table.height = 0.3,
      risk.table.col = "strata", 
      risk.table.y.text.col = TRUE,
      fontsize = 2.75,
      surv.median.line = "hv",
      xlab = "Time (months)",
      ylab = "OS Probability",
      ggtheme = theme_bw(),
      palette = c("black", "red")
      )
    
    # Add p-value
    km_plot$plot = km_plot$plot + 
      annotate(
        geom = "text", 
        x = 26,
        y = 0.3,
        label = "",
        color = "black") + 
      annotate(
        geom = "text", 
        x = 50,
        y = 0.2,
        label = pval,
        color = "black")
    
    # Customize risk table title 
    km_plot$table = 
      km_plot$table +
      theme(plot.title = element_text(size = 10))
    
    return(km_plot)
  }

}
```


```{r}
print(paste("There are", length(id_only_dvt), "patients that suffered only DVT."))
```


```{r}
# Obtain median survival and 18 month overall survival rate
# Plot survival curves VTE vs no VTE
surv_curve_vte = multivariate_km_log_rank_test(
  tu1_var = "tstart", 
  tu2_var = "tstop", 
  event_var = "death", 
  covariates = c("dvt_event"), 
  dataset = all_types_vte_tmerge, 
  plot_curve = TRUE
  )
```

First, we need to exclude the patients that have more than one type of VTE, because having multiple types of VTE probably decreases survival and the effect of one type of VTE ends up being overestimated.  


```{r}
dvt_no_vte_tmerge = all_types_vte_tmerge %>%
  group_by(id) %>%
  filter(id %in% c(id_only_dvt, id_no_vte)) 
```


```{r}
dvt_no_vte_survfit = survfit(Surv(
  time = tstart, 
  time2 = tstop,
  event = death) ~ dvt_event + cluster(id), 
    id = id,
  data = dvt_no_vte_tmerge)

dvt_no_vte_survfit
```


```{r}
# Cox regression
dvt_no_vte_cph_fit = coxph(Surv(
  time = tstart, 
  time2 = tstop,
  event = death) ~ 
    dvt_event + 
    age_when_cancer_dx + # Adjust
    gender + 
    primary_tumor_simplified +
    tnm_stage_grouped +
    #strata(primary_tumor_simplified, tnm_stage_grouped) +
    cluster(id), 
  id = id,
  data = dvt_no_vte_tmerge)

cox.zph(dvt_no_vte_cph_fit)

# Log-rank p-value
dvt_no_vte_cph_tests_summary = summary(dvt_no_vte_cph_fit)

dvt_no_vte_pval_robust = format(
  dvt_no_vte_cph_tests_summary$robscore[["pvalue"]], 
  digits = 1)

dvt_no_vte_pval_robust

```


```{r}
#jpeg("./graphs/paper/dvt_no_vte_surv_analysis_KM_curve.jpeg", width = 7, height = 4, units = 'in', res = 700)
dvt_no_vte_km_plot = 
  ggsurvplot(fit = dvt_no_vte_survfit,
           data = dvt_no_vte_tmerge,
           xscale = 30,
           break.time.by = 60,
           censor = FALSE,
           conf.int = FALSE,
           #conf.int.alpha = 0.15,
           xlim = c(0, 540),
           surv.plot.height = 1.25, 
           tables.height = 0.25,
           legend.title = " ",
           legend.labs = c("No VTE", "DVT"),
           risk.table = "abs_pct", 
           risk.table.height = 0.35, 
           fontsize = 3,
           risk.table.col = "strata", 
           risk.table.y.text.col = TRUE,
           #linetype = "strata", 
           surv.median.line = "hv",
           xlab = "Time (months)",
           ylab = "OS Probability",
           ggtheme = theme_bw(),
           palette = c("black", "red"))

# Add p-value and n of patient subgroups
dvt_no_vte_km_plot$plot = dvt_no_vte_km_plot$plot + 
  # Log-rank and p-value
  annotate(
    geom = "text", 
    size = 3.5,
    x = 26,
    y = 0.3,
    label = "Log-rank",
    color = "black") + 
  annotate(
    geom = "text", 
    size = 3.5,
    x = 28,
    y = 0.2,
    label = "<0.0001",
    color = "black") + 
  # N subgroups
  annotate(
    geom = "text",
    x = 540, 
    y = 0.95, 
    label = paste0("n=", length(id_no_vte)), 
    color = "black") + 
  annotate(
    geom = "text",
    x = 540, 
    y = 0.85, 
    label = paste0("n=", length(id_only_dvt)), 
    color = "red")

# Customize risk table title 
dvt_no_vte_km_plot$table = 
  dvt_no_vte_km_plot$table +
  theme(plot.title = element_text(size = 10))

dvt_no_vte_km_plot 

#dev.off()
```


```{r}
dvt_no_vte_cph_fit %>% 
  gtsummary::tbl_regression(exp = TRUE)
```

```{r}
summary(survfit(Surv(
  time = tstart, 
  time2 = tstop,
  event = death) ~ dvt_event + cluster(id), 
    id = id,
  data = dvt_no_vte_tmerge), times = 18*30.25)
```



## DVT interactions

```{r}
# DVT effect per cancer type
dvt_interaction_primary_tumor = coxph(Surv(
  time = tstart, 
  time2 = tstop,
  event = death) ~ 
    dvt_event:primary_tumor_simplified + 
    age_when_cancer_dx + # Adjust
    gender + 
    #primary_tumor_simplified +
    tnm_stage_grouped +
    #strata(primary_tumor_simplified, tnm_stage_grouped) +
    cluster(id), 
  id = id,
  data = dvt_no_vte_tmerge) 

dvt_interaction_primary_tumor %>% 
  gtsummary::tbl_regression(exp = TRUE)
```

```{r}
cox.zph(dvt_interaction_primary_tumor, terms = FALSE)
```

```{r}
#ggcoxzph(cox.zph(dvt_interaction_primary_tumor, terms = FALSE))[5]
```


```{r}
# DVT effect per stage
dvt_interaction_stage = coxph(Surv(
  time = tstart, 
  time2 = tstop,
  event = death) ~ 
    dvt_event:tnm_stage_grouped + 
    age_when_cancer_dx + # Adjust
    gender + 
    primary_tumor_simplified +
    #tnm_stage_grouped +
    #strata(primary_tumor_simplified, tnm_stage_grouped) +
    cluster(id), 
  id = id,
  data = dvt_no_vte_tmerge) 

dvt_interaction_stage %>% 
  gtsummary::tbl_regression(exp = TRUE)
```

```{r}
cox.zph(dvt_interaction_stage, terms = FALSE)
```


```{r}
# DVT effect per stage
dvt_interaction_primary_tumor_stage = coxph(Surv(
  time = tstart, 
  time2 = tstop,
  event = death) ~ 
    dvt_event:primary_tumor_simplified:tnm_stage_grouped + 
    age_when_cancer_dx + # Adjust
    gender + 
    #primary_tumor_simplified +
    #tnm_stage_grouped +
    #strata(primary_tumor_simplified, tnm_stage_grouped) +
    cluster(id), 
  id = id,
  data = dvt_no_vte_tmerge) 

dvt_interaction_primary_tumor_stage %>% 
  gtsummary::tbl_regression(exp = TRUE)
```


```{r}
cox.zph(dvt_interaction_primary_tumor_stage, terms = FALSE)
```

```{r}
table(dvt_no_vte_tmerge$primary_tumor_simplified, dvt_no_vte_tmerge$dvt_event)
table(dvt_no_vte_tmerge$tnm_stage_grouped, dvt_no_vte_tmerge$dvt_event)
```

No DVT events happened in patients with oesophago-gastric cancer. Similarly, no DVT events happened in patients with stages I-II.  

Only 1 DVT happened in a patient with pancreatic cancer, so HR cannot be computed.  


# PTE vs no VTE

```{r}
pte_no_vte_tmerge = all_types_vte_tmerge %>%
  group_by(id) %>%
  filter(id %in% c(id_only_pte, id_no_vte)) 
```


```{r}
pte_no_vte_survfit = survfit(Surv(
  time = tstart, 
  time2 = tstop,
  event = death) ~ pte_event + cluster(id), 
    id = id,
  data = pte_no_vte_tmerge)

pte_no_vte_survfit
```


```{r}
# Cox regression
pte_no_vte_cph_fit = coxph(Surv(
  time = tstart, 
  time2 = tstop,
  event = death) ~ 
    pte_event + 
    age_when_cancer_dx + # Adjust by age and gender
    gender + 
    primary_tumor_simplified +
    tnm_stage_grouped +
    #strata(primary_tumor_simplified, tnm_stage_grouped) +
    cluster(id), 
  id = id,
  data = pte_no_vte_tmerge)

cox.zph(pte_no_vte_cph_fit)

# Log-rank p-value
pte_no_vte_cph_tests_summary = summary(pte_no_vte_cph_fit)

pte_no_vte_pval_robust = format(
  pte_no_vte_cph_tests_summary$robscore[["pvalue"]], 
  digits = 1)

pte_no_vte_pval_robust
```


```{r}
#jpeg("../graphs/paper/pte_no_vte_td_surv_analysis_KM_curve.jpeg", width = 7, height = 4, units = 'in', res = 700)
pte_no_vte_km_plot = 
  ggsurvplot(fit = pte_no_vte_survfit,
           data = pte_no_vte_tmerge,
           xscale = 30,
           break.time.by = 60,
           censor = FALSE,
           conf.int = FALSE,
           #conf.int.alpha = 0.15,
           xlim = c(0, 540),
           surv.plot.height = 1.25, 
           tables.height = 0.25,
           legend.title = " ",
           legend.labs = c("No VTE", "PE"),
           risk.table = "abs_pct", 
           risk.table.height = 0.3, 
           fontsize = 3,
           risk.table.col = "strata", 
           risk.table.y.text.col = TRUE,
           surv.median.line = "hv",
           xlab = "Time (months)",
           ylab = "OS Probability",
           ggtheme = theme_bw(),
           palette = c("black", "red"))

# Add p-value
pte_no_vte_km_plot$plot = 
  pte_no_vte_km_plot$plot + 
  # Log-rank and p-value
  annotate(
    geom = "text", 
    size = 3.5, 
    x = 26,
    y = 0.3,
    label = "Log-rank",
    color = "black") + 
  annotate(
    geom = "text", 
    size = 3.5,
    x = 28,
    y = 0.2,
    label = "<0.0001",
    color = "black") + 
  # N subgroups
  annotate(
    geom = "text",
    x = 540, 
    y = 0.95, 
    label = paste0("n=", length(id_no_vte)), 
    color = "black") + 
  annotate(
    geom = "text",
    x = 540, 
    y = 0.85, 
    label = paste0("n=", length(id_only_pte)), 
    color = "red")

# Customize risk table title 
pte_no_vte_km_plot$table = 
  pte_no_vte_km_plot$table +
  theme(plot.title = element_text(size = 10))

pte_no_vte_km_plot 

#dev.off()
```

```{r}
pte_no_vte_cph_fit %>% 
  gtsummary::tbl_regression(exp = TRUE)
```


```{r}
summary(survfit(Surv(
  time = tstart, 
  time2 = tstop,
  event = death) ~ pte_event + cluster(id), 
    id = id,
  data = pte_no_vte_tmerge), times = 18*30.25)
```

## PTE interactions

```{r}
# PTE effect per cancer type
pte_interaction_primary_tumor = coxph(Surv(
  time = tstart, 
  time2 = tstop,
  event = death) ~ 
    pte_event:primary_tumor_simplified + 
    age_when_cancer_dx + # Adjust
    gender + 
    #primary_tumor_simplified +
    tnm_stage_grouped +
    #strata(primary_tumor_simplified, tnm_stage_grouped) +
    cluster(id), 
  id = id,
  data = pte_no_vte_tmerge) 

pte_interaction_primary_tumor %>% 
  gtsummary::tbl_regression(exp = TRUE)
```


```{r}
cox.zph(pte_interaction_primary_tumor, terms = FALSE)
```

```{r}
# This is why HR for PTE and pancreatic cancer cannot be computed
# There is only one event
table(pte_no_vte_tmerge$primary_tumor_simplified, pte_no_vte_tmerge$pte_event)
```



```{r}
#ggcoxzph(cox.zph(dvt_interaction_primary_tumor, terms = FALSE))[5]
```


```{r}
# PTE effect per stage
pte_interaction_stage = coxph(Surv(
  time = tstart, 
  time2 = tstop,
  event = death) ~ 
    pte_event:tnm_stage_grouped + 
    age_when_cancer_dx + # Adjust
    gender + 
    primary_tumor_simplified +
    #tnm_stage_grouped +
    #strata(primary_tumor_simplified, tnm_stage_grouped) +
    cluster(id), 
  id = id,
  data = pte_no_vte_tmerge) 

pte_interaction_stage %>% 
  gtsummary::tbl_regression(exp = TRUE)
```

Suffering a PTE in stage IV increases 10-fold the probability of dying. 


```{r}
cox.zph(pte_interaction_stage, terms = FALSE)
```

```{r}
# PTE effect per cancer and stage
pte_interaction_primary_tumor_stage = coxph(Surv(
  time = tstart, 
  time2 = tstop,
  event = death) ~ 
    pte_event:primary_tumor_simplified:tnm_stage_grouped + 
    age_when_cancer_dx + # Adjust
    gender + 
    #primary_tumor_simplified +
    #tnm_stage_grouped +
    #strata(primary_tumor_simplified, tnm_stage_grouped) +
    cluster(id), 
  id = id,
  data = pte_no_vte_tmerge) 

pte_interaction_primary_tumor_stage %>% 
  gtsummary::tbl_regression(exp = TRUE)
```

There is multicollinearity in this analysis.  

```{r}
cox.zph(pte_interaction_primary_tumor_stage)
```


# Visceral thromboses vs no VTE

```{r}
visceral_thromb_no_vte_tmerge = all_types_vte_tmerge %>%
  group_by(id) %>%
  filter(id %in% c(id_only_visceral_thromb, id_no_vte)) 
```


```{r}
visceral_thromb_no_vte_survfit = survfit(Surv(
  time = tstart, 
  time2 = tstop,
  event = death) ~ visceral_thromb_event + cluster(id), 
    id = id,
  data = visceral_thromb_no_vte_tmerge)

visceral_thromb_no_vte_survfit
```


```{r}
# Cox regression
visceral_thromb_no_vte_cph_fit = coxph(Surv(
  time = tstart, 
  time2 = tstop,
  event = death) ~ 
    visceral_thromb_event +
    age_when_cancer_dx + # Adjust 
    gender + 
    primary_tumor_simplified +
    tnm_stage_grouped +
    #strata(primary_tumor_simplified, tnm_stage_grouped) +
    cluster(id), 
  id = id,
  data = visceral_thromb_no_vte_tmerge)

cox.zph(visceral_thromb_no_vte_cph_fit)

# Log-rank p-value
visceral_thromb_no_vte_cph_tests_summary = summary(visceral_thromb_no_vte_cph_fit)

visceral_thromb_no_vte_pval_robust = format(
  visceral_thromb_no_vte_cph_tests_summary$robscore[["pvalue"]], 
  digits = 1)

visceral_thromb_no_vte_pval_robust

```


```{r}
#jpeg("./graphs/paper/visceral_thromb_no_vte_td_surv_analysis_KM_curve.jpeg", width = 7, height = 4, units = 'in', res = 700)

visceral_thromb_no_vte_km_plot = 
  ggsurvplot(fit = visceral_thromb_no_vte_survfit,
           data = visceral_thromb_no_vte_tmerge,
           xscale = 30,
           break.time.by = 60,
           censor = FALSE,
           conf.int = FALSE,
           xlim = c(0, 540),
           surv.plot.height = 1.25, 
           tables.height = 0.25,
           legend.title = " ",
           legend.labs = c("No VTE", "VT"),
           risk.table = "abs_pct", 
           risk.table.height = 0.3, 
           fontsize = 3,
           risk.table.col = "strata", 
           risk.table.y.text.col = TRUE,
           surv.median.line = "hv",
           xlab = "Time (months)",
           ylab = "OS Probability",
           ggtheme = theme_bw(),
           palette = c("black", "red")
           )

# Add p-value
visceral_thromb_no_vte_km_plot$plot = 
  visceral_thromb_no_vte_km_plot$plot + 
  # Log-rank and p-value
  annotate(
    geom = "text", 
    size = 3.5,
    x = 26,
    y = 0.3,
    label = "Log-rank",
    color = "black") + 
  annotate(
    geom = "text",
    size = 3.5,
    x = 28,
    y = 0.2,
    label = "<0.0001",
    color = "black") + 
  # N subgroups
  annotate(
    geom = "text",
    x = 540, 
    y = 0.95, 
    label = paste0("n=", length(id_no_vte)), 
    color = "black") + 
  annotate(
    geom = "text",
    x = 540, 
    y = 0.85, 
    label = paste0("n=", length(id_only_visceral_thromb)), 
    color = "red")

# Customize risk table title 
visceral_thromb_no_vte_km_plot$table = 
  visceral_thromb_no_vte_km_plot$table +
  theme(plot.title = element_text(size = 10))

visceral_thromb_no_vte_km_plot 

#dev.off()
```

```{r}
visceral_thromb_no_vte_cph_fit %>% 
  gtsummary::tbl_regression(exp = TRUE)
```

VT is not proven to be a risk factor.  

```{r}
summary(survfit(Surv(
  time = tstart, 
  time2 = tstop,
  event = death) ~ visceral_thromb_event + cluster(id), 
    id = id,
  data = visceral_thromb_no_vte_tmerge), times = 18*30.25)
```


## VT interactions

```{r}
# VT effect per cancer type
visceral_thromb_interaction_primary_tumor = coxph(Surv(
  time = tstart, 
  time2 = tstop,
  event = death) ~ 
    visceral_thromb_event:primary_tumor_simplified + 
    age_when_cancer_dx + # Adjust
    gender + 
    #primary_tumor_simplified +
    tnm_stage_grouped +
    #strata(primary_tumor_simplified, tnm_stage_grouped) +
    cluster(id), 
  id = id,
  data = visceral_thromb_no_vte_tmerge) 

visceral_thromb_interaction_primary_tumor %>% 
  gtsummary::tbl_regression(exp = TRUE)
```


```{r}
cox.zph(visceral_thromb_interaction_primary_tumor, terms = FALSE)
```

There is only 1 visceral thrombosis in NSCLC patients and 2 in oesophago-gastric patients.  

```{r}
# VT effect per stage
visceral_thromb_interaction_stage = coxph(Surv(
  time = tstart, 
  time2 = tstop,
  event = death) ~ 
    visceral_thromb_event:tnm_stage_grouped + 
    age_when_cancer_dx + # Adjust
    gender + 
    primary_tumor_simplified +
    #tnm_stage_grouped +
    #strata(primary_tumor_simplified, tnm_stage_grouped) +
    cluster(id), 
  id = id,
  data = visceral_thromb_no_vte_tmerge) 

visceral_thromb_interaction_stage %>% 
  gtsummary::tbl_regression(exp = TRUE)
```

```{r}
cox.zph(visceral_thromb_interaction_stage, terms = FALSE)
```


There are 2 visceral thrombosis in stages I-II, so the corresponding results are not very reliable.  

```{r}
# VT effect per stage
visceral_thromb_interaction_primary_tumor_stage = coxph(Surv(
  time = tstart, 
  time2 = tstop,
  event = death) ~ 
    visceral_thromb_event:primary_tumor_simplified:tnm_stage_grouped + 
    age_when_cancer_dx + # Adjust
    gender + 
    #primary_tumor_simplified +
    #tnm_stage_grouped +
    #strata(primary_tumor_simplified, tnm_stage_grouped) +
    cluster(id), 
  id = id,
  data = visceral_thromb_no_vte_tmerge) 

visceral_thromb_interaction_primary_tumor_stage %>% 
  gtsummary::tbl_regression(exp = TRUE)
```

```{r}
cox.zph(visceral_thromb_interaction_primary_tumor_stage, terms = FALSE)
```


```{r}
table(visceral_thromb_no_vte_tmerge$primary_tumor_simplified, visceral_thromb_no_vte_tmerge$visceral_thromb_event)

table(visceral_thromb_no_vte_tmerge$tnm_stage_grouped, visceral_thromb_no_vte_tmerge$visceral_thromb_event)
```

There are only 1 VT in a NSCLC patient and 2 VT in oesophago-gastric patients. Again, only 2 VT in stages I-III.  


# Other thromboses vs no VTE

```{r}
other_thromb_grouped_no_vte_tmerge = all_types_vte_tmerge %>%
  group_by(id) %>%
  filter(id %in% c(id_only_other_thromb_grouped, id_no_vte)) 
```


```{r}
other_thromb_grouped_no_vte_survfit = survfit(Surv(
  time = tstart, 
  time2 = tstop,
  event = death) ~ other_thromb_grouped_event + cluster(id), 
    id = id,
  data = other_thromb_grouped_no_vte_tmerge)

other_thromb_grouped_no_vte_survfit
```


```{r}
# Cox regression
other_thromb_grouped_no_vte_cph_fit = coxph(Surv(
  time = tstart, 
  time2 = tstop,
  event = death) ~ 
    other_thromb_grouped_event + 
    age_when_cancer_dx + # Adjust
    gender + 
    primary_tumor_simplified + 
    tnm_stage_grouped +
    #strata(primary_tumor_simplified, tnm_stage_grouped) +
    cluster(id), 
  id = id,
  data = other_thromb_grouped_no_vte_tmerge)

cox.zph(other_thromb_grouped_no_vte_cph_fit)

# Log-rank p-value
other_thromb_grouped_no_vte_cph_tests_summary = summary(other_thromb_grouped_no_vte_cph_fit)

other_thromb_grouped_no_vte_pval_robust = format(
  other_thromb_grouped_no_vte_cph_tests_summary$robscore[["pvalue"]], 
  digits = 1)

other_thromb_grouped_no_vte_pval_robust
```


```{r}
#jpeg("./graphs/paper/other_thromb_grouped_no_vte_td_surv_analysis_KM_curve.jpeg", width = 7, height = 4, units = 'in', res = 700)
other_thromb_grouped_no_vte_km_plot = 
  ggsurvplot(fit = other_thromb_grouped_no_vte_survfit,
           data = other_thromb_grouped_no_vte_tmerge,
           xscale = 30,
           break.time.by = 60,
           censor = FALSE,
           conf.int = FALSE,
           #conf.int.alpha = 0.15,
           xlim = c(0, 540),
           surv.plot.height = 1.25, 
           tables.height = 0.25,
           legend.title = " ",
           legend.labs = c("No VTE", "Other VTEs"),
           risk.table = "abs_pct", 
           risk.table.height = 0.3, 
           fontsize = 3,
           risk.table.col = "strata", 
           risk.table.y.text.col = TRUE,
           surv.median.line = "hv",
           xlab = "Time (months)",
           ylab = "OS Probability",
           ggtheme = theme_bw(),
           palette = c("black", "red"))

# Count n by subgroups
n_no_vte = 301
n_only_other_thromb_grouped = length(unique(other_thromb_grouped_no_vte_tmerge$id)) - n_no_vte

# Add p-value
other_thromb_grouped_no_vte_km_plot$plot = other_thromb_grouped_no_vte_km_plot$plot + 
  annotate(
    geom = "text",
    size = 3.5,
    x = 26,
    y = 0.3,
    label = "Log-rank",
    color = "black") + 
  annotate(
    geom = "text", 
    size = 3.5,
    x = 28,
    y = 0.2,
    label = "<0.0001",
    color = "black") + 
  annotate(
    geom = "text",
    x = 540, 
    y = 0.95, 
    label = paste0("n=", length(id_no_vte)), 
    color = "black") + 
  annotate(
    geom = "text",
    x = 540, 
    y = 0.85, 
    label = paste0("n=", length(id_only_other_thromb_grouped)), 
    color = "red")

# Customize risk table title 
other_thromb_grouped_no_vte_km_plot$table = 
  other_thromb_grouped_no_vte_km_plot$table +
  theme(plot.title = element_text(size = 10))

other_thromb_grouped_no_vte_km_plot 

#dev.off()
```


```{r}
other_thromb_grouped_no_vte_cph_fit %>% 
  gtsummary::tbl_regression(exp = TRUE)
```

Other types of VTE (CRT, SVT, etc.) are not proven to be risk factors.  

```{r}
summary(survfit(Surv(
  time = tstart, 
  time2 = tstop,
  event = death) ~ other_thromb_grouped_event + cluster(id), 
    id = id,
  data = other_thromb_grouped_no_vte_tmerge), times = 18*30.25)
```


## Other thromboses interactions

## PTE interactions

```{r}
# PTE effect per cancer type
other_thromb_grouped_interaction_primary_tumor = coxph(Surv(
  time = tstart, 
  time2 = tstop,
  event = death) ~ 
    other_thromb_grouped_event:primary_tumor_simplified + 
    age_when_cancer_dx + # Adjust
    gender + 
    #primary_tumor_simplified +
    tnm_stage_grouped +
    #strata(primary_tumor_simplified, tnm_stage_grouped) +
    cluster(id), 
  id = id,
  data = other_thromb_grouped_no_vte_tmerge) 

other_thromb_grouped_interaction_primary_tumor %>% 
  gtsummary::tbl_regression(exp = TRUE)
```


```{r}
cox.zph(other_thromb_grouped_interaction_primary_tumor, terms = FALSE)
```


```{r}
# This is why HR for PTE and pancreatic cancer cannot be computed
# There is only one event
table(other_thromb_grouped_no_vte_tmerge$primary_tumor_simplified, other_thromb_grouped_no_vte_tmerge$other_thromb_grouped_event)
```



```{r}
# PTE effect per stage
other_thromb_grouped_interaction_stage = coxph(Surv(
  time = tstart, 
  time2 = tstop,
  event = death) ~ 
    other_thromb_grouped_event:tnm_stage_grouped + 
    age_when_cancer_dx + # Adjust
    gender + 
    primary_tumor_simplified +
    #tnm_stage_grouped +
    #strata(primary_tumor_simplified, tnm_stage_grouped) +
    cluster(id), 
  id = id,
  data = other_thromb_grouped_no_vte_tmerge) 

other_thromb_grouped_interaction_stage %>% 
  gtsummary::tbl_regression(exp = TRUE)
```


```{r}
cox.zph(other_thromb_grouped_interaction_stage, terms = FALSE)
```

```{r}
# PTE effect per cancer and stage
other_thromb_grouped_interaction_primary_tumor_stage = coxph(Surv(
  time = tstart, 
  time2 = tstop,
  event = death) ~ 
    other_thromb_grouped_event:primary_tumor_simplified:tnm_stage_grouped + 
    age_when_cancer_dx + # Adjust
    gender + 
    #primary_tumor_simplified +
    #tnm_stage_grouped +
    #strata(primary_tumor_simplified, tnm_stage_grouped) +
    cluster(id), 
  id = id,
  data = other_thromb_grouped_no_vte_tmerge) 

other_thromb_grouped_interaction_primary_tumor_stage %>% 
  gtsummary::tbl_regression(exp = TRUE)
```

There is multicollinearity in this analysis.  

```{r}
cox.zph(other_thromb_grouped_interaction_primary_tumor_stage)
```

# All single-type VTEs vs no VTE

```{r}
single_type_vte_tmerge = all_types_vte_tmerge %>%
  group_by(id) %>%
  filter(id %in% c(id_no_vte, 
                   id_only_dvt, 
                   id_only_pte, 
                   id_only_visceral_thromb, 
                   id_only_other_thromb_grouped))
```


```{r}
single_type_vte_no_vte_survfit = survfit(Surv(
  time = tstart, 
  time2 = tstop,
  event = death) ~ dvt_event +
    pte_event + 
    visceral_thromb_event + 
    other_thromb_grouped_event +
    cluster(id), 
    id = id,
  data = single_type_vte_tmerge)

#single_type_vte_no_vte_survfit
```


```{r}
# Cox regression
single_type_vte_no_vte_cph_fit = coxph(Surv(
  time = tstart, 
  time2 = tstop,
  event = death) ~ 
    dvt_event + 
    pte_event + 
    visceral_thromb_event + 
    other_thromb_grouped_event +
    age_when_cancer_dx + # Adjust
    gender + 
    #primary_tumor_simplified +
    #tnm_stage_grouped +
    strata(primary_tumor_simplified, tnm_stage_grouped) +
    cluster(id), 
  id = id,
  data = single_type_vte_tmerge)

cox.zph(single_type_vte_no_vte_cph_fit)

# Log-rank p-value
single_type_vte_no_vte_cph_tests_summary = summary(single_type_vte_no_vte_cph_fit)

single_type_vte_no_vte_pval_robust = format(
  single_type_vte_no_vte_cph_tests_summary$robscore[["pvalue"]], 
  digits = 1)

single_type_vte_no_vte_pval_robust

```


```{r}
#jpeg("./graphs/time_dependent_covariate_analysis/single_type_vte_no_vte_surv_analysis_KM_curve.jpeg", width = 6.5, height = 4, units = 'in', res = 700)
single_type_vte_no_vte_km_plot = 
  ggsurvplot(
    fit = single_type_vte_no_vte_survfit,
    data = single_type_vte_tmerge,
    xscale = 30,
    break.time.by = 60,
    censor = FALSE,
    conf.int = FALSE,
    #conf.int.alpha = 0.15,
    xlim = c(0, 540),
    surv.plot.height = 1.25, 
    tables.height = 0.25,
    legend.title = " ",
    legend.labs = c("No VTE", "Other", "VT", "PTE", "DVT"),
    risk.table = "abs_pct", 
    risk.table.height = 0.35, 
    fontsize = 3.5,
    risk.table.col = "strata", 
    risk.table.y.text.col = TRUE,
    #linetype = "strata", 
    surv.median.line = "hv",
    xlab = "Time (months)",
    ylab = "OS Probability",
    ggtheme = theme_bw(),
    palette = c("black", "red", "orange", "pink", "blue")
    )

# Add p-value and n of patient subgroups
single_type_vte_no_vte_km_plot$plot = 
  single_type_vte_no_vte_km_plot$plot + 
  # Log-rank and p-value
  annotate(
    geom = "text", 
    x = 26,
    y = 0.3,
    label = "Log-rank",
    color = "black") + 
  annotate(
    geom = "text", 
    x = 28,
    y = 0.2,
    label = "0.003",
    color = "black") 

# Customize risk table title 
single_type_vte_no_vte_km_plot$table = 
  single_type_vte_no_vte_km_plot$table +
  theme(plot.title = element_text(size = 10))

single_type_vte_no_vte_km_plot 

#dev.off()
```

```{r}
single_type_vte_no_vte_cph_fit %>% 
  gtsummary::tbl_regression(exp = TRUE)
```



# Multiple types of VTE vs no VTE


```{r}
vte_combinations_no_vte_tmerge = survival_vte_tmerge %>%
  group_by(id) %>%
  filter(id %in% c(id_no_vte, 
                   id_dvt_and_pte, 
                   id_dvt_and_visceral_thromb, 
                   id_dvt_and_other_thromb_grouped, 
                   id_pte_and_visceral_thromb, 
                   id_vt_and_other_thromb_grouped, 
                   id_dvt_and_pte_and_other_thromb_grouped, 
                   id_all_types_vte)) 
```


```{r}
vte_combinations_no_vte_survfit = survfit(Surv(
  time = tstart, 
  time2 = tstop,
  event = death) ~ vte_event +
    cluster(id), 
    id = id,
  data = vte_combinations_no_vte_tmerge)

vte_combinations_no_vte_survfit
```


```{r}
# Cox regression
vte_combinations_no_vte_cph_fit = coxph(Surv(
  time = tstart, 
  time2 = tstop,
  event = death) ~ 
    vte_event + 
    age_when_cancer_dx + # Adjust
    gender + 
    primary_tumor_simplified + 
    tnm_stage_grouped +
    #strata(primary_tumor_simplified, tnm_stage_grouped) +
    cluster(id), 
  id = id,
  data = vte_combinations_no_vte_tmerge)

cox.zph(vte_combinations_no_vte_cph_fit)

# Log-rank p-value
vte_combinations_no_vte_cph_tests_summary = summary(vte_combinations_no_vte_cph_fit)

vte_combinations_no_vte_pval_robust = format(
  vte_combinations_no_vte_cph_tests_summary$robscore[["pvalue"]], 
  digits = 1)

vte_combinations_no_vte_pval_robust
```


```{r}
#jpeg("./graphs/paper/multiple_vte_types_no_vte_td_surv_analysis_KM_curve.jpeg", width = 7, height = 4, units = 'in', res = 700)
vte_combinations_no_vte_km_plot = 
  ggsurvplot(fit = vte_combinations_no_vte_survfit,
           data = vte_combinations_no_vte_tmerge,
           xscale = 30,
           break.time.by = 60,
           censor = FALSE,
           conf.int = FALSE,
           #conf.int.alpha = 0.15,
           xlim = c(0, 540),
           surv.plot.height = 1.25, 
           tables.height = 0.25,
           legend.title = " ",
           legend.labs = c("No VTE", "Multiple VTEs"),
           risk.table = "abs_pct", 
           risk.table.height = 0.3, 
           fontsize = 3,
           risk.table.col = "strata", 
           risk.table.y.text.col = TRUE,
           surv.median.line = "hv",
           xlab = "Time (months)",
           ylab = "OS Probability",
           ggtheme = theme_bw(),
           palette = c("black", "red"))

# Count n by subgroups
n_no_vte = 301
n_only_vte_combinations = length(unique(vte_combinations_no_vte_tmerge$id)) - n_no_vte

# Add p-value
vte_combinations_no_vte_km_plot$plot = 
  vte_combinations_no_vte_km_plot$plot + 
  annotate(
    geom = "text", 
    size = 3.5,
    x = 26,
    y = 0.3,
    label = "Log-rank",
    color = "black") + 
  annotate(
    geom = "text", 
    size = 3.5,
    x = 28,
    y = 0.2,
    label = "<0.0001",
    color = "black") + 
  annotate(
    geom = "text",
    x = 540, 
    y = 0.95, 
    label = paste0("n=", length(id_no_vte)), 
    color = "black") + 
  annotate(
    geom = "text",
    x = 540, 
    y = 0.85, 
    label = "n=23", 
    color = "red")

# Customize risk table title 
vte_combinations_no_vte_km_plot$table = 
  vte_combinations_no_vte_km_plot$table +
  theme(plot.title = element_text(size = 10))

vte_combinations_no_vte_km_plot 

#dev.off()
```

```{r}
vte_combinations_no_vte_cph_fit %>% 
  gtsummary::tbl_regression(exp = TRUE)
```


```{r}
summary(survfit(Surv(
  time = tstart, 
  time2 = tstop,
  event = death) ~ vte_event +
    cluster(id), 
    id = id,
  data = vte_combinations_no_vte_tmerge), times = 18*30.25)
```

# KM plot with single and multiple VTE types, all together


```{r}
# marrangeGrob(list(dvt_no_vte_km_plot$plot, 
#   pte_no_vte_km_plot$plot, 
#   visceral_thromb_no_vte_km_plot$plot, 
#   other_thromb_grouped_no_vte_km_plot$plot,
#   vte_combinations_no_vte_km_plot$plot), nrow = 5, ncol = 1)
```



---

```{r}
dvt_primary_tumor_cph_fit = coxph(Surv(
  time = tstart, 
  time2 = tstop,
  event = death) ~ 
    age_when_cancer_dx +
    gender +
    #primary_tumor_simplified + 
    #tnm_stage_grouped +
    #strata(tnm_stage_grouped) + 
    dvt_event:primary_tumor_simplified:tnm_stage_two_groups + 
    #pte_event + 
    #visceral_thromb_event + 
    #other_thromb_grouped_event + 
    cluster(id), 
  id = id,
  data = all_types_vte_tmerge)

dvt_primary_tumor_cph_fit %>% 
  gtsummary::tbl_regression(exp = TRUE)

```

We can draw some conclusions from here for patients in stage IV. But the data is fishy.  


```{r}
cox.zph(dvt_primary_tumor_cph_fit)
```


```{r}
concordance(dvt_primary_tumor_cph_fit)
```


```{r}
pte_primary_tumor_cph_fit = coxph(Surv(
  time = tstart, 
  time2 = tstop,
  event = death) ~ 
    age_when_cancer_dx +
    gender +
    #primary_tumor_simplified + 
    #strata(tnm_stage_grouped) + 
    #dvt_event + 
    pte_event:primary_tumor_simplified:tnm_stage_two_groups + 
    #visceral_thromb_event + 
    #other_thromb_grouped_event + 
    cluster(id), 
  id = id,
  data = all_types_vte_tmerge)
 
pte_primary_tumor_cph_fit %>% 
  gtsummary::tbl_regression(exp = TRUE)

```

We can draw interesting information from stage IV tumors and its correlation with PTE and death.  

```{r}
cox.zph(pte_primary_tumor_cph_fit)
```


```{r}
table(all_types_vte_tmerge$primary_tumor_simplified, all_types_vte_tmerge$pte_event, useNA = "always")
```

```{r}
table(all_types_vte_tmerge$tnm_stage_grouped, all_types_vte_tmerge$pte_event, useNA = "always")
```


```{r}
visceral_thromb_primary_tumor_cph_fit = coxph(Surv(
  time = tstart, 
  time2 = tstop,
  event = death) ~ 
    age_when_cancer_dx +
    gender +
    #primary_tumor_simplified + 
    #strata(tnm_stage_grouped) + 
    #dvt_event + 
    #pte_event + 
    visceral_thromb_event:primary_tumor_simplified:tnm_stage_two_groups + 
    #other_thromb_grouped_event + 
    cluster(id), 
  id = id,
  data = all_types_vte_tmerge)
 
visceral_thromb_primary_tumor_cph_fit %>% 
  gtsummary::tbl_regression(exp = TRUE)

```


```{r}
cox.zph(visceral_thromb_primary_tumor_cph_fit)
```


```{r}
other_thromb_grouped_primary_tumor_cph_fit = coxph(Surv(
  time = tstart, 
  time2 = tstop,
  event = death) ~ 
    age_when_cancer_dx +
    gender +
    #primary_tumor_simplified + 
    #strata(tnm_stage_grouped) + 
    #dvt_event + 
    #pte_event + 
    #visceral_thromb_event + 
    other_thromb_grouped_event:primary_tumor_simplified:tnm_stage_two_groups + 
    cluster(id), 
  id = id,
  data = all_types_vte_tmerge) 

other_thromb_grouped_primary_tumor_cph_fit %>% 
  gtsummary::tbl_regression(exp = TRUE)

```


```{r}
cox.zph(other_thromb_grouped_primary_tumor_cph_fit)
```


Best to do Fisher test to find correlation between cancer type and type of VTE; and do the same with stage.  


# KM incidental vs symptomatic

```{r}
# %<>%
id_no_vte_only_one_type_vte = oncoth1_data %>%
  filter(n_all_vte == 0 | n_all_vte == 1) %>%
  select(id) %>%
  unlist()
```


```{r}
covariates_df_inc_sympt = oncoth1_data %>%
  filter(id %in% id_no_vte_only_one_type_vte) %>%
  select(id, 
         age_when_cancer_dx_65, 
         gender, 
         primary_tumor_simplified, 
         tnm_stage_grouped, 
         tnm_stage_two_groups, 
         contains("_type_"),
         n_all_vte,
         tu_1st_vte_all_from_cancer_diagnosis_days_corrected, 
         tu_1st_vte_all_from_cancer_diagnosis_months_corrected, 
         censored_patient,
         follow_up_length_days_from_cancer_diagnosis, 
         follow_up_length_months_from_cancer_diagnosis) %>%
  mutate(first_vte_inc_symp = coalesce(
    cancer_associated_ledvt_type_vte, 
    cancer_associated_uedvt_type_vte, 
    cancer_associated_pte_type_vte, 
    cancer_associated_visceral_thromb_type_vte, 
    eval_crt_type_vte_along_study,
    eval_ledvt_type_vte_along_study, 
    eval_uedvt_type_vte_along_study, 
    eval_pte_type_vte_along_study,
    eval_svt_type_vte_along_study, 
    eval_visceral_thromb_type_vte_along_study, 
    eval_other_thromb_type_vte_along_study), 
    .after = n_all_vte) %>%
  filter(!c(n_all_vte == 1 & is.na(first_vte_inc_symp))) %>%
  mutate(first_vte_inc_symp = case_when(
    n_all_vte == 0 ~ "No CAT", 
    TRUE ~ first_vte_inc_symp)) %>%
  mutate(first_vte_inc_symp = factor(
    x = first_vte_inc_symp, 
    levels = c("No CAT", "Incidental", "Symptomatic"), 
    labels = c("No CAT", "Incidental", "Symptomatic")))
```


```{r}
survival_vte_tmerge_inc_sympt = tmerge(
  data1 = covariates_df_inc_sympt, 
  data2 = covariates_df_inc_sympt, 
  id = id, 
  tstop = follow_up_length_days_from_cancer_diagnosis, 
  # event() is the survival analysis event of interest (in our case, death)
  death = event(follow_up_length_days_from_cancer_diagnosis, censored_patient), 
  # tdc() are for time-dependent covariates (VTEs)
  vte_event = tdc(tu_1st_vte_all_from_cancer_diagnosis_days_corrected)
  )
```


```{r}
inc_symp_km_estimator_fit = survfit(Surv(
  time = tstart, 
  time2 = tstop, 
  event = death) ~ 
    first_vte_inc_symp + 
    cluster(id), 
  id = id,
  data = survival_vte_tmerge_inc_sympt)

inc_symp_km_estimator_fit
```


```{r}
# We perform the log-rank test separately, because we need to manually insert the
# p-value into the KM plot
inc_symp_cph_fit = coxph(Surv(
  time = tstart, 
  time2 = tstop, 
  event = death) ~ 
    first_vte_inc_symp + 
    #strata(tnm_stage_grouped) + #test this primary_tumor_simplified
    cluster(id),
  data = survival_vte_tmerge_inc_sympt)

inc_symp_cph_tests_summary = summary(inc_symp_cph_fit)

inc_symp_cph_tests_summary$robscore[["pvalue"]]

```


```{r}
#jpeg("./graphs/time_dependent_covariate_analysis/vte_td_surv_analysis_KM_curve.jpeg", width = 6.5, height = 4, units = 'in', res = 700)
inc_symp_td_survival_km_plot = 
  ggsurvplot(
    fit = inc_symp_km_estimator_fit,
    data = survival_vte_tmerge_inc_sympt,
    xscale = 30,
    break.time.by = 60,
    censor = FALSE, 
    conf.int = FALSE,
    #conf.int.alpha = 0.15,
    xlim = c(0, 540),
    surv.plot.height = 1.25, 
    tables.height = 0.25,
    legend.title = " ",
    legend.labs = c("No CAT", "Incidental", "Symptomatic"),
    risk.table = "abs_pct", 
    risk.table.height = 0.3, 
    fontsize = 3,
    risk.table.col = "strata", 
    risk.table.y.text.col = TRUE,
    #linetype = "strata", 
    surv.median.line = "hv",
    xlab = "Time (months)",
    ylab = "OS Probability",
    ggtheme = theme_bw(),
    palette = c("black", "magenta", "blue"))

# Add p-value
inc_symp_td_survival_km_plot$plot = inc_symp_td_survival_km_plot$plot + 
  annotate(
    geom = "text", 
    x = 26,
    y = 0.3,
    label = "Log-rank",
    color = "black") + 
  annotate(
    geom = "text", 
    x = 28,
    y = 0.2,
    label = as.character(format(
      inc_symp_cph_tests_summary$robscore[["pvalue"]], 
      scientific = FALSE, 
      digits = 1)),
    color = "black")

# Customize risk table title 
inc_symp_td_survival_km_plot$table = 
  inc_symp_td_survival_km_plot$table +
  theme(plot.title = element_text(size = 10))

inc_symp_td_survival_km_plot

#dev.off()
```


```{r}
# We perform the log-rank test separately, because we need to manually insert the
# p-value into the KM plot
adj_inc_symp_cph_fit = coxph(Surv(
  time = tstart, 
  time2 = tstop, 
  event = death) ~ 
    first_vte_inc_symp:vte_event + 
    age_when_cancer_dx_65 +
    gender + 
    # create stratified model -- different cancer types and stages have different impact on OS
    strata(primary_tumor_simplified, tnm_stage_grouped) + 
    cluster(id),
  data = survival_vte_tmerge_inc_sympt)

cox.zph(adj_inc_symp_cph_fit)

adj_inc_symp_cph_fit %>% 
  gtsummary::tbl_regression(exp = TRUE)

```


---


# Survival analysis for patients with no VTE *vs* patients with VTE but no recurrence *vs* patients with VTE recurrence

Show three survival lines in the same KM plot.  

Recurrence is also a time-dependent covariate. 

```{r}
no_vte_vs_vte_vs_recurrence_fit = survfit(Surv(
  time = tstart, 
  time2 = tstop,
  event = death) ~ vte_recurrence + cluster(id), 
    id = id,
  data = survival_vte_tmerge)

no_vte_vs_vte_vs_recurrence_fit
```


```{r}
no_vte_vs_vte_vs_recurrence_cph_reg = coxph(Surv(
  time = tstart, 
  time2 = tstop, 
  event = death) ~ vte_event:vte_recurrence + 
    cluster(id),
  id = id,
  data = survival_vte_tmerge)

no_vte_vs_vte_vs_recurrence_cph_test_summary = summary(no_vte_vs_vte_vs_recurrence_cph_reg)

no_vte_vs_vte_vs_recurrence_pval_robust = format(
  no_vte_vs_vte_vs_recurrence_cph_test_summary$robscore[["pvalue"]], 
  digits = 1)

no_vte_vs_vte_vs_recurrence_pval_robust
```


```{r}
# jpeg("../graphs/paper/KM_cat_recurrence.jpeg", width = 6.5, height = 4, units = 'in', res = 700)

no_vte_vs_vte_vs_recurrence_km_plot = 
  ggsurvplot(
    fit = no_vte_vs_vte_vs_recurrence_fit,
    data = survival_vte_tmerge,
    xscale = 30,
    break.time.by = 60,
    censor = FALSE,
    conf.int = FALSE,
    xlim = c(0, 540),
    #surv.plot.height = 1.25, 
    #tables.height = 0.25,
    legend.title = " ",
    legend.labs = c("No CAT", "CAT w/o recurrence", "CAT recurrence"),
    risk.table = "abs_pct", 
    risk.table.height = 0.3, 
    fontsize = 3,
    risk.table.col = "strata", 
    tables.y.text = FALSE,
    #tables.y.text.col = TRUE,
    risk.table.y.text.col = TRUE,
    #linetype = "strata", 
    surv.median.line = "hv",
    xlab = "Time (months)",
    ylab = "OS Probability",
    ggtheme = theme_bw(),
    palette = c("black", "red", "#FF6600")
    )

# Count number of patients of each category
n_vte_no_recurrence = sum(oncoth1_data$vte_recurrence == "No VTE recurrence")
n_vte_recurrence = sum(oncoth1_data$vte_recurrence == "VTE recurrence")


# Customize risk table title 
no_vte_vs_vte_vs_recurrence_km_plot$table = 
  no_vte_vs_vte_vs_recurrence_km_plot$table +
  theme(plot.title = element_text(size = 10))

# Add p-value and n of patient subgroups
no_vte_vs_vte_vs_recurrence_km_plot$plot = 
  no_vte_vs_vte_vs_recurrence_km_plot$plot + 
  # Log-rank p-value
  annotate(
    geom = "text", 
    x = 26,
    y = 0.3,
    label = "Log-rank",
    color = "black") + 
  annotate(
    geom = "text", 
    x = 28,
    y = 0.2,
    label = "<0.001",
    color = "black") + 
  # n subgroups
  annotate(
    geom = "text",
    size = 3.5,
    x = 530, 
    y = 0.99, 
    label = "n=302", 
    color = "black") + 
  annotate(
    geom = "text",
    size = 3.5,
    x = 530, 
    y = 0.91, 
    label = paste0("n=", n_vte_no_recurrence), 
    color = "red") +
  annotate(
    geom = "text",
    size = 3.5,
    x = 530, 
    y = 0.83, 
    label = paste0("n=", n_vte_recurrence), 
    color = "#FF6600")

no_vte_vs_vte_vs_recurrence_km_plot

# dev.off()
```


```{r}
cph_recurrence = coxph(Surv(
  time = tstart, 
  time2 = tstop, 
  event = death) ~ 
    age_when_cancer_dx + 
    gender +
    primary_tumor_simplified +
    tnm_stage_grouped +
    vte_event:vte_recurrence + # VTE recurrence only makes sense if there is VTE
    #strata(primary_tumor_simplified) +
    cluster(id),
  id = id,
  data = survival_vte_tmerge) 

cph_recurrence %>% 
  gtsummary::tbl_regression(exp = TRUE)
```


```{r}
#cph_recurrence_test = coxph(Surv(
#  time = tstart, 
#  time2 = tstop, 
#  event = death) ~ 
#    age_when_cancer_dx + 
#    gender +
#    primary_tumor_simplified +
#    tnm_stage_grouped +
#    vte_event:vte_recurrence + # VTE recurrence only makes sense if there is VTE
#    tt(vte_recurrence) +
#    cluster(id),
#    id = id,
#    data = survival_vte_tmerge, 
#    tt = function(x, t, ...) {
#        mtrx <- model.matrix(~x)[,-1]
#        mtrx * (t)
#        }) 

#cph_recurrence_test %>% 
#  gtsummary::tbl_regression(exp = TRUE)
```


```{r}
cox.zph(cph_recurrence, terms = FALSE)
```


```{r}
#ggcoxzph(cox.zph(cph_recurrence, terms = FALSE))[[9]]
```


```{r}
concordance(cph_recurrence)
```


```{r}
# 18-month OS rate
summary(survfit(Surv(
  time = tstart, 
  time2 = tstop, 
  event = death) ~ vte_recurrence + cluster(id), 
  id = id, 
  data = survival_vte_tmerge), times = 18*30.25)
```

## Interactions

Not enough data to compute HR that make sense.  

```{r}
# Age group
cph_recurrence_age = coxph(Surv(
  time = tstart, 
  time2 = tstop, 
  event = death) ~ 
    #age_when_cancer_dx + 
    #gender +
    #primary_tumor_simplified +
    #tnm_stage_grouped +
    strata(gender, primary_tumor_simplified, tnm_stage_grouped) + 
    vte_event:vte_recurrence:age_when_cancer_dx_65 + 
    cluster(id),
  id = id,
  data = survival_vte_tmerge) 

cph_recurrence_age %>% 
  gtsummary::tbl_regression(exp = TRUE)
```


```{r}
cox.zph(cph_recurrence_age, terms = FALSE)
```


```{r}
table(oncoth1_data$vte_recurrence, oncoth1_data$age_when_cancer_dx_65, useNA = "always")
```


```{r}
# Gender
cph_recurrence_gender = coxph(Surv(
  time = tstart, 
  time2 = tstop, 
  event = death) ~ 
    #age_when_cancer_dx + 
    #gender +
    strata(age_when_cancer_dx, primary_tumor_simplified, tnm_stage_grouped) +
    #tnm_stage_grouped +
    vte_event:vte_recurrence:gender + 
    cluster(id),
  id = id,
  data = survival_vte_tmerge) 

cph_recurrence_gender %>% 
  gtsummary::tbl_regression(exp = TRUE)
```

```{r}
cox.zph(cph_recurrence_gender, terms = FALSE)
```



```{r}
table(oncoth1_data$vte_recurrence, oncoth1_data$gender, useNA = "always")
```


```{r}
# Primary tumor
cph_recurrence_primary_tumor = coxph(Surv(
  time = tstart, 
  time2 = tstop, 
  event = death) ~ 
    #age_when_cancer_dx + 
    #gender +
    #primary_tumor_simplified +
    #tnm_stage_grouped +
    strata(age_when_cancer_dx, gender, tnm_stage_grouped) +
    vte_event:vte_recurrence:primary_tumor_simplified + 
    cluster(id),
  id = id,
  data = survival_vte_tmerge) 

cph_recurrence_primary_tumor %>% 
  gtsummary::tbl_regression(exp = TRUE)
```


```{r}
cox.zph(cph_recurrence_primary_tumor, terms = FALSE)
```

NSCLC presents no CAT recurrence. Only the results from pancreatic cancer seem to make some sense, but the calculation of HR is not very reliable due to the low number of recurrences in this subgroup. 


```{r}
table(oncoth1_data$primary_tumor_simplified, oncoth1_data$vte_recurrence, useNA = "always")
```
  


```{r}
# Primary tumor
cph_recurrence_stage = coxph(Surv(
  time = tstart, 
  time2 = tstop, 
  event = death) ~ 
    age_when_cancer_dx + 
    #gender +
    #primary_tumor_simplified +
    strata(gender, primary_tumor_simplified) +
    #tnm_stage_grouped +
    vte_event:vte_recurrence:tnm_stage_two_groups + 
    cluster(id),
  id = id,
  data = survival_vte_tmerge) 

cph_recurrence_stage %>% 
  gtsummary::tbl_regression(exp = TRUE)
```

There is multicollinearity between VTE recurrence and stage.  

```{r}
cox.zph(cph_recurrence_stage, terms = FALSE)
```

```{r}
#ggcoxzph(cox.zph(cph_recurrence_stage, terms = FALSE))
```




