---
title: "Cancer-associated venous thrombosis (CAT) cumulative incidence estimation for ONCOTHROMB12-01 cohort"
author: "Bárbara Lobato Delgado"
output:
  html_document:
    df_print: paged
  pdf_document: default
---


```{r}
# Deactivate scientific notation
options(scipen = 999, na.action = "na.pass", max.print = 10000)
```


```{r, message=FALSE}
# Load libraries
pacman::p_load(
  an9elproject, 
  tidyverse, 
  lubridate, 
  survival, 
  pec, 
  survminer, 
  cmprsk, 
  install = FALSE, update = FALSE
)
```


# Introduction

In this analysis we estimate the cancer-associated venous thrombosis (CAT) cumulative incidence in competing risks with death, and plot the cumulative incidence over time for the ONCOTHROMB12-01 cohort.

# Cumulative incidence estimation with competing risks — death *vs* CAT

Death and CAT are events in competing risks, given that death impedes CAT — but not the other way around. 

Cumulative incidence describes the probability of a specific event occurring within a certain time frame. It is often used in the context of competing risks, where there are multiple possible events that can occur, and the focus is on estimating the probability of each event type.

In the context of survival analysis with competing risks, the cumulative incidence function (CIF) is used to represent the cumulative probability of experiencing a particular type of event before a given time point, while accounting for the presence of other competing events that could prevent the occurrence of the event of interest.

Mathematically, the cumulative incidence function for a specific event type is defined as:

CIF(t) = P(Event of interest occurs before or at time t)

The cumulative incidence at a specific time point represents the proportion of individuals who have experienced the event of interest up to that time. It's important to note that as time progresses, the cumulative incidence can increase, but it cannot decrease.

When there are competing risks, each event type will have its own cumulative incidence function. *The sum of the cumulative incidence functions for all event types at any given time point cannot exceed 1, as it represents the entire probability space.*

Cumulative incidence functions are commonly used to compare event rates between different groups or categories, and statistical tests like Gray's test can be used to assess whether there are significant differences in the cumulative incidence functions across groups. Gray's test helps you determine if there are statistically significant differences between these CIFs across different groups. It's an extension of the classical log-rank test used in traditional survival analysis, adapted to the competing risks setting.

```{r}
# Load DB from an9elproject
# oncoth1 = get_project("oncothr1", version = "0.0.8003")
load("/mnt/ir-bioinf02/home/blobato/oncothromb01/oncothr1_versions/oncothr1.RData_0.0.8003")
oncoth1 = obj
```


```{r}
# Create dataset with information for estimation con cumulative incidence
cat_cuminc_data = oncoth1$data %>% 
  select(id, 
         patient_group,
         primary_tumor_simplified,
         n_appointment_patient_became_case,
         date_cancer_dx, 
         date_dropout, 
         date_death,
         date_study_begins, 
         date_study_ends, 
         patient_status_at_end_study_simplified, 
         follow_up_length_from_cancer_diagnosis_days, 
         tu_1st_vte_all_from_cancer_diagnosis_days_corrected) %>%
  # Patients that already suffered CAT at time 0 or prior must be excluded
  filter(!n_appointment_patient_became_case == "Beginning of study")

```

A total of 30 patients are excluded in order to estimate CAT cumulative incidence, given that they already suffered CAT at time = 0.  

```{r}
# Now create variables of status (CAT, death w/o CAT or censored) and time until event (CAT or death w/o CAT)
cat_cuminc_data %<>%
  # Status (0 for censored, 1 for CAT, 2 for dead with no CAT)
  mutate(event_status = case_when(
    patient_group == "Control" & 
      patient_status_at_end_study_simplified == "Dead" ~ 2, 
    patient_group == "Case" ~ 1, 
    TRUE ~ 0)) %>%
  # Time until event
  mutate(tu_event = case_when(
    event_status == 0 ~ follow_up_length_from_cancer_diagnosis_days, # censored patients
    event_status == 2 ~ follow_up_length_from_cancer_diagnosis_days, # patients dead w/o CAT
    event_status == 1 ~ tu_1st_vte_all_from_cancer_diagnosis_days_corrected # patients with CAT
  ))
```


```{r}
# Estimate CAT and death cumulative incidence functions from competing risks data
cat_cuminc = cuminc(
  ftime = cat_cuminc_data$tu_event, 
  fstatus = cat_cuminc_data$event_status, 
  cencode = 0 # censored patients are encoded as zero
  )

# $est: cumulative incidence estimates at different time points for each type of event (days)
# i is the group index and j is the event
# $var: variances of the cumulative incidence estimates
cat_cuminc
```

These are the cumulative incidence functions (CIFs) for developing CAT (1 1) and dying without CAT (1 2) and their respective variances. In spite of the notation, patients are not grouped. 

```{r}
# Estimate CAT and death cumulative estimates functions (CIFs) for group of patients with CAT and w/o CAT
# Get Tests
cat_cuminc_test = cuminc(
  ftime = cat_cuminc_data$tu_event, 
  fstatus = cat_cuminc_data$event_status, 
  group = cat_cuminc_data$event_status, # patients with CAT vs patients dead w/o CAT
  cencode = 0)[["Tests"]]

# Show the test statistic value, the p-value and the degrees of freedom resulting from comparison of CIFs between groups
cat_cuminc_test
```

Group 1 are patients with CAT and group 2 are patients that die w/o CAT. 

```{r}
# Show results in a more intelligible way
cat_cuminc_customised <- cat_cuminc %>%
  # Apply a function to each element of the list or atomic vector and show in a dataframe
  map_df(`[`, c("time", "est"), .id = "id") %>%
  # Add a simplified name for groups using case_match()
  mutate(id = case_match(
    id, 
    "1 1" ~ "CAT",
    "1 2" ~ "Death w/o CAT"
  )) %>%
  # Select event of interest, in this case, CAT
  filter(id == "CAT")


# cat_cuminc_customised = cat_cuminc %>% 
#   # Remove Tests from list
#   list_modify("Tests" = NULL) %>% 
#   # Apply a function to each element of list or atomic vector and show in a dataframe
#   # Subset "time" and "est" and add id column
#   map_df(`[`, c("time", "est"), .id = "id") %>% 
#   # Add name to groups
#   mutate(id = recode(
#     id, 
#     "1 1" = "CAT", 
#     "1 2" = "Death w/o CAT")) %>% 
#   # Select event of interest, in this case, CAT
#   filter(id == "CAT")  

# All time points in CAT CIF estimation
cat_cuminc_customised
```


```{r}
# Estimation of cumulative incidence with competing risks
# Similar information as above, but providing standard error and 95% CI
tidycmprsk::cuminc(Surv(tu_event, as.factor(event_status)) ~ 1, cat_cuminc_data)
```


```{r}
# Create survival curve for CAT in competing risks
# Covariates are not being treated as time-dependent
cat_fit <- survfit(
  Surv(tu_event, event_status == 1) ~ 1 + cluster(id), 
  data = cat_cuminc_data
)

# I can use this table to 
cat_fit_summary = surv_summary(cat_fit)
cat_fit_summary 
```

```{r}
get_cumulative_incidence = function(fit_summary, event_code = 1){
  
  # Function to print the cumulative incidence and 95% confidence interval
  # of CAT from summary of fit object
  
  # CAT cumulative incidence at given times
  # Time in days
  twelve_months_in_days <- 12 * 30.25
  eighteen_months_in_days <- 18 * 30.25
  window_size <- 10
  
  # 12-month CAT cumulative incidence
  cuminc_12month_stats <- fit_summary %>% 
    filter(n.event == event_code) %>%
    slice(which.min(abs(time - twelve_months_in_days)))
  
  print(paste("12-month CAT cumulative incidence is", round(1 - cuminc_12month_stats$surv, digits = 4) * 100, "% (CI", round(1 - cuminc_12month_stats$upper, digits = 4) * 100, "% -", round(1 - cuminc_12month_stats$lower, digits = 4) * 100, "%)"))
  
  
  # 18-month CAT cumulative incidence
  cuminc_18month_stats <- fit_summary %>% 
    filter(n.event == event_code) %>%
    slice(which.min(abs(time - eighteen_months_in_days)))
  
  print(paste("18-month CAT cumulative incidence is", round(1 - cuminc_18month_stats$surv, digits = 4) * 100, "% (CI", round(1 - cuminc_18month_stats$upper, digits = 4) * 100, "% -", round(1 - cuminc_18month_stats$lower, digits = 4) * 100, "%)"))
}
```


```{r}
# Print CAT cumulative incidence at 12 and 18 months for ONCOTHROMB1 cohort
get_cumulative_incidence(cat_fit_summary)
```


```{r}
# Count number of patients affected by CAT 
n_cat_patients = sum(cat_fit_summary$n.event)

print(paste("A total", n_cat_patients, "patients suffered CAT. Remind that VTE concomitant to cancer diagnosis is not accounted for in the cumulative incidence estimation."))

# There is a total of 88 patients that suffered CAT in the ONCOTHROMB1 cohort.
# Thirty patients are excluded for the cumulative incidence estimation
# However, the rest must be accounted for
stopifnot(n_cat_patients + sum(oncoth1$data$n_appointment_patient_became_case == "Beginning of study") == sum(oncoth1$data$patient_group == "Case"))
```


```{r}
# Plot cumulative incidence curve
cat_cuminc_with_table = ggsurvplot(
  fit = cat_fit,
  data = cat_cuminc_data,
  fun = "event",
  censor = FALSE,
  conf.int = FALSE,
  xscale = 30,
  break.time.by = 60,
  surv.plot.height = 1.5,
  cumevents = TRUE,
  cumevents.y.text = FALSE,
  cumevents.height = 0.15,
  #tables.col = "strata",
  fontsize = 3.5,
  tables.y.text = FALSE, # color rectangle instead of text with strata name
  tables.theme = theme_cleantable(),
  legend = "none",
  xlab = "Time (months)",
  ylab = "Cumulative incidence of CAT", 
  title = "CAT cumulative incidence for ONCOTHROMB12-01 cohort",
  # Extend until month 20 so that all 59 VTE events can be seen in plot table
  xlim = c(0, 600), # 540
  ylim = c(0, 0.2),
  ggtheme = theme_bw(),
  palette = "red"
  ) 

# Add percentages in y axis
cat_cuminc_with_table$plot = cat_cuminc_with_table$plot +
  scale_y_continuous(labels = scales::percent, limits = c(0, 0.2))

# Customize risk table title 
# Add new title to table
cat_cuminc_with_table$cumevents$labels$title = "Cumulative number of patients affected by CAT"
# Change table title font size
cat_cuminc_with_table$cumevents = cat_cuminc_with_table$cumevents +
  theme(plot.title = element_text(size = 10))

cat_cuminc_with_table
```

```{r}
# Save plot with table

# Combine the main plot and the cumulative events table
cat_cum_inc_combined_plot <- cowplot::plot_grid(
  cat_cuminc_with_table$plot,
  cat_cuminc_with_table$cumevents,
  ncol = 1, 
  rel_heights = c(1.5, 0.3)  # Adjust heights to fit both plots
)

# Save plot
ggsave(
  filename = "../graphs/CAT_cum_inc.jpeg",
  plot = cat_cum_inc_combined_plot, 
  width = 6, 
  height = 4,
  units = "in"
  )
```


## CAT cumulative incidence per cancer type

We want to estimate CAT cumulative incidence for each type of cancer. 

```{r}
# Account for patients with concomitant VTEs per cancer type
conc_VTE_per_cancer_type <- table(oncoth1$data$primary_tumor_simplified, oncoth1$data$vte_before_entering_study)
conc_VTE_per_cancer_type <- as.data.frame(conc_VTE_per_cancer_type)
colnames(conc_VTE_per_cancer_type) <- c("cancer_type", "conc_VTE", "n_patients_conc_VTE")
conc_VTE_per_cancer_type %<>% 
  select(-conc_VTE)
conc_VTE_per_cancer_type
```

### Colorectal cancer

```{r}
# Subset patients with colorectal cancer
cat_cuminc_data_colorectal <- cat_cuminc_data %>%
  filter(primary_tumor_simplified == "Colorectal")
```


```{r}
# Estimate CAT and death cumulative incidence functions from competing risks data
cat_cuminc_colorectal <- cuminc(
  ftime = cat_cuminc_data_colorectal$tu_event, 
  fstatus = cat_cuminc_data_colorectal$event_status, 
  cencode = 0 # censored patients are encoded as zero
  )

# $est: cumulative incidence estimates at different time points for each type of event (days)
# i is the group index and j is the event
# $var: variances of the cumulative incidence estimates
cat_cuminc_colorectal
```

These are the cumulative incidence functions (CIFs) for developing CAT (1 1) and dying without CAT (1 2) and their respective variances for cancer patients in ONCOTHROMB1. In spite of the notation, patients are not grouped. 

```{r}
# Estimate CAT and death cumulative estimates functions (CIFs) for group of patients with CAT and w/o CAT
# Get Tests
cat_cuminc_test_colorectal <- cuminc(
  ftime = cat_cuminc_data_colorectal$tu_event, 
  fstatus = cat_cuminc_data_colorectal$event_status, 
  group = cat_cuminc_data_colorectal$event_status, # patients with CAT vs patients dead w/o CAT
  cencode = 0)[["Tests"]]

# Show the test statistic value, the p-value and the degrees of freedom resulting from comparison of CIFs between groups
cat_cuminc_test_colorectal
```

Group 1 are colorectal cancer patients with CAT and group 2 are patients that die w/o CAT. 

```{r}
# Show results in a more intelligible way
cat_cuminc_customised_colorectal <- cat_cuminc_colorectal %>%
  # Apply a function to each element of the list or atomic vector and show in a dataframe
  map_df(`[`, c("time", "est"), .id = "id") %>%
  # Add a simplified name for groups using case_match()
  mutate(id = case_match(
    id, 
    "1 1" ~ "CAT",
    "1 2" ~ "Death w/o CAT"
  )) %>%
  # Select event of interest, in this case, CAT
  filter(id == "CAT")

# All time points in CAT CIF estimation
cat_cuminc_customised_colorectal
```

```{r}
# CAT cumulative incidence at given times for colorectal cancer patients
# Time in days

# 12-month CAT cumulative incidence
# Inspect table and choose incidence approximate to specific time
cat_cuminc_customised_colorectal 
```


```{r}
# Estimation of cumulative incidence with competing risks for colorectal cancer patients
# Similar information as above, but providing standard error and 95% CI
tidycmprsk::cuminc(Surv(tu_event, as.factor(event_status)) ~ 1, cat_cuminc_data_colorectal)
```

```{r}
# Create survival curve for CAT in competing risks
# Covariates are not being treated as time-dependent
cat_fit_colorectal <- survfit(
  Surv(tu_event, event_status == 1) ~ 1 + cluster(id), 
  data = cat_cuminc_data_colorectal
)

# I can use this table to 
cat_fit_summary_colorectal = surv_summary(cat_fit_colorectal)
cat_fit_summary_colorectal
```

```{r}
cat_fit_summary_colorectal %>% 
  filter(n.event == 1)
```



```{r}
# Print CAT cumulative incidence at 12 and 18 months for colorectal cancer patients in ONCOTHROMB1 cohort
get_cumulative_incidence(cat_fit_summary_colorectal)
```


```{r}
# Count number of patients affected by CAT 
n_cat_colorectal_patients = sum(cat_fit_summary_colorectal$n.event)

print(paste("A total", n_cat_colorectal_patients, "colorectal cancer patients suffered CAT. Remind that VTE concomitant to cancer diagnosis is not accounted for in the cumulative incidence estimation."))

# There is a total of 88 patients that suffered CAT in the ONCOTHROMB1 cohort.
# Thirty patients are excluded for the cumulative incidence estimation
# However, the rest must be accounted for
stopifnot(n_cat_colorectal_patients + conc_VTE_per_cancer_type %>%filter(cancer_type == "Colorectal") %>% select(n_patients_conc_VTE) %>% unname() == sum(cat_cuminc_data$patient_group == "Case"))
```


```{r}
# Plot cumulative incidence curve
cat_cuminc_colorectal_with_table = ggsurvplot(
  fit = cat_fit_colorectal,
  data = cat_cuminc_data_colorectal,
  fun = "event",
  censor = FALSE,
  conf.int = FALSE,
  xscale = 30,
  break.time.by = 60,
  surv.plot.height = 1.5,
  cumevents = TRUE,
  cumevents.y.text = FALSE,
  cumevents.height = 0.15,
  #tables.col = "strata",
  fontsize = 3.5,
  tables.y.text = FALSE, # color rectangle instead of text with strata name
  tables.theme = theme_cleantable(),
  legend = "none",
  xlab = "Time (months)",
  ylab = "Cumulative incidence of CAT", 
  title = "CAT cumulative incidence for colorectal cancer patients",
  # Extend until month 20 so that all 59 VTE events can be seen in plot table
  xlim = c(0, 600), # 540
  ylim = c(0, 0.2),
  ggtheme = theme_bw(),
  palette = "red"
  ) 

# Add percentages in y axis
cat_cuminc_colorectal_with_table$plot = cat_cuminc_colorectal_with_table$plot +
  scale_y_continuous(labels = scales::percent, limits = c(0, 0.2))

# Customize risk table title 
# Add new title to table
cat_cuminc_colorectal_with_table$cumevents$labels$title = "Cumulative number of colorectal cancer patients affected by CAT"
# Change table title font size
cat_cuminc_colorectal_with_table$cumevents = cat_cuminc_colorectal_with_table$cumevents +
  theme(plot.title = element_text(size = 10))

cat_cuminc_colorectal_with_table
```

```{r}
# Save plot with table

# Combine the main plot and the cumulative events table
cat_cum_inc_colorectal_combined_plot <- cowplot::plot_grid(
  cat_cuminc_colorectal_with_table$plot,
  cat_cuminc_colorectal_with_table$cumevents,
  ncol = 1, 
  rel_heights = c(1.5, 0.3)  # Adjust heights to fit both plots
)

# Save plot
ggsave(
  filename = "../graphs/CAT_cum_inc_colorectal.jpeg",
  plot = cat_cum_inc_colorectal_combined_plot, 
  width = 6, 
  height = 4,
  units = "in"
  )
```


### NSCLC

```{r}
# Subset patients with NSCLC
cat_cuminc_data_nsclc <- cat_cuminc_data %>%
  filter(primary_tumor_simplified == "NSCLC")
```


```{r}
# Estimate CAT and death cumulative incidence functions from competing risks data
cat_cuminc_nsclc <- cuminc(
  ftime = cat_cuminc_data_nsclc$tu_event, 
  fstatus = cat_cuminc_data_nsclc$event_status, 
  cencode = 0 # censored patients are encoded as zero
  )

# $est: cumulative incidence estimates at different time points for each type of event (days)
# i is the group index and j is the event
# $var: variances of the cumulative incidence estimates
cat_cuminc_nsclc
```

These are the cumulative incidence functions (CIFs) for developing CAT (1 1) and dying without CAT (1 2) and their respective variances for cancer patients in ONCOTHROMB1. In spite of the notation, patients are not grouped. 

```{r}
# Estimate CAT and death cumulative estimates functions (CIFs) for group of patients with CAT and w/o CAT
# Get Tests
cat_cuminc_test_nsclc <- cuminc(
  ftime = cat_cuminc_data_nsclc$tu_event, 
  fstatus = cat_cuminc_data_nsclc$event_status, 
  group = cat_cuminc_data_nsclc$event_status, # patients with CAT vs patients dead w/o CAT
  cencode = 0)[["Tests"]]

# Show the test statistic value, the p-value and the degrees of freedom resulting from comparison of CIFs between groups
cat_cuminc_test_nsclc
```

Group 1 are NSCLC patients with CAT and group 2 are patients that die w/o CAT. 

```{r}
# Show results in a more intelligible way
cat_cuminc_customised_nsclc <- cat_cuminc_nsclc %>%
  # Apply a function to each element of the list or atomic vector and show in a dataframe
  map_df(`[`, c("time", "est"), .id = "id") %>%
  # Add a simplified name for groups using case_match()
  mutate(id = case_match(
    id, 
    "1 1" ~ "CAT",
    "1 2" ~ "Death w/o CAT"
  )) %>%
  # Select event of interest, in this case, CAT
  filter(id == "CAT")

# All time points in CAT CIF estimation
cat_cuminc_customised_nsclc
```


```{r}
# CAT cumulative incidence at given times for NSCLC patients
# Time in days

# 12-month CAT cumulative incidence
# Inspect table and choose incidence approximate to specific time
cat_cuminc_customised_nsclc 
```


```{r}
# Estimation of cumulative incidence with competing risks for NSCLC patients
# Similar information as above, but providing standard error and 95% CI
tidycmprsk::cuminc(Surv(tu_event, as.factor(event_status)) ~ 1, cat_cuminc_data_nsclc)
```


```{r}
# Create survival curve for CAT in competing risks
# Covariates are not being treated as time-dependent
cat_fit_nsclc <- survfit(
  Surv(tu_event, event_status == 1) ~ 1 + cluster(id), 
  data = cat_cuminc_data_nsclc
)

# I can use this table to 
cat_fit_summary_nsclc = surv_summary(cat_fit_nsclc)
cat_fit_summary_nsclc
```


```{r}
# Print CAT cumulative incidence at 12 and 18 months for NSCLC patients in ONCOTHROMB1 cohort
get_cumulative_incidence(cat_fit_summary_nsclc)
```


```{r}
# Count number of patients affected by CAT 
n_cat_nsclc_patients = sum(cat_fit_summary_nsclc$n.event)

print(paste("A total", n_cat_nsclc_patients, "NSCLC patients suffered CAT. Remind that VTE concomitant to cancer diagnosis is not accounted for in the cumulative incidence estimation."))

# There is a total of 88 patients that suffered CAT in the ONCOTHROMB1 cohort.
# Thirty patients are excluded for the cumulative incidence estimation
# However, the rest must be accounted for
stopifnot(n_cat_nsclc_patients + conc_VTE_per_cancer_type %>%filter(cancer_type == "NSCLC") %>% select(n_patients_conc_VTE) %>% unname() == sum(cat_cuminc_data$patient_group == "Case"))
```


```{r}
# Plot cumulative incidence curve
cat_cuminc_nsclc_with_table = ggsurvplot(
  fit = cat_fit_nsclc,
  data = cat_cuminc_data_nsclc,
  fun = "event",
  censor = FALSE,
  conf.int = FALSE,
  xscale = 30,
  break.time.by = 60,
  surv.plot.height = 1.5,
  cumevents = TRUE,
  cumevents.y.text = FALSE,
  cumevents.height = 0.15,
  #tables.col = "strata",
  fontsize = 3.5,
  tables.y.text = FALSE, # color rectangle instead of text with strata name
  tables.theme = theme_cleantable(),
  legend = "none",
  xlab = "Time (months)",
  ylab = "Cumulative incidence of CAT", 
  title = "CAT cumulative incidence for NSCLC cancer patients",
  # Extend until month 20 so that all 59 VTE events can be seen in plot table
  xlim = c(0, 600), # 540
  ylim = c(0, 0.2),
  ggtheme = theme_bw(),
  palette = "red"
  ) 

# Add percentages in y axis
cat_cuminc_nsclc_with_table$plot = cat_cuminc_nsclc_with_table$plot +
  scale_y_continuous(labels = scales::percent, limits = c(0, 0.2))

# Customize risk table title 
# Add new title to table
cat_cuminc_nsclc_with_table$cumevents$labels$title = "Cumulative number of NSCLC patients affected by CAT"
# Change table title font size
cat_cuminc_nsclc_with_table$cumevents = cat_cuminc_nsclc_with_table$cumevents +
  theme(plot.title = element_text(size = 10))

cat_cuminc_nsclc_with_table
```


```{r}
# Save plot with table

# Combine the main plot and the cumulative events table
cat_cum_inc_nsclc_combined_plot <- cowplot::plot_grid(
  cat_cuminc_nsclc_with_table$plot,
  cat_cuminc_nsclc_with_table$cumevents,
  ncol = 1, 
  rel_heights = c(1.5, 0.3)  # Adjust heights to fit both plots
)

# Save plot
ggsave(
  filename = "../graphs/CAT_cum_inc_nsclc.jpeg",
  plot = cat_cum_inc_nsclc_combined_plot, 
  width = 6, 
  height = 4,
  units = "in"
  )
```


### Esophago-gastric cancer

```{r}
# Subset patients with esophagogastric cancer
cat_cuminc_data_esophagogastric <- cat_cuminc_data %>%
  filter(primary_tumor_simplified == "Oesophago-gastric")
```


```{r}
# Estimate CAT and death cumulative incidence functions from competing risks data
cat_cuminc_esophagogastric <- cuminc(
  ftime = cat_cuminc_data_esophagogastric$tu_event, 
  fstatus = cat_cuminc_data_esophagogastric$event_status, 
  cencode = 0 # censored patients are encoded as zero
  )

# $est: cumulative incidence estimates at different time points for each type of event (days)
# i is the group index and j is the event
# $var: variances of the cumulative incidence estimates
cat_cuminc_esophagogastric
```

These are the cumulative incidence functions (CIFs) for developing CAT (1 1) and dying without CAT (1 2) and their respective variances for esophago-gastric cancer patients in ONCOTHROMB1. In spite of the notation, patients are not grouped. 

```{r}
# Estimate CAT and death cumulative estimates functions (CIFs) for group of patients with CAT and w/o CAT
# Get Tests
cat_cuminc_test_esophagogastric <- cuminc(
  ftime = cat_cuminc_data_esophagogastric$tu_event, 
  fstatus = cat_cuminc_data_esophagogastric$event_status, 
  group = cat_cuminc_data_esophagogastric$event_status, # patients with CAT vs patients dead w/o CAT
  cencode = 0)[["Tests"]]

# Show the test statistic value, the p-value and the degrees of freedom resulting from comparison of CIFs between groups
cat_cuminc_test_esophagogastric
```

Group 1 are esophago-gastric cancer patients with CAT and group 2 are patients that die w/o CAT. 

```{r}
# Show results in a more intelligible way
cat_cuminc_customised_esophagogastric <- cat_cuminc_esophagogastric %>%
  # Apply a function to each element of the list or atomic vector and show in a dataframe
  map_df(`[`, c("time", "est"), .id = "id") %>%
  # Add a simplified name for groups using case_match()
  mutate(id = case_match(
    id, 
    "1 1" ~ "CAT",
    "1 2" ~ "Death w/o CAT"
  )) %>%
  # Select event of interest, in this case, CAT
  filter(id == "CAT")

# All time points in CAT CIF estimation
cat_cuminc_customised_esophagogastric
```

```{r}
# CAT cumulative incidence at given times for esophago-gastric cancer patients
# Time in days

# 12-month CAT cumulative incidence
# Inspect table and choose incidence approximate to specific time
cat_cuminc_customised_esophagogastric 
```


```{r}
# Estimation of cumulative incidence with competing risks for esophago-gastric cancer patients
# Similar information as above, but providing standard error and 95% CI
tidycmprsk::cuminc(Surv(tu_event, as.factor(event_status)) ~ 1, cat_cuminc_data_esophagogastric)
```


```{r}
# Create survival curve for CAT in competing risks
# Covariates are not being treated as time-dependent
cat_fit_esophagogastric <- survfit(
  Surv(tu_event, event_status == 1) ~ 1 + cluster(id), 
  data = cat_cuminc_data_esophagogastric
)

# I can use this table to 
cat_fit_summary_esophagogastric = surv_summary(cat_fit_esophagogastric)
cat_fit_summary_esophagogastric
```


```{r}
# Print CAT cumulative incidence at 12 and 18 months for esophago-gastric cancer patients in ONCOTHROMB1 cohort
get_cumulative_incidence(cat_fit_summary_esophagogastric)
```


```{r}
# Count number of patients affected by CAT 
n_cat_esophagogastric_patients = sum(cat_fit_summary_esophagogastric$n.event)

print(paste("A total", n_cat_esophagogastric_patients, "esophago-gastric cancer patients suffered CAT. Remind that VTE concomitant to cancer diagnosis is not accounted for in the cumulative incidence estimation."))

# There is a total of 88 patients that suffered CAT in the ONCOTHROMB1 cohort.
# Thirty patients are excluded for the cumulative incidence estimation
# However, the rest must be accounted for
stopifnot(n_cat_esophagogastric_patients + conc_VTE_per_cancer_type %>%filter(cancer_type == "Oesophago-gastric") %>% select(n_patients_conc_VTE) %>% unname() == sum(cat_cuminc_data$patient_group == "Case"))
```


```{r}
# Plot cumulative incidence curve
cat_cuminc_esophagogastric_with_table = ggsurvplot(
  fit = cat_fit_esophagogastric,
  data = cat_cuminc_data_esophagogastric,
  fun = "event",
  censor = FALSE,
  conf.int = FALSE,
  xscale = 30,
  break.time.by = 60,
  surv.plot.height = 1.5,
  cumevents = TRUE,
  cumevents.y.text = FALSE,
  cumevents.height = 0.15,
  #tables.col = "strata",
  fontsize = 3.5,
  tables.y.text = FALSE, # color rectangle instead of text with strata name
  tables.theme = theme_cleantable(),
  legend = "none",
  xlab = "Time (months)",
  ylab = "Cumulative incidence of CAT", 
  title = "CAT cumulative incidence for esophago-gastric cancer patients",
  # Extend until month 20 so that all 59 VTE events can be seen in plot table
  xlim = c(0, 600), # 540
  ylim = c(0, 0.2),
  ggtheme = theme_bw(),
  palette = "red"
  ) 

# Add percentages in y axis
cat_cuminc_esophagogastric_with_table$plot = cat_cuminc_esophagogastric_with_table$plot +
  scale_y_continuous(labels = scales::percent, limits = c(0, 0.2))

# Customize risk table title 
# Add new title to table
cat_cuminc_esophagogastric_with_table$cumevents$labels$title = "Cumulative number of esophago-gastric cancer patients affected by CAT"
# Change table title font size
cat_cuminc_esophagogastric_with_table$cumevents = cat_cuminc_esophagogastric_with_table$cumevents +
  theme(plot.title = element_text(size = 10))

cat_cuminc_esophagogastric_with_table
```


```{r}
# Save plot with table

# Combine the main plot and the cumulative events table
cat_cum_inc_esophagogastric_combined_plot <- cowplot::plot_grid(
  cat_cuminc_esophagogastric_with_table$plot,
  cat_cuminc_esophagogastric_with_table$cumevents,
  ncol = 1, 
  rel_heights = c(1.5, 0.3)  # Adjust heights to fit both plots
)

# Save plot
ggsave(
  filename = "../graphs/CAT_cum_inc_esophagogastric.jpeg",
  plot = cat_cum_inc_esophagogastric_combined_plot, 
  width = 6, 
  height = 4,
  units = "in"
  )
```



### Pancreatic cancer

```{r}
# Subset patients with pancreatic cancer
cat_cuminc_data_pancreatic <- cat_cuminc_data %>%
  filter(primary_tumor_simplified == "Pancreatic")
```


```{r}
# Estimate CAT and death cumulative incidence functions from competing risks data
cat_cuminc_pancreatic <- cuminc(
  ftime = cat_cuminc_data_pancreatic$tu_event, 
  fstatus = cat_cuminc_data_pancreatic$event_status, 
  cencode = 0 # censored patients are encoded as zero
  )

# $est: cumulative incidence estimates at different time points for each type of event (days)
# i is the group index and j is the event
# $var: variances of the cumulative incidence estimates
cat_cuminc_pancreatic
```

These are the cumulative incidence functions (CIFs) for developing CAT (1 1) and dying without CAT (1 2) and their respective variances for pancreatic cancer patients in ONCOTHROMB1. In spite of the notation, patients are not grouped. 

```{r}
# Estimate CAT and death cumulative estimates functions (CIFs) for group of patients with CAT and w/o CAT
# Get Tests
cat_cuminc_test_pancreatic <- cuminc(
  ftime = cat_cuminc_data_pancreatic$tu_event, 
  fstatus = cat_cuminc_data_pancreatic$event_status, 
  group = cat_cuminc_data_pancreatic$event_status, # patients with CAT vs patients dead w/o CAT
  cencode = 0)[["Tests"]]

# Show the test statistic value, the p-value and the degrees of freedom resulting from comparison of CIFs between groups
cat_cuminc_test_pancreatic
```

Group 1 are pancreatic cancer patients with CAT and group 2 are patients that die w/o CAT. 

```{r}
# Show results in a more intelligible way
cat_cuminc_customised_pancreatic <- cat_cuminc_pancreatic %>%
  # Apply a function to each element of the list or atomic vector and show in a dataframe
  map_df(`[`, c("time", "est"), .id = "id") %>%
  # Add a simplified name for groups using case_match()
  mutate(id = case_match(
    id, 
    "1 1" ~ "CAT",
    "1 2" ~ "Death w/o CAT"
  )) %>%
  # Select event of interest, in this case, CAT
  filter(id == "CAT")

# All time points in CAT CIF estimation
cat_cuminc_customised_pancreatic
```


```{r}
# CAT cumulative incidence at given times for pancreatic cancer patients
# Time in days

# 12-month CAT cumulative incidence
# Inspect table and choose incidence approximate to specific time
cat_cuminc_customised_pancreatic 
```


```{r}
# Estimation of cumulative incidence with competing risks for pancreatic cancer patients
# Similar information as above, but providing standard error and 95% CI
tidycmprsk::cuminc(Surv(tu_event, as.factor(event_status)) ~ 1, cat_cuminc_data_pancreatic)
```


```{r}
# Create survival curve for CAT in competing risks
# Covariates are not being treated as time-dependent
cat_fit_pancreatic <- survfit(
  Surv(tu_event, event_status == 1) ~ 1 + cluster(id), 
  data = cat_cuminc_data_pancreatic
)

# I can use this table to 
cat_fit_summary_pancreatic = surv_summary(cat_fit_pancreatic)
cat_fit_summary_pancreatic
```


```{r}
# Print CAT cumulative incidence at 12 and 18 months for pancreatic cancer patients in ONCOTHROMB1 cohort
get_cumulative_incidence(cat_fit_summary_pancreatic)
```


```{r}
# Count number of patients affected by CAT 
n_cat_pancreatic_patients = sum(cat_fit_summary_pancreatic$n.event)

print(paste("A total", n_cat_pancreatic_patients, "pancreatic cancer patients suffered CAT. Remind that VTE concomitant to cancer diagnosis is not accounted for in the cumulative incidence estimation."))

# There is a total of 88 patients that suffered CAT in the ONCOTHROMB1 cohort.
# Thirty patients are excluded for the cumulative incidence estimation
# However, the rest must be accounted for
stopifnot(n_cat_pancreatic_patients + conc_VTE_per_cancer_type %>%filter(cancer_type == "Pancreatic") %>% select(n_patients_conc_VTE) %>% unname() == sum(cat_cuminc_data$patient_group == "Case"))
```


```{r}
# Plot cumulative incidence curve
cat_cuminc_pancreatic_with_table = ggsurvplot(
  fit = cat_fit_pancreatic,
  data = cat_cuminc_data_pancreatic,
  fun = "event",
  censor = FALSE,
  conf.int = FALSE,
  xscale = 30,
  break.time.by = 60,
  surv.plot.height = 1.5,
  cumevents = TRUE,
  cumevents.y.text = FALSE,
  cumevents.height = 0.15,
  #tables.col = "strata",
  fontsize = 3.5,
  tables.y.text = FALSE, # color rectangle instead of text with strata name
  tables.theme = theme_cleantable(),
  legend = "none",
  xlab = "Time (months)",
  ylab = "Cumulative incidence of CAT", 
  title = "CAT cumulative incidence for pancreatic cancer patients",
  # Extend until month 20 so that all 59 VTE events can be seen in plot table
  xlim = c(0, 600), # 540
  ylim = c(0, 0.4),
  ggtheme = theme_bw(),
  palette = "red"
  ) 

# Add percentages in y axis
cat_cuminc_pancreatic_with_table$plot = cat_cuminc_pancreatic_with_table$plot +
  scale_y_continuous(labels = scales::percent, limits = c(0, 0.4))

# Customize risk table title 
# Add new title to table
cat_cuminc_pancreatic_with_table$cumevents$labels$title = "Cumulative number of pancreatic cancer patients affected by CAT"
# Change table title font size
cat_cuminc_pancreatic_with_table$cumevents = cat_cuminc_pancreatic_with_table$cumevents +
  theme(plot.title = element_text(size = 10))

cat_cuminc_pancreatic_with_table
```


```{r}
# Save plot with table

# Combine the main plot and the cumulative events table
cat_cum_inc_pancreatic_combined_plot <- cowplot::plot_grid(
  cat_cuminc_pancreatic_with_table$plot,
  cat_cuminc_pancreatic_with_table$cumevents,
  ncol = 1, 
  rel_heights = c(1.5, 0.3)  # Adjust heights to fit both plots
)

# Save plot
ggsave(
  filename = "../graphs/CAT_cum_inc_pancreatic.jpeg",
  plot = cat_cum_inc_pancreatic_combined_plot, 
  width = 6, 
  height = 4,
  units = "in"
  )
```


